### 2014-10-18 draft.tex
Initializing
Start of Draft

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage[margin=0.9in]{geometry}
%\usepackage{amsfonts}
%\usepackage{amsthm}
\usepackage{subfigure}
\usepackage{mathtools}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{cite}
\usepackage{hyperref}
\usepackage{url}
\urlstyle{same}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\makeatletter
\let\c@lemma=\c@theorem
\let\c@corollary=\c@theorem
\let\c@fact=\c@theorem
\makeatother

\let\realendproof=\endproof
\def\endproof{\hspace*{\fill}$\Box$\realendproof}
\title{Self Referencing Sequences}
\author{Perry Kleinhenz, Fermi Ma and Erik Waingarten}
\date{}							% Activate to display a given date or no date

\begin{document}

\title{Self Referencing Sequences}
\titlerunning{Self Referencing Sequences}

\author{Perry Kleinhenz \and Fermi Ma \and Erik Waingarten}
%
\authorrunning{Perry Kleinhenz \and Fermi Ma \and Erik Waingarten}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{
\protect\url{{pkleinhe, fermima,eaw}@mit.edu}}

\maketitle

\section{Introduction}

\section{Definitions and Notations}
In this section we introduce

\section{Prefixes}
\emph{Written by Perry Kleinhenz, edited by Fermi Ma and Erik Waingarten}
\begin{definition} We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a prefix of length $n$ if 
\begin{equation*}
a_{n+i} = b_{i},
\end{equation*}
for $i$ any positive integer. if no such $n$ exists then we say that the two sequences are independent.
\end{definition}
Note that a pair of sequences can have prefixes of different length. For example if we have 
\begin{align*}
&\{a_i\} = \{1,2,3,4,5,6,1,2,3,4,5,6, \ldots \} \\  
&\{b_i\}= \{7,8,9,1,2,3,4,5,6,1,2,3,4,5,6, \ldots\},
\end{align*}
 then $a_i$ and $b_i$ differ by prefixes of length $3+6k$ for $k$ a nonnegative integer. Because of this we make the following definition.
 \begin{definition} 
 We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a minimal prefix of length $n$ if they differ by a prefix of length $n$, but do not differ by a prefix of length $m$ for all $0<m<n$.
 \end{definition}
We note that if a pair of sequences differs by a prefix they differ by a minimal prefix. In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$.

\begin{theorem} The self-referencing sequence beginning with a $1$ generated over $\{1, c_1, c_2, \ldots, c_n\}$ differs from the self-referencing sequence beginning with $c_1$ generated over $\{1, c_1, c_2, \ldots, c_n\}$ by a minimal prefix of length 1.
\end{theorem}
\begin{proof}
Let $a_i$ refer to the $i$th term of the sequence beginning with a $1$ and $b_i$ refer to the $i$th term of the sequence beginning with $c_1$. We know that $a_1=1$, this means that the first block in $a_i$ must be of length 1. In other words we must have $a_2=c_1$, but since $a_2$ is now part of the second block of $a_i$ the value of $a_3$ is not specified by $a_1$. Thus we could specify the rest of $a_i$ only knowing the value of $a_2$, but $b_1=c_1$, so this is exactly how the values of $b_i$ is specified. Therefore
\begin{equation*}
a_{1+i} = b_{i},
\end{equation*}
so $a_i$ and $b_i$ differ by a prefix of length 1, but $a_{1} \neq b_{1}$ so it is a minimal prefix. 
\end{proof}

\begin{theorem} If $1<a<b$ then the self-referencing sequence beginning with $a$ generated over $\{1, a, b\}$ and the self-referencing sequence beginning with $c_2$ generated over $\{1, a, b\}$ are independent.
\end{theorem}
\begin{proof}
We proceed by contradiction. If the two sequences are not independent then they must differ by some minimal prefix of length $n$. If we let $a_i$ be the sequence beginning with $a$ and $b_i$ be the sequence beginning with $b$ the
\begin{equation*}
a_{n+k} = b_{k}.
\end{equation*}
We know that the first $b$ terms of the sequence which begins with a $b$ are $b$'s. Because $b>a>1$ no block can be longer then $b$. Thus the terms immediately preceding and following this block in the sequence $a_i$ must not be $b$'s. Because of this the first block of $b_i$ is the $m+1$th block of $a_i$, where $m$ is some positive integer. Since the value of the sequence at position $k$ is also the length of the $k$th block this means that $a_{m+1}=b_1$. In fact the $k$th block of $b_i$ is the $(m+k)$th block of $a_i$ and so 
\begin{equation*}
a_{m+k} = b_k
\end{equation*}

We now claim that the prefix has length strictly larger than $m$. Assume otherwise so $n \leq m$. 
This means the entry in the $a_i$ which gives the length of the $m+1$th block occurs at or after the beginning of the $m+1$th block. By our proof that self-referential sequences are well defined we know that the second case cannot occur. Therefore the only other option would be for the $m+1$th block to start at position $m+1$, but this would require every block to have length exactly 1, but this is a contradiction because $a_1=a>1$, that is the first block has length greater than one. Therefore the prefix must have length strictly larger than $m$.

We note that based on a previous step the two sequences differ by a prefix starting at $m$. Since $m<n$ we have produced a prefix which starts before $n$ which is a contradiction. Therefore the two sequences are independent. 
\end{proof}
\end{document}

### 2014-10-18 buildsequence.py

@@ -1,0 +1,22 @@
+def recurbuildseq(seq):
+    counter=0
+    oneortwo=0
+    seqout=[]
+    while counter<len(seq):
+        seqout = seqout + [oneortwo+1]*seq[counter]
+        counter = counter + 1
+        oneortwo = (oneortwo + 1) % 2
+    return seqout
+
+def buildseq(seq,n):
+    output=seq
+    iteration=0
+    while iteration < n:
+        output = recurbuildseq(output)
+        iteration = iteration +1
+    return output
+
+print buildseq([1,2],8)
+
+    
+    

### 2014-10-26 draft.tex
added def and notations section

@@
 \section{Definitions and Notations}
-In this section we introduce
+
+\emph{Written by Perry Kleinhenz, edited by Fermi Ma and Erik Waingarten}
+
+In this section, we introduce the notation we will use for the remainder of this paper. Consider the sequence
+\begin{align*}
+1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,1,2,1,1,...
+\end{align*}
+If we break the sequence up into contiguous blocks, where a block is a stretch of the sequence where only one number is used, we find that the block lengths of the sequence reproduce the original sequence.
+The first block length is 1, the second block length is 2, the third block length is 3, and so on.
+
+We say that this sequence is \emph{generated} by the set of numbers used in the sequence, so this sequence is generated by $\{1,2\}$.
+We call this set the \emph{generating set}.
+Note that here, it is unambiguous which number to use ``next".
+After a block of 1's, the only choice for the next block is the number 2.
+However, if a sequence had the generating set $\{1,2,3\}$, for example, a block of 1's could be followed by either a block of 2's or a block of 3's.
+Thus, we will assume that a \emph{rule} is given whenever a sequence is generated by more than 2 numbers.
+In particular, the rule will be to use the next number (in cyclic order) in the generating set.
+Note here that the set is not a set in the formal sense, as the order does not matter.
+However, we avoid naming it a ``sequence" to avoid possible confusion with its corresponding self-referencing sequence.
+
+With this rule, a the self-referencing sequence generated by $\{1,2,3\}$ has blocks of 1's followed by blocks of 2's followed by blocks of 3's, followed by blocks of 1's, and so on.
+Thus, the sequence is
 
@@
 
-\begin{theorem} If $1<a<b$ then the self-referencing sequence beginning with $a$ generated over $\{1, a, b\}$ and the self-referencing sequence beginning with $c_2$ generated over $\{1, a, b\}$ are independent.
+\begin{theorem} If $1<a<b$ then the self-referencing sequence beginning with $a$ generated over $\{1, a, b\}$ and the self-referencing sequence beginning with $b$ generated over $\{1, a, b\}$ are independent.
 \end{theorem}
@@
 \end{proof}
+
+\section{Equivalence}
+
+\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+
+The above analysis shows that, for example, the s
+
 \end{document}

### 2014-10-26 draft.tex
added a lot more to introduction

--- 2014-10-18 draft.tex
@@
 
-\section{Definitions and Notations}
-In this section we introduce
+\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+
+\subsection{Problem Setup}
+Consider the sequence
+\begin{align*}
+1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
+\end{align*}
+We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers:
+\begin{align*}
+[1],[2,2],[1,1],[2],[1],[2,2],[1],[2,2],[1,1],[2],[1,1],[2,2],\dots
+\end{align*}
+The block lengths, read from left to right, reproduce the original sequence. Thus, this sequence is \emph{self-referential}, and for the remainder of this paper we study properties of such sequences.
+
+We say that this sequence is \emph{generated} by the set of numbers used in the sequence, so the above sequence is generated by $\{1,2\}$.
+We will refer to this set as the \emph{generating set}.
+Observe that it is unambiguous which number to use next.
+After a block of 1's, the next block must be a block of 2's.
+However, in a sequence with generating set $\{1,2,3\}$, for example, a block of 1's could be followed by either a block of 2's or a block of 3's.
+Thus, we will assume that the next number to use is the following element (in cyclic order) in the generating set.
+(Note that because order matters, the \emph{generating set} is technically not a set, but we do not call it a ``sequence" to avoid possible confusion with its corresponding self-referencing sequence.)
+
+With this rule, a self-referencing sequence generated by $\{1,2,3\}$ has blocks of 1's followed by blocks of 2's followed by blocks of 3's, followed by blocks of 1's, and so on.
+Thus, $\{1,2,3\}$ can generate
+\begin{align*}
+1,2,2,3,3,1,1,1,2,2,2,3,2,1,2,2,3,3,1,1,\dots
+\end{align*}
+but not
+\begin{align*}
+1,3,3,3,2,2,2,1,1,1,3,3,2,2,1,1,3,2,1,3,\dots
+\end{align*}
+as the second sequence has the numbers in the wrong order.
+
+\subsection{Overview}
+
+In this paper, we address a variety of questions regarding these sequences.
+In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined.
+After that, we consider how the starting number for the sequence affects what the sequence can be.
+In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers.
+In the following Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.
+
+In Section~\ref{iterativeprocess}, we take a different approach to analyzing these sequences by showing that there is an alternate way of generating them with an iterative process.
+In Section~\ref{substitutionrules}, we show how a large number of these sequences can actually be generated with a simple set of substitution rules.
+We then introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating sets.
+Unfortunately, we find that the open problem of determining the density for the sequence of 1's and 2's is hard to solve using our methods.
+
+\section{Determinism}
+\label{determinism}
 
 \section{Prefixes}
+\label{prefixes}
 \emph{Written by Perry Kleinhenz, edited by Fermi Ma and Erik Waingarten}
@@
 \end{proof}
+
+\section{Equivalence}
+\label{equivalence}
+
+\section{An Iterative Process}
+\label{iterativeprocess}
+
+\section{Substitution Rules}
+\label{substitutionrules}
+
+\section{Density}
+\label{density}
+
+\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+
+The above analysis shows that, for example, the s
+
 \end{document}

### 2014-10-26 draft.tex
added more stuff

@@
 
+\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+
+We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number.
+To see why this is true, we can simply give an algorithm to build the sequence.
+Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$.
+
+\textbf{Algorithm:}
+Assume that the sequence $\{s\}$ is initially just $a_i$.
+At the next step, add an additional $a_i -1$ repetitions of $a_i$ to the end of the sequence.
+Now if there is a second term in the sequence, read that term and write down that many copies of $a_{i+1}$.
+If there is no term, write down $a_{i+1}$.
+
+NEED TO FINISH THIS.
+
+Since we know that a self-referencing sequence is uniquely determined by the generating set and the starting number, we introduce the notation $s(A,a)$ to denote the unique sequence generated by set $A$ with starting number $a \in A$.
+
 \section{Prefixes}
@@
 
-\begin{theorem} If $1<a<b$ then the self-referencing sequence beginning with $a$ generated over $\{1, a, b\}$ and the self-referencing sequence beginning with $b$ generated over $\{1, a, b\}$ are independent.
+\begin{theorem}
+\label{independentsequences}
+If $1<a<b$ then the self-referencing sequence beginning with $a$ generated over $\{1, a, b\}$ and the self-referencing sequence beginning with $b$ generated over $\{1, a, b\}$ are independent.
 \end{theorem}
@@
 \label{equivalence}
+\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+
+In this section, we consider a slightly weaker definition of equivalence.
+Consider, for example, sequences generated by $\{1,2,3\}$.
+Theorem~\ref{independentsequences} shows that the sequence starting with 2
+\begin{align*}
+2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+\end{align*}
+and the sequence starting with 3
+\begin{align*}
+3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+\end{align*}
+are independent in the sense that neither sequence is exactly contained in the other.
+However, if we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we get
+\begin{align*}
+3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+\end{align*}
+in both cases. (Need to further justify why these two are the same, probably with a reading distance argument)
+
+Thus, since both sequences are essentially the same if certain prefixes are removed, we call these sequences \emph{equivalent}.
+To formalize, this, we say that two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n,k \geq 0$ such that
+\begin{align*}
+a_{n+i} = b_{k+i}
+\end{align*}
+
+We now give the following conjecture
+
+\begin{conjecture} Any sequence generated by a set $S$ is equivalent to any other sequence generated by the same set $S$.
+\end{conjecture}
+
+This conjecture is partly motivated by the fact that the property holds for $\{1,2,3\}$.
+For generating sets with larger numbers, we have been unable to verify or refute the conjecture.
+However, in Section~\ref{Density}, we show that the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
 
@@
 \label{iterativeprocess}
+
+\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+
+We change focus in this section and consider how to generate self-referential sequences with an iterative process. We first look at the sequences generated by $\{1,2\}$.
+
+One possible iterative process is as follows.
+Let the sequence $\{s^0\}$ be 2, and let the sequence $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$.
+This gives:
+\begin{align*}
+\{s^{(0)}\} &= 2\\
+\{s^{(1)}\} &= 2,2\\
+\{s^{(2)}\} &= 2,2,1,1\\
+\{s^{(3)}\} &= 2,2,1,1,2,1\\
+\{s^{(4)}\} &= 2,2,1,1,2,1,2,2,1\\
+&\cdots
+\end{align*}
+Let $s^\infty$ be the limit of this process. We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2.
+\begin{proof}
+We show that this is true by simply showing that if $s^{(i)}$ is a prefix of
+\end{proof}
 

### 2014-10-26 draft.tex
edited iterative process

@@
 Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$.
+If we ever write down some $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
+\begin{quote}
+\textbf{Algorithm:}\\
+Suppose the first number of the sequence is $a_i$.
+Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block.
+If $a_i$ is not 1, then there is a second element of that sequence.
+We read that second element, and then we add a block to the end of that sequence with length dictated by the value of the second element.
+If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
 
-\textbf{Algorithm:}
-Assume that the sequence $\{s\}$ is initially just $a_i$.
-At the next step, add an additional $a_i -1$ repetitions of $a_i$ to the end of the sequence.
-Now if there is a second term in the sequence, read that term and write down that many copies of $a_{i+1}$.
-If there is no term, write down $a_{i+1}$.
+This procedure can continue deterministically so long as there is always an element to read.
+However, every time we write down a block, we read 1 element and write down at least 1 element, so the end of the sequence will always be after the spot we are reading.
+The only exception is when $a_i = 1$, but then since $a_{i+1} \neq 1$, this issue will never occur again.
+\end{quote}
 
-NEED TO FINISH THIS.
-
-Since we know that a self-referencing sequence is uniquely determined by the generating set and the starting number, we introduce the notation $s(A,a)$ to denote the unique sequence generated by set $A$ with starting number $a \in A$.
+Thus, a self-referencing sequence is uniquely determined by the generating set and the starting number.
+From this point on, we will denote by $s(A,a)$ the unique sequence generated by set $A$ with starting number $a \in A$.
 
@@
 \begin{proof}
-We show that this is true by simply showing that if $s^{(i)}$ is a prefix of
+We prove this with induction.
+We show that if $s^{(i)}$ is a prefix of $s(\{1,2\},2)$, then $s^{(i+1)}$ is a longer prefix of $s(\{1,2\},2)$.
+The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(\{1,2\},2)$.
+
+For the inductive step, observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence.
+We know that it describes a longer prefix because the prefix it describes has length equal to the sum of the original prefix.
+In other words, the length of $s^{(i+1)}$ must be equal to the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1).
+This holds true because $s^{(i)}$ merely contains a collection of lengths, which are summed up to give the next prefix.
 \end{proof}
+
+It turns out that this same argument holds in general.
+
+\begin{theorem}
+Let $\{s^0\} = a$ for some $a \in A$ where $a \neq 1$, and let $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$.
+Then the limit of this process, $s^\infty$, is the sequence $S(A,a)$.
+\end{theorem}
+
+Note that we must specify that $a \neq 1$, because that is the only case in which the length of $s^{(i+1)}$ is not strictly greater than the length of $s^{(i)}$.
+The proof of this theorem is omitted, as it is essentially the same proof as for the $A = \{1,2\}, a = 2$ case.
+
+It turns out that we can be less restrictive about what the starting sequence is.
+Going back to the case of $\{1,2\}$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2.
+For example, suppose $s^{(0)} = 2,1,1,1,2$.
+The rules for generating large $s^{(i)}$ give:
+\begin{align*}
+\{s^{(0)}\} &= 2,1,1,1,2\\
+\{s^{(1)}\} &= 2,2,1,2,1,2,2\\
+\{s^{(2)}\} &= 2,2,1,1,2,1,1,2,1,1,2,2\\
+\{s^{(3)}\} &= 2,2,1,1,2,1,2,2,1,2,1,1,2,1,2,2,1,1\\
+&\cdots
+\end{align*}
 

### 2014-10-26 draft.tex
added more stuff about weak and strong convergence

@@
 Let $\{s^0\} = a$ for some $a \in A$ where $a \neq 1$, and let $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$.
-Then the limit of this process, $s^\infty$, is the sequence $S(A,a)$.
+Then the limit of this process, $s^\infty$, is the sequence $s(A,a)$.
 \end{theorem}
@@
 \begin{align*}
-\{s^{(0)}\} &= 2,1,1,1,2\\
-\{s^{(1)}\} &= 2,2,1,2,1,2,2\\
-\{s^{(2)}\} &= 2,2,1,1,2,1,1,2,1,1,2,2\\
-\{s^{(3)}\} &= 2,2,1,1,2,1,2,2,1,2,1,1,2,1,2,2,1,1\\
+\{s^{(0)}\} &= \mathbf{2},1,1,1,2\\
+\{s^{(1)}\} &= \mathbf{2,2,1},2,1,2,2\\
+\{s^{(2)}\} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
+\{s^{(3)}\} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
 &\cdots
 \end{align*}
+Here, we highlight in bold the numbers that form a prefix of $s(\{1,2\},2)$.
+We note that $s^\infty = s(A,a)$ simply because $s^{(0)}$ contains a 2 at the beginning, whch then causes a 2,2 to appear at the start of $s^{(1)}$, which causes $2,2,1,1$ to appear at the start of $s^{(2)}$, and so on.
+The rest of each $s^{(i)}$ that is not a prefix of $s({1,2},2)$ we call the \emph{tail}.
+
+We now consider two different forms of convergence.
+We know that $s^{(i)}$ in some sense approaches $s(A,a)$, as arbitrarily many elements at the beginning of the sequence will match up with the beginning of $s(A,a)$ as $i$ increases.
+However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(A,a)$.
+Thus, we can formalize two types of convergence: \emph{weak convergence} and \emph{strong convergence}.
+Weak convergence is satisfied if $s^\infty = s(A,a)$, and strong convergence is satisfied if $s^{(i)}$ is a prefix of $s(A,a)$ for all $i \geq k$ for some $k$.
 

### 2014-10-26 draft.tex
started substitution rules

@@
 
+\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+
+The iterative process given in Section~\ref{iterativeprocess} is one way to generate self-referencing sequences.
+However, for certain generating sets $A$, there is a simpler ways to perform this iteration using a fixed set of substitution rules.
+Take $A = \{1,3\}$ as an example, and consider the sequence that starts with a 3:
+\begin{align*}
+333111333131333111333\dots
+\end{align*}
+
 \section{Density}
@@
 
-\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
-
-The above analysis shows that, for example, the s
-
 \end{document}

### 2014-10-26 draft.tex
added more to substitution rules section

@@
 \end{align*}
+We can use the iterative process from Section~\ref{iterativeprocess}, and suppose that $s^{(0)} = 33$. We get:
+\begin{align*}
+\{s^{(0)}\} &= 3,3\\
+\{s^{(1)}\} &= 3,3,3,1,1,1\\
+\{s^{(2)}\} &= 3,3,3,1,1,1,3,3,3,1,3,1\\
+\{s^{(3)}\} &= 3,3,3,1,1,1,3,3,3,1,3,1,3,3,3,1,1,1,3,3,3,1,3,3,3,1\\
+&\cdots
+\end{align*}
+It turns out that we can model this evolution with the following substitution rules:
+\begin{align*}
+\text{Rule }A: 3,3 &\to 3,3,3,1,1,1 \\
+\text{Rule }B: 3,1 &\to 3,3,3,1\\
+\text{Rule }C: 1,1 &\to 3,1
+\end{align*}
+This is to be interpreted as follows.
+Starting from 3,3, the only possible rule to apply is Rule A, which gives $3,3,3,1,1,1$ at the following iteration.
+We then apply Rule A to $3,3$, Rule B to $3,1$ and Rule C to $1,1$.
+Note that at every step, we break up the sequence into chunks of 2 and then simultaneously apply \emph{all} the rules.
+This will always be possible, since the rules preserve the fact that these sequences have even length.
+So at the following step, we have 3,3,3,1,1,1,3,3,3,1,3,1, which is $s^{(2)}$.
+Notice that in general, after the $i$th application of the substitution rules, the sequence is equal to $s^{(i)}$.
 

### 2014-10-26 draft.tex
Edited equivalence
Adding stuff on Reading distance

@@
 \label{equivalence}
-\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+\emph{Written by Fermi Ma and Perry Kleinhenz, edited by Erik Waingarten}
 
@@
 
+We now introduce a notion which will help us categorize sequences as equivalent. 
+\begin{definition} Consider a sequence $x$.
+We say define the read distance for an element $x_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
+We denote this as $Rd(x_i)$.
+\end{definition}
+The read distance can be any nonnegative integer. For example if we consider the sequence $a_i$ generated by $\{1,2,3\}$ beginning with a 2  
+\begin{align*}
+2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+\end{align*}
+the read distance for the $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number.
+We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0,1,2,4,6,6,6,6,7,8,9,\ldots$.
+From this we might suspect that the read distance for any self referencing sequence will also be monotonically increasing.
+We can in fact show a much stronger result.
+\begin{lemma} 
+\label{readdistformula}
+Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$ for $i>1$ is 
+\begin{equation*}
+\sum_{k=1}^{i-1}( a_k - 1)
+\end{equation*} 
+\end{lemma}
+\begin{proof}
+We will prove this using induction.
+We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number.
+For the base case $i=2$ we must show that the read distance is $a_1-1$.
+Well we know that the first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at \emph{position} $a_1$.
+The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.
+
+Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the $l$th block and the number at position $l$ is 
+\begin{equation*}
+\sum_{k=1}^{l-1}( a_k - 1).
+\end{equation*} 
+This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block.
+Also note that $a_{l+1}$ is one position to the right of $a_l$.
+Therefore the beginning of the $l+1$th block is
+\begin{equation*}
+a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
+\end{equation*} 
+positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
+\end{proof}
+Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and the read distance is monotonically increasing.
+Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a bound on the number of consecutive times the read distance can stay constant.
+
+\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max{ a_1, a_2, \ldots, a_n}$.
+Suppose the read distance for $x_i$ is $d$, then the read distance for $x_{i+a}$ is at least $d+1$.
+\end{lemma}
+\begin{proof}
+Assume otherwise.
+By our Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$.
+By our assumption it must be exactly $d$.
+Lemma ~\ref{readdistformula} also tells us that $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$.
+But this means that we have a block of length $a+1$, which is a contradiction of the self referencing property of the sequence.
+Therefore the read distance for $x_{i}$ is at least $d+1$.
+\end{proof}
+
+\begin{theorem} If two sequences are equivalent then the first element of the shared sequence must have the same read distance in both sequences.
+\end{theorem}
+\begin{proof}
+Assume otherwise. So if the two sequences are $a_i$ and $b_i$ and we have 
+\begin{equation*}
+a_{i+l} = b_{i+k} = c_{i}
+\end{equation*}
+
+\end{proof}
 \section{An Iterative Process}

### 2014-10-27 draft.tex
More updates to Equivalence

@@
 \begin{proof}
-Assume otherwise. So if the two sequences are $a_i$ and $b_i$ and we have 
+Assume otherwise. So if the two sequences are $a_j$ and $b_j$ and we have 
 \begin{equation*}
@@
 \end{equation*}
+such that if $m$ or $l$ were made smaller this relation would not hold.
+Suppose the read distance for $b_{1+k}$ is $x$ and the read distance for $a_{1+l}$ is $y$.
+Without loss of generality let us assume that $x>y$.
+Therefore we know there are two separate blocks of length $c_1$ in the sequence, one which corresponds to $a_{i+l}$ and another which corresponds to $b_{i+k}$.
+Because the two sequences are identical there exists some index $n$ such that $c_{n+1}=a_{m}$ is the element of sequence $a$ which corresponds to the second block of length $c_1$. 
+Because the sequences are self referential we know that $c_1, c_2, \ldots, c_{n}$  must appears in $b$ prior to element $i+k$.
+In fact the self referential nature of the sequence guarantees that $b_{i+k-1}=c_{n}$, that is this subsequence must end just before the two sequences become the same.
+Thus our sequence $b$ is
+\begin{equation*}
+\cdots c_1 c_2 c_3 \cdots c_{n} c_1 c_2 \cdots c_n \cdots
+\end{equation*}
+where the prefix begins at the second $c_1$. We now claim that the common sequence cycles through
+\begin{equation*}
+c_1 c_2 \cdots c_n.
+\end{equation*}
+Well by Lemma ~\ref{readdistformula} since the read distance for $b_{1+k}$ is greater than the read distance for $a_{1+l}$, and the sequences are identical for subsequent indices, we know that the read distance for $b_{i+k}$ is longer than the read distance for $a_{i+l}$ by exactly $x-y$.
+Thus $b_{k+i}$ corresponds to the same block as $a_{l+(x-y)+i}$.
+But $b_{k+i} = c_{i}$ and $a_{l+(x-y)+i}=c_{x-y+i}$ and so $c_{i} = c_{x-y+i}$ and so the common sequence cycles.
+Because the common sequence cycles and $b$ has the same cycle, but one which begins before the common sequence $l$ could be made smaller and have the equivalence relation hold.
+This is a contradiction and so the first element of the shared sequence must have the same read distance in both sequences.
+\end{proof}
 
+We also have a sufficient condition for equivalence 
+\begin{lemma} If two sequences $a_i, b_i$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are the same and 
+\begin{equation*}
+a_{m+k} = b_{l+k}
+\end{equation*}
+where $k$ ranges from 0 to the read distance for $a_m$ then $a_i$ and $b_i$ are equivalent. 
+\end{lemma}
+\begin{proof}
+We will prove this using a simple induction.
+Suppose the read distance for $a_m$ is $d$ then know that $a_{m+d} = b_{l+d}$ because $a_{m+d-1} = b_{l+d-1}$ and $a_{m+d}$ is the beginning of a new block.
+Furthermore we have that the read distance for $a_{m+d}$ equals the read distance for $b_{l+d}$.
+Now suppose that we have for $k \leq j$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$ and we would like to show that this holds for $k = j+1$.
+By Lemma ~\ref{readdistformula} we actually already know that the read distance for the two is the same, as it is the same for the preceding terms, which are equal.
+
+If $a_{m+k}$ is the last term in a block then $b_{m+k}$ must also be the last term in its block, because the length of blocks is determined by earlier terms which we know to be equal and to have the same read distance.
+Since $a_{m+k} = b_{l+k}$  in this case we have $a_{m+k+1} = b_{l+k+1}$.
+
+Otherwise $a_{m+k}$ and $a_{m+k+1}$ are in the same block, then we only need to show that $b_{l+k}$ and $b_{l+k+1}$ are in the same block as well.
+But as we argued above the length of blocks is determined by terms earlier in the sequence which we know to be equal and have the same read distance.
+Therefore we have $a_{m+k+1} = b_{l+k+1}$ for this case as well.
+This completes our induction and so $a_i$ and $b_i$ are equivalent.
 \end{proof}
+These conditions for equivalence lead us to make the following conjecture on a sufficient condition for nonequivalence.
+
+\begin{conjecture} If two sequences $a_i$ and $b_i$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
+\end{conjecture}
+
 \section{An Iterative Process}

### 2014-10-27 draft.tex
edited up to theorem 1, which I changed to lemma 1.

@@
 \end{align*}
-The block lengths, read from left to right, reproduce the original sequence. Thus, this sequence is \emph{self-referential}, and for the remainder of this paper we study properties of such sequences.
+The block lengths, read from left to right, reproduce the original sequence. Thus, this sequence is \emph{self-referential}. For the remainder of this paper we study properties of such sequences.
 
@@
 After a block of 1's, the next block must be a block of 2's.
-However, in a sequence with generating set $\{1,2,3\}$, for example, a block of 1's could be followed by either a block of 2's or a block of 3's.
-Thus, we will assume that the next number to use is the following element (in cyclic order) in the generating set.
-(Note that because order matters, the \emph{generating set} is technically not a set, but we do not call it a ``sequence" to avoid possible confusion with its corresponding self-referencing sequence.)
+There does exist ambiguity in a sequence with generating set $\{1,2,3\}$, since after a block of 1's, ther can be a block of 2's or a block of 3's.
+We assign an order to the generating set and resolve the ambiguity.
+The order we will use will be the order of the elements presented, for example, $\{1, 2, 3 \}$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.
+Note that because order matters, the \emph{generating set} is technically not a set, but we do not call it a ``sequence" to avoid possible confusion with its corresponding self-referencing sequence.
 
@@
  \end{definition}
-We note that if a pair of sequences differs by a prefix they differ by a minimal prefix. In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$.
+We note that if a pair of sequences differs by a prefix they differ by a minimal prefix.
+In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$.
+The following lemma is a trivial consequence of our algorithm for generating self-referential sequences:
 
-\begin{theorem} The self-referencing sequence beginning with a $1$ generated over $\{1, c_1, c_2, \ldots, c_n\}$ differs from the self-referencing sequence beginning with $c_1$ generated over $\{1, c_1, c_2, \ldots, c_n\}$ by a minimal prefix of length 1.
-\end{theorem}
+\begin{lemma} Let $C = \{1, c_1, c_2, \ldots, c_n\}$, $s(C, 1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
+\end{lemma}
 \begin{proof}
-Let $a_i$ refer to the $i$th term of the sequence beginning with a $1$ and $b_i$ refer to the $i$th term of the sequence beginning with $c_1$.
-We know that $a_1=1$, this means that the first block in $a_i$ must be of length 1.
-In other words we must have $a_2=c_1$, but since $a_2$ is now part of the second block of $a_i$ the value of $a_3$ is not specified by $a_1$.
-Thus we could specify the rest of $a_i$ only knowing the value of $a_2$, but $b_1=c_1$, so this is exactly how the values of $b_i$ is specified.
-Therefore
-\begin{equation*}
-a_{1+i} = b_{i},
-\end{equation*}
-so $a_i$ and $b_i$ differ by a prefix of length 1, but $a_{1} \neq b_{1}$ so it is a minimal prefix. 
+Let $a_i$ refer to the $i$th term of $s(C, 1)$ and $b_i$ refer to the $i$th term of $s(C, c_1)$.
+If we use our algorithm for generating the sequence, we note that $a_1 = 1$, and so $a_2 = c_1$, and that this is the beginning of a new block of length $c_1$.
+Furthermore, $b_1 = c_1$ and it indicated the beginning of a block of length $c_1$.
+Note that our algorithm is ``memory-less'' in the sense that it does not maintain any state about previous elements.
+Because the number and the begining of the block is the same.
+The algorithm will generate the same sequence, shifted by 1.
+Therefore, $s(C,1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
 \end{proof}

### 2014-10-27 draft.tex
gave explanation for why sequence of 1,2,3 starting at 2 and starting at 3 are the same. edited up to conjecture 1

@@
 In this section, we consider a slightly weaker definition of equivalence.
-Consider, for example, sequences generated by $\{1,2,3\}$.
+Consider, for example, sequences generated by $C = \{1,2,3\}$.
 Theorem~\ref{independentsequences} shows that the sequence starting with 2
@@
 \end{align*}
-in both cases. (Need to further justify why these two are the same, probably with a reading distance argument)
+in both cases.
+Lets look at the state of the algorithm that generates both sequences.
+After the first block, the execution of the algorithm of $s(C, 2)$ right after placing the $3$'s (in the 4th and 5th position), the algorithm is in state corresponding to:
+\begin{itemize}
+\item block size $3$
+\item no elements on the current block
+\item starting number is $1$
+\end{itemize}.   
+If we look at the execution of the algorithm that generates $s(C, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
+\begin{itemize}
+\item block size $3$
+\item no elements on the current block
+\item starting number is $1$
+\end{itemize}
+Therefore, since the states are the same, the same sequence will be generated after the 4th position of $s(C, 2)$ and after the $3rd$ position of $s(C, 3)$ will be the same.
+By inspection, we can see we can include $2$ of the 3's.
 
@@
 \begin{definition} Consider a sequence $x$.
-We say define the read distance for an element $x_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
+We define the read distance for an element $x_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
 We denote this as $Rd(x_i)$.

### 2014-10-27 draft.tex
edited iterative process section
addded some rules stuff

@@
 \end{align*}
-Let $s^\infty$ be the limit of this process. We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2.
+Let $s^\infty$ be the limit of this process.
+By limit, we mean the limit with respect to finitely many terms $s^{(i)}_j \rightarrow s^\infty_j$ as$i \rightarrow \infty$.
+We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2.
 \begin{proof}
@@
 
+Why does this work?
+Well, first of all, notice that the rules are the natural applications of the self-referential sequence.
+Also, note that we have decided to make the rules of length 2; if we had made the self-referential rules $1 \rightarrow 1$ and $3 \rightarrow 333$, this would not have worked.
+In addition, we know that each block length preserves parity.
+
+So we can ask, when can we do this? We would like to be able to simulate our original algorithm with this current one. We had established that our algorithm had a state corresponding to 
+\begin{itemize}
+\item current block length number
+\item how many elements already written down in the block
+\item which number we need to write.
+\end{itemize}
+
+In an algorithm with rules, we cannot remember anything about the previous rules.
+Therefore, each rule needs to encompas the block length numbers, the number of elements already written, and which number we need to write.
+The first condition is trivial to keep, since the rules themselves will say the block length numbers, in order to keep track of the numbers already written, each rule should be self-contained.
+That means that the rules should read off the block numbers.
+We would also need the length of the rules to be divisible by the size of the generating set.
+This is so that at the start of each rule, we always know which numbers to write.
+
 \section{Density}

### 2014-10-27 draft.tex
asdasdfasdfasdf
more rule stuff

@@
 \label{independentsequences}
-If $1<a<b$ then the self-referencing sequence beginning with $a$ generated over $\{1, a, b\}$ and the self-referencing sequence beginning with $b$ generated over $\{1, a, b\}$ are independent.
+Let $C = \{ 1, a, b\}$ where $1<a<b$. $s(C, a)$ and $s(C, b)$ are independent.
 \end{theorem}
@@
 
+Therefore, we can formulate a set of sufficient conditions to make rules to generate the self-referential sequence with generating set $G$.
+We want to have a set of rules $R$ where each rule $r \in R$ has length $|r| = l$, such that all $|G|^{l}$ strings are in the rule set and for each $r \in R$
+\begin{itemize}
+\item each rule reads off the block lengths in the same order of $G$
+\item $\sum_{i \in r} i$ is divisible by $l$
+\item $l$ is divisible by $|G|$
+\end{itemize}
+Given these conditions on every rule, we can guarantee that the self-referential sequence is generated.
+Note that in this case, the above example does not satisfy our rules until we say that $1, 3 \to 3, 1, 1, 1$.
+
+Another exaple, we can have $G = \{ 2, 4 \}$, and we can have that the rules be
+\begin{align*}
+\text{Rule }A: 2,2 &\to 2, 2, 4, 4 \\
+\text{Rule }B: 2,4 &\to 2, 2, 4,4,4,4\\
+\text{Rule }C: 4,2 &\to 2,2,2,2,4,4\\
+\text{Rule }D: 4,4 &\to 2, 2, 2, 2, 4, 4, 4, 4
+\end{align*}
+
 \section{Density}

### 2014-10-27 draft.tex
adding density

@@
 
+One of the reasons why generating a rule set for $\{1, 2\}$ is complicated is that since $1, ..., 1$ must be in the rule set, the length of the rules must be even (since $|G| = |\{1, 2\}| = 2$).
+On the other hand, $2, 1, ..., 1$ must also be in the set and so the rule length must be odd.
+
 \section{Density}
 \label{density}
+\emph{Written by Erik Waingarten, edited by Perry Kleinhenz and Fermi Ma}
 
+In this section, we analyze two things: the first is how to take the rule based methods of sequence generation to show that if a density exists, we can find what it is.
+We will explain the method and give an example applied to the case above of $G = \{ 1, 3\}$ and $G = \{ 2, 4 \}$.
+In the second part, we will show an ad hoc way to bound the density of $G = \{ 1, 2\}$ in a way that suggests that the density converges to 2; however, this method is computationally intractable for smaller bounds since the number of cases to check increases exponentially.
+
+\subsection{Determining the density}
+
+We have found a technique to compute the density of numbers in a self-referential sequence (if one exists) for some generating sets.
+In particular, suppose we have a generating set $G$ with a set of rules $R$ that satisfy the conditions from the section above.
+For each rule $r \in R$, we will say denote the rule as $r_l \rightarrow r_r$ as the inputs and the outputs.
+
+Now we can count the fraction of times $r_r$ appears in the sequence at each step.
+Since we know the rules, these will become a set of $|R|$ equalities that reference each other.
+We can take their ratios and solve for the fraction of times that each $r_r$ appears and compute the density by assuming a stable state.
+
+For example, lets look at the rules we have determined for $G = \{ 1, 3 \}$. 
+\begin{align*}
+\text{Rule }A: 3,3 &\to 3,3,3,1,1,1 \\
+\text{Rule }B: 3,1 &\to 3,3,3,1\\
+\text{Rule }C: 1,1 &\to 3,1
+\end{align*}
+Then we can let $3,3,3,1,1,1 = A$, $3,3,3,1 = B$ and $3, 1 = C$. If $A_i$ is the number of occurences of $A$ in $s^{(i)}$, then 
+\[ A_i = B_{i-1} + A_{i-1} \]
+Likewise, we have
+\[ B_i = A_{i-1} + B_{i-1} + C_{i-1} \]
+\[ C_i = A_{i-1} \]
+So now we can take the fraction by dividing by the total number of them:
+\[ A_i' = \dfrac{B_{i-1}' + A_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
+\[ B_i' = \dfrac{A_{i-1}' + B_{i-1}' + C_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
+\[ C_i' = \dfrac{A_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
+So if we assume a density, there will be a stable state which will be the density, so in the case where $A_i'=A_{i-1}', B_i' = B_{i-1}', C_i'=C_{i-1}'$, we can solve for the fractional appearence of each and get the density of the numbers.
+
+In particular, the density of 3's is approxinately $0.602$ by computing the solution to the set of quadratic equations.
+
+\subsection{Density bounds when $G = \{1, 2\}$}
+
+As shown in the previous section, we cannot establish the set of rules we want when $G = \{1 , 2\}$; however, there is some structure that we can exploit.
+In particular, we know that the sequence $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence.
+This is because they would be part of a block of size at least $3$.
+
+What does this mean? Well, we know that if there is a density of $1$'s and $2$'s, then that density $d_2$ is bounded between 
+\[ \dfrac{1}{3} \leq d_2 \leq \dfrac{2}{3} \]
+
+In general, this gives a procedure for computing the limiting density.
+For example, now we know that there must be at least one $2$ for every three numbers.
+So now we can look at how many $2$'s are in any given string of length $9$.
+We know there must be at least $3$, but each consecutive $3$ elements must contain a $3$.
+There are only two cases where there are three 2's in a segment of $9$ numbers in the sequence, namely
+\[ 2, 1, 1, 2, 1, 1 ,2 , 1, 1 \]
+\[ 1, 2, 1, 1, 2, 1, 1, 2, 1 \]
+\[ 1, 1, 2, 1, 1, 2, 1, 1, 2 \]
+In each of these cases, we can read the subsequence that must have generated them part of them. In order, they are
+\[ 2, 1, 2, 1, 2 \]
+\[ 1, 2, 1, 2, 1\]
+\[ 2, 1, 2, 1, 2 \]
+And these are generated by $1, 1, 1$.
+Which means that none of these are possible, so in addition to having one 2 for every group of 3, there is another 2 for every group of 9.
+Therefore, we have shown that the density must be bounded by
+\[ \dfrac{1}{3} + \dfrac{1}{9} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} \]
+In general, we can continue this process, and we did, to show that
+\[ \dfrac{1}{3} + \dfrac{1}{9} + \dfrac{1}{27} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} - \dfrac{1}{27} \]
+Of course, the number of cases to check increases very quickly: there are $3$ choices to put the $2$ in the groups of $3$, $3$ choices to put the 3 groups of 2's and 2 positions within these groups for the groups of 9, and so on.
+Its not hard to see that this number grows exponentially fast, making it computationally difficult to analyze more cases.
+
+We believe that if there is a way to show this, we will be able to show that the density is $\frac{1}{2}$, since the density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number.
+However, generalizing the procedure to improve the bound seems dificult.
 \end{document}

### 2014-10-28 draft.tex
Edited Intro and Determinism

@@
 In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined.
-After that, we consider how the starting number for the sequence affects what the sequence can be.
+After that, we consider how the starting number for the sequence affects the long term behavior of the sequence.
 In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers.
@@
 We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number.
-To see why this is true, we can simply give an algorithm to build the sequence.
+To see why this is true, we give an algorithm to build the sequence.
 Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$.
@@
 Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block.
-If $a_i$ is not 1, then there is a second element of that sequence.
-We read that second element, and then we add a block to the end of that sequence with length dictated by the value of the second element.
+If $a_i$ is not 1, then there is a second element of that block.
+We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence.
 If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
 
-This procedure can continue deterministically so long as there is always an element to read.
-However, every time we write down a block, we read 1 element and write down at least 1 element, so the end of the sequence will always be after the spot we are reading.
-The only exception is when $a_i = 1$, but then since $a_{i+1} \neq 1$, this issue will never occur again.
+This procedure continues deterministically so long as there is an element to read.
+Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
 \end{quote}
@@
 Thus, a self-referencing sequence is uniquely determined by the generating set and the starting number.
-From this point on, we will denote by $s(A,a)$ the unique sequence generated by set $A$ with starting number $a \in A$.
+From this point on, we will denote by $s(G,g)$ the unique sequence generated by set $G$ with starting number $g \in G$.
 

### 2014-10-28 draft.tex
Edited Prefixes

@@
 \end{equation*}
-for $i$ any positive integer. if no such $n$ exists then we say that the two sequences are independent.
+for $i$ any positive integer. We say that two sequences are independent if they do not differ by a prefix for any $n$, in either order. 
 \end{definition}
-Note that a pair of sequences can have prefixes of different length. For example if we have 
+Note that a pair of sequences can differ by multiple prefixes of different length. For example if we have 
 \begin{align*}
@@
 In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$.
-The following lemma is a trivial consequence of our algorithm for generating self-referential sequences:
+The following lemma is a direct consequence of our algorithm for generating self-referential sequences:
 
-\begin{lemma} Let $C = \{1, c_1, c_2, \ldots, c_n\}$, $s(C, 1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
+\begin{lemma} Let $G = \{1, g_1, g_2, \ldots, g_n\}$, $s(G, 1)$ and $s(G, g_1)$ differ by a minimal prefix of length at most 1.
 \end{lemma}
 \begin{proof}
-Let $a_i$ refer to the $i$th term of $s(C, 1)$ and $b_i$ refer to the $i$th term of $s(C, c_1)$.
-If we use our algorithm for generating the sequence, we note that $a_1 = 1$, and so $a_2 = c_1$, and that this is the beginning of a new block of length $c_1$.
-Furthermore, $b_1 = c_1$ and it indicated the beginning of a block of length $c_1$.
+Let $a_i$ refer to the $i$th term of $s(G, 1)$ and $b_i$ refer to the $i$th term of $s(G, g_1)$.
+If we use our algorithm for generating the sequence, we note that $a_1 = 1$, and so $a_2 = g_1$, and that this is the beginning of a new block of length $g_1$.
+Furthermore, $b_1 = g_1$ and it indicated the beginning of a block of length $g_1$.
 Note that our algorithm is ``memory-less'' in the sense that it does not maintain any state about previous elements.
@@
 The algorithm will generate the same sequence, shifted by 1.
-Therefore, $s(C,1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
+Therefore, $s(G,1)$ and $s(G, g_1)$ differ by a minimal prefix of length at most 1.
 \end{proof}
-
+We now give a 
 \begin{theorem}
 \label{independentsequences}
-Let $C = \{ 1, a, b\}$ where $1<a<b$. $s(C, a)$ and $s(C, b)$ are independent.
+Let $G = \{ 1, a, b\}$ where $1<a<b$. $s(G, a)$ and $s(G, b)$ are independent.
 \end{theorem}
@@
 If the two sequences are not independent then they must differ by some minimal prefix of length $n$.
-If we let $a_i$ be the sequence beginning with $a$ and $b_i$ be the sequence beginning with $b$ the
+If we let $a_i$ be the sequence beginning with $a$ and $b_i$ be the sequence beginning with $b$ then
 \begin{equation*}
@@
 \end{equation*}
-We know that the first $b$ terms of the sequence which begins with a $b$ are $b$'s.
+We know that the first $b$ terms of $b_i$ are $b$'s.
 Because $b>a>1$ no block can be longer then $b$.
-Thus the terms immediately preceding and following this block in the sequence $a_i$ must not be $b$'s.
-Because of this the first block of $b_i$ is the $m+1$th block of $a_i$, where $m$ is some positive integer.
-Since the value of the sequence at position $k$ is also the length of the $k$th block this means that $a_{m+1}=b_1$.
+Thus the terms immediately preceding and following this block in  $a_i$ cannot be $b$'s.
+Because of this the first block of $b_i$ is the $m+1$th block of $a_i$, for some $m$ a positive integer.
+Since the value of the sequence at position $k$ is the length of the $k$th block this means that $a_{m+1}=b_1$.
 In fact the $k$th block of $b_i$ is the $(m+k)$th block of $a_i$ and so
@@
 We now claim that the prefix has length strictly larger than $m$. Assume otherwise so $n \leq m$. 
-This means the entry in the $a_i$ which gives the length of the $m+1$th block occurs at or after the beginning of the $m+1$th block.
+This means that $a_{m+1}$ which gives the length of the $m+1$th block occurs at or after the beginning of the $m+1$th block, which begins at position $a_{n+1}$.
 By our proof that self-referential sequences are well defined we know that the second case cannot occur.
-Therefore the only other option would be for the $m+1$th block to start at position $m+1$, but this would require every block to have length exactly 1, but this is a contradiction because $a_1=a>1$, that is the first block has length greater than one.
+Therefore we must have the $m+1$th block start at position $m+1$.
+This would require every block to have length exactly 1, but this is a contradiction because the first block has length $a_1=a>1$.
 Therefore the prefix must have length strictly larger than $m$.

### 2014-10-28 draft.tex
Edited equivalence

@@
 In this section, we consider a slightly weaker definition of equivalence.
-Consider, for example, sequences generated by $C = \{1,2,3\}$.
+Consider, for example, sequences generated by $G = \{1,2,3\}$.
 Theorem~\ref{independentsequences} shows that the sequence starting with 2
@@
 Lets look at the state of the algorithm that generates both sequences.
-After the first block, the execution of the algorithm of $s(C, 2)$ right after placing the $3$'s (in the 4th and 5th position), the algorithm is in state corresponding to:
+After the first block, the execution of the algorithm of $s(G, 2)$ right after placing the first $3$'s (in the 3rd and 4th position), the algorithm is in state corresponding to:
 \begin{itemize}
 \item block size $3$
-\item no elements on the current block
+\item no elements in the current block
 \item starting number is $1$
 \end{itemize}.   
-If we look at the execution of the algorithm that generates $s(C, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
+If we look at the execution of the algorithm that generates $s(G, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
 \begin{itemize}
 \item block size $3$
-\item no elements on the current block
+\item no elements in the current block
 \item starting number is $1$
 \end{itemize}
-Therefore, since the states are the same, the same sequence will be generated after the 4th position of $s(C, 2)$ and after the $3rd$ position of $s(C, 3)$ will be the same.
-By inspection, we can see we can include $2$ of the 3's.
+Therefore, since the states are the same, the sequence generated after the 4th position of $s(G, 2)$ and after the $3rd$ position of $s(G, 3)$ will be the same.
+By inspection, we see we can also include two of the 3's.
 
-Thus, since both sequences are essentially the same if certain prefixes are removed, we call these sequences \emph{equivalent}.
+Thus, since both sequences are essentially the same if a prefix is removed from each of them, we call these sequences \emph{equivalent}.
 To formalize, this, we say that two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n,k \geq 0$ such that
@@
 For generating sets with larger numbers, we have been unable to verify or refute the conjecture.
-However, in Section~\ref{Density}, we show that the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
+However, in Section~\ref{density}, we show that for some generating sets the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
 
 We now introduce a notion which will help us categorize sequences as equivalent. 
-\begin{definition} Consider a sequence $x$.
-We define the read distance for an element $x_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
-We denote this as $Rd(x_i)$.
+\begin{definition} Consider a sequence $a$. We define the read distance for an element $a_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
 \end{definition}
@@
 \end{align*}
-the read distance for the $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number.
+the read distance for $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number.
 We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0,1,2,4,6,6,6,6,7,8,9,\ldots$.
 From this we might suspect that the read distance for any self referencing sequence will also be monotonically increasing.
-We can in fact show a much stronger result.
+We can in fact compute the read distance exactly.
 \begin{lemma} 
 \label{readdistformula}
-Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$ for $i>1$ is 
+Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$, for $i>1$, is 
 \begin{equation*}
@@
 We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number.
-For the base case $i=2$ we must show that the read distance is $a_1-1$.
-Well we know that the first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at \emph{position} $a_1$.
+When $i=2$ we must show that the read distance is $a_1-1$.
+The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$.
 The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.
 
-Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the $l$th block and the number at position $l$ is 
+Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the start of the $l$th block and position $l$ is 
 \begin{equation*}
@@
 Also note that $a_{l+1}$ is one position to the right of $a_l$.
-Therefore the beginning of the $l+1$th block is
+Therefore the beginning of the $(l+1)$th block is
 \begin{equation*}
@@
 \end{proof}
-Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and the read distance is monotonically increasing.
-Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a bound on the number of consecutive times the read distance can stay constant.
+Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing.
+Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a simple bound on the number of consecutive times the read distance can stay constant.
 
-\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max{ a_1, a_2, \ldots, a_n}$.
+\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max(a_1, a_2, \ldots, a_n)$.
 Suppose the read distance for $x_i$ is $d$, then the read distance for $x_{i+a}$ is at least $d+1$.
@@
 Assume otherwise.
-By our Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$.
-By our assumption it must be exactly $d$.
-Lemma ~\ref{readdistformula} also tells us that $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$.
+By Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$.
+Thus it must be exactly $d$.
+Lemma ~\ref{readdistformula} also tells us that if this is the case then $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$.
 But this means that we have a block of length $a+1$, which is a contradiction of the self referencing property of the sequence.

### 2014-10-28 draft.tex
Edits to Iterative Process and Substitution Rules

@@
 One possible iterative process is as follows.
-Let the sequence $\{s^0\}$ be 2, and let the sequence $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$.
+Let the sequence $s^0$ be 2, and let the sequence $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$.
 This gives:
 \begin{align*}
-\{s^{(0)}\} &= 2\\
-\{s^{(1)}\} &= 2,2\\
-\{s^{(2)}\} &= 2,2,1,1\\
-\{s^{(3)}\} &= 2,2,1,1,2,1\\
-\{s^{(4)}\} &= 2,2,1,1,2,1,2,2,1\\
-&\cdots
+s^{(0)}&= 2\\
+s^{(1)} &= 22\\
+s^{(2)} &= 2211\\
+s^{(3)}&= 221121\\
+s^{(4)} &= 221121221\\
+&\vdots
 \end{align*}
 Let $s^\infty$ be the limit of this process. 
-By limit, we mean the limit with respect to finitely many terms $s^{(i)}_j \rightarrow s^\infty_j$ as$i \rightarrow \infty$.
+By limit, we mean the limit with respect to finitely many terms $s^{(i)}_j \rightarrow s^\infty_j$ as $i \rightarrow \infty$.
 We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2.
-\begin{proof}
-We prove this with induction.
-We show that if $s^{(i)}$ is a prefix of $s(\{1,2\},2)$, then $s^{(i+1)}$ is a longer prefix of $s(\{1,2\},2)$.
-The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(\{1,2\},2)$.
-
-For the inductive step, observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence.
-We know that it describes a longer prefix because the prefix it describes has length equal to the sum of the original prefix.
-In other words, the length of $s^{(i+1)}$ must be equal to the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1).
-This holds true because $s^{(i)}$ merely contains a collection of lengths, which are summed up to give the next prefix.
-\end{proof}
-
-It turns out that this same argument holds in general.
+In fact we can prove a more general result about this iterative process.
 
 \begin{theorem}
-Let $\{s^0\} = a$ for some $a \in A$ where $a \neq 1$, and let $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$.
+Let $s^0 = a$ for some $a \in A$ where $a \neq 1$, and let $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$.
 Then the limit of this process, $s^\infty$, is the sequence $s(A,a)$.
@@
 
+\begin{proof}
+We prove this with induction.
+We show that if $s^{(i)}$ is a prefix of $s(A,a)$, then $s^{(i+1)}$ is a longer prefix of $s(A,a)$.
+The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(A,a)$.
+
+For the inductive step, we observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence.
+We know that the sequence it describes is longer because the sequence it describes has length equal to the sum of the original prefix.
+In other words, the length of $s^{(i+1)}$ equals the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1).
+
+The resultant sequence is a prefix of the self referencing sequence because if we apply our process to any prefix of the self referencing sequence it will produce a prefix of the self referencing sequence.
+\end{proof}
+
 Note that we must specify that $a \neq 1$, because that is the only case in which the length of $s^{(i+1)}$ is not strictly greater than the length of $s^{(i)}$.
-The proof of this theorem is omitted, as it is essentially the same proof as for the $A = \{1,2\}, a = 2$ case.
 
-It turns out that we can be less restrictive about what the starting sequence is.
+It turns out that we can be less restrictive about what our starting sequence is.
 Going back to the case of $\{1,2\}$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2.
 For example, suppose $s^{(0)} = 2,1,1,1,2$.
-The rules for generating large $s^{(i)}$ give:
+The rules for generating $s^{(i)}$ give:
 \begin{align*}
-\{s^{(0)}\} &= \mathbf{2},1,1,1,2\\
-\{s^{(1)}\} &= \mathbf{2,2,1},2,1,2,2\\
-\{s^{(2)}\} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
-\{s^{(3)}\} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
-&\cdots
+s^{(0)}&= \mathbf{2},1,1,1,2\\
+s^{(1)} &= \mathbf{2,2,1},2,1,2,2\\
+s^{(2)} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
+s^{(3)} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
+&\vdots
 \end{align*}
-Here, we highlight in bold the numbers that form a prefix of $s(\{1,2\},2)$.
-We note that $s^\infty = s(A,a)$ simply because $s^{(0)}$ contains a 2 at the beginning, whch then causes a 2,2 to appear at the start of $s^{(1)}$, which causes $2,2,1,1$ to appear at the start of $s^{(2)}$, and so on.
-The rest of each $s^{(i)}$ that is not a prefix of $s({1,2},2)$ we call the \emph{tail}.
+Here, we bold the numbers that form a prefix of $s(\{1,2\},2)$.
+We note that $s^\infty = s(A,a)$ because because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(A,a)$ in the limit.
+We call the rest of each $s^{(i)}$ that is not a prefix of $s({1,2},2)$ the \emph{tail}.
 
@@
 However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(A,a)$.
-Thus, we can formalize two types of convergence: \emph{weak convergence} and \emph{strong convergence}.
-Weak convergence is satisfied if $s^\infty = s(A,a)$, and strong convergence is satisfied if $s^{(i)}$ is a prefix of $s(A,a)$ for all $i \geq k$ for some $k$.
+Thus, we formalize two types of convergence
+\begin{definition} We say that a sequence $s^{i}$ is weakly convergent to $s(A,a)$ if $s^\infty = s(A,a)$.
+We say that the sequence $s^{i}$ is strongly convergent if $s^{(i)}$ is a prefix of $s(A,a)$ for all $i \geq k$ for some $k$.
+\end{definition}
+It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence, although based on numerical simulation we make the following conjecture.
+
+\begin{conjecture} A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 2)$ if the starting sequence begins with a $2$ and contains no more than 3 elements.
+A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 1)$ if the starting sequence begins with a $1$ and contains between 2 and 3 elements.
+A sequence $s^i$ is weakly convergent but not strongly convergent to $s(\{1,2\}, a)$ if the starting sequence begins with an $a \in \{1,2,\}$ and contains 4 or more elements.
+\end{conjecture}
+We make this conjecture on the basis on computer simulations which executes the iterative process to at least 20 iterations.
+More iterations were not completed because after this point the simulations began to take impractically long amounts of time to complete.
 
@@
 
-So we can ask, when can we do this? We would like to be able to simulate our original algorithm with this current one. We had established that our algorithm had a state corresponding to 
+We would like to know if we can do this for any generating set $G$ and starting number $g$.
+In particular we would like to be able to simulate our original algorithm with a substitution rule.
+Previously we established that our algorithm had a state corresponding to
 \begin{itemize}
-\item current block length number
-\item how many elements already written down in the block
-\item which number we need to write.
+\item current block size
+\item number of elements already written down in the current block
+\item what number the current block is composed of 
 \end{itemize}
 
-In an algorithm with rules, we cannot remember anything about the previous rules.
-Therefore, each rule needs to encompas the block length numbers, the number of elements already written, and which number we need to write.
-The first condition is trivial to keep, since the rules themselves will say the block length numbers, in order to keep track of the numbers already written, each rule should be self-contained.
-That means that the rules should read off the block numbers.
-We would also need the length of the rules to be divisible by the size of the generating set.
+When we apply our substitution rules we should not need to know what happened to blocks other than the one we are currently substituting for.
+Therefore, each rule must somehow encode the length of  blocks, the number of elements already written in the current block, and what number the current block is composed of.
+The first condition is simple to implement, as the rules will specify the length of blocks.
+In order to keep track of the numbers already written, each rule should be self-contained.
+That is each rule should write complete blocks.
+In order to comply with the third condition it is enough for the size of the generating set to divide the length of the rules.
 This is so that at the start of each rule, we always know which numbers to write.
@@
 \begin{itemize}
-\item each rule reads off the block lengths in the same order of $G$
+\item each rule writes blocks in the same order of $G$
 \item $\sum_{i \in r} i$ is divisible by $l$
@@
 
-Another exaple, we can have $G = \{ 2, 4 \}$, and we can have that the rules be
+Another example, we can have $G = \{ 2, 4 \}$, and we can have that the rules be
 \begin{align*}

### 2014-10-28 draft.tex
Edited Density

@@
 
-In this section, we analyze two things: the first is how to take the rule based methods of sequence generation to show that if a density exists, we can find what it is.
-We will explain the method and give an example applied to the case above of $G = \{ 1, 3\}$ and $G = \{ 2, 4 \}$.
-In the second part, we will show an ad hoc way to bound the density of $G = \{ 1, 2\}$ in a way that suggests that the density converges to 2; however, this method is computationally intractable for smaller bounds since the number of cases to check increases exponentially.
+A topic of interest when considering a self referencing sequence is the relative density of the various elements of the generating set.
+It is unclear if such a density even exists and if so how to compute it for arbitrary generating sets.
+In this section, we show how to use substitution rules to compute relative densities for certain sequences if we assume such densities exists.
+We explain the method and compute the relative density of 3's for $G = \{ 1, 3\}$.
+We will also produce bounds on the density of $2$'s $G = \{ 1, 2\}$ in a way that suggests that the density converges to 2; however, this method is computationally difficult for smaller bounds since the number of cases to check increases exponentially.
 
-\subsection{Determining the density}
-
-We have found a technique to compute the density of numbers in a self-referential sequence (if one exists) for some generating sets.
+To begin we will discuss a technique that can be used to compute the densities for some generating sets.
 In particular, suppose we have a generating set $G$ with a set of rules $R$ that satisfy the conditions from the section above.
@@
 \end{align*}
-Then we can let $3,3,3,1,1,1 = A$, $3,3,3,1 = B$ and $3, 1 = C$. If $A_i$ is the number of occurences of $A$ in $s^{(i)}$, then 
+Then we can let $3,3,3,1,1,1 = A$, $3,3,3,1 = B$ and $3, 1 = C$. If $A_i$ is the number of occurrences of $A$ in $s^{(i)}$, then 
 \[ A_i = B_{i-1} + A_{i-1} \]
@@
 \[ C_i' = \dfrac{A_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
-So if we assume a density, there will be a stable state which will be the density, so in the case where $A_i'=A_{i-1}', B_i' = B_{i-1}', C_i'=C_{i-1}'$, we can solve for the fractional appearence of each and get the density of the numbers.
+So if we assume a density, there will be a stable state which will be the density, so in the case where $A_i'=A_{i-1}', B_i' = B_{i-1}', C_i'=C_{i-1}'$, we can solve for the fractional appearance of each and get the density of the numbers.
 
-In particular, the density of 3's is approxinately $0.602$ by computing the solution to the set of quadratic equations.
+We can compute solutions to this set of quadratic equations and find the density of 3's is approximately $0.602$.
 
-\subsection{Density bounds when $G = \{1, 2\}$}
-
-As shown in the previous section, we cannot establish the set of rules we want when $G = \{1 , 2\}$; however, there is some structure that we can exploit.
-In particular, we know that the sequence $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence.
-This is because they would be part of a block of size at least $3$.
+As discussed in the previous section, it is unclear how to establish a set of rules for $G = \{1 , 2\}$; however, there is some structure that we can exploit.
+In particular, we know that the sequences $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence.
+This is because such sequences would be part of a block of size at least $3$.
 
@@
 So now we can look at how many $2$'s are in any given string of length $9$.
-We know there must be at least $3$, but each consecutive $3$ elements must contain a $3$.
-There are only two cases where there are three 2's in a segment of $9$ numbers in the sequence, namely
+We know there must be at least three $2$'s.
+There are only three cases where there are three 2's in a segment of $9$ numbers in the sequence, namely
 \[ 2, 1, 1, 2, 1, 1 ,2 , 1, 1 \]
@@
 \[ 2, 1, 2, 1, 2 \]
-And these are generated by $1, 1, 1$.
-Which means that none of these are possible, so in addition to having one 2 for every group of 3, there is another 2 for every group of 9.
+but each of these are generated by $1, 1, 1$.
+This means that none of the sequences of length 9 are possible, so in addition to having one 2 for every group of 3, there is another 2 for every group of 9.
 Therefore, we have shown that the density must be bounded by
@@
 We believe that if there is a way to show this, we will be able to show that the density is $\frac{1}{2}$, since the density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number.
-However, generalizing the procedure to improve the bound seems dificult.
+However, generalizing the procedure to improve the bound seems difficult.
 \end{document}

### 2014-10-28 draft.tex
changed margins
added open problems and haikus

@@
 \usepackage{verbatim}
-\usepackage[margin=0.9in]{geometry}
+\usepackage[margin=1.25in]{geometry}
 %\usepackage{amsfonts}
@@
 
-\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+\emph{Written by Fermi Ma and Erik Waingarten, edited by Perry Kleinhenz}
 
@@
 However, generalizing the procedure to improve the bound seems difficult.
+
+\section{Unresolved Problems}
+\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+
+Our work leaves a number of unresolved problems.
+
+\begin{itemize}
+\item In Section~\ref{equivalence}, we propose the idea that in some sense, the number that we start a sequence with does not matter.
+In other words, any sequence that uses a certain generating set is equivalent to any other sequence created by the same generating set.
+\item In the same section, we propose a condition for determining when two sequences are not equivalent.
+We conjecture that if two sequences have unequal numbers with the same read distances, then they are not equivalent.
+It seems from trying cases that this condition is sufficient, but we have been unable to prove it.
+\item In Section~\ref{iterativeprocess} we make a conjecture on when sequences are strongly and weakly convergent.
+We have not yet been able to investigate strong and weak convergence in great detail, as we have been unable to make precise mathematical statements about these types of convergence.
+Most of what we know about them comes from analyzing numerical data.
+\item In Section~\ref{substitutionrules}, we look at substitution rules for certain ``nice" generating sets of numbers, such as $\{1,3\}$ and $\{2,4\}$.
+These sets allow for substitution rules because they satisfy the trio of properties we outline in that section.
+However, we do not know that these are the only properties.
+We leave unresolved the question of determining precisely which sets allow for substitution rules and which ones ones do not.
+The major difficulty here seems to come in proving that a certain generating set does \emph{not} allow for substitution rules.
+\item In Section~\ref{density} we were able to compute the density of any sequence with generating set $\{1,3\}$, and we were able to give bounds on the density of any sequence with generating set $\{1,2\}$.
+However, we leave unresolved the question of what the density for $\{1,2\}$ actually is, or even if we can obtain tighter bounds.
+\end{itemize}
+
+\section{Haikus}
+\emph{Written by Fermi Ma}\\
+
+\noindent One, Two, Two, One, One,\\
+Two, One, Two, Two, One, Two, Two,\\
+One, One, Two, One, One...\\
+
+\noindent For the final draft,\\
+we will improve the haikus\\
+(and the whole paper).
 \end{document}

### 2014-10-31 draft.tex
sdfh

--- 2014-10-28 draft.tex
@@
 In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined.
-After that, we consider how the starting number for the sequence affects the long term behavior of the sequence.
+After that, we consider how the starting number for the sequence affects what the sequence can be.
 In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers.
@@
 We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number.
-To see why this is true, we give an algorithm to build the sequence.
+To see why this is true, we can simply give an algorithm to build the sequence.
 Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$.
@@
 Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block.
-If $a_i$ is not 1, then there is a second element of that block.
-We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence.
+If $a_i$ is not 1, then there is a second element of that sequence.
+We read that second element, and then we add a block to the end of that sequence with length dictated by the value of the second element.
 If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
 
-This procedure continues deterministically so long as there is an element to read.
-Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
+This procedure can continue deterministically so long as there is always an element to read.
+However, every time we write down a block, we read 1 element and write down at least 1 element, so the end of the sequence will always be after the spot we are reading.
+The only exception is when $a_i = 1$, but then since $a_{i+1} \neq 1$, this issue will never occur again.
 \end{quote}
@@
 Thus, a self-referencing sequence is uniquely determined by the generating set and the starting number.
-From this point on, we will denote by $s(G,g)$ the unique sequence generated by set $G$ with starting number $g \in G$.
+From this point on, we will denote by $s(A,a)$ the unique sequence generated by set $A$ with starting number $a \in A$.
 
@@
 \end{equation*}
-for $i$ any positive integer. We say that two sequences are independent if they do not differ by a prefix for any $n$, in either order. 
+for $i$ any positive integer. if no such $n$ exists then we say that the two sequences are independent.
 \end{definition}
-Note that a pair of sequences can differ by multiple prefixes of different length. For example if we have 
+Note that a pair of sequences can have prefixes of different length. For example if we have 
 \begin{align*}
@@
 In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$.
-The following lemma is a direct consequence of our algorithm for generating self-referential sequences:
+The following lemma is a trivial consequence of our algorithm for generating self-referential sequences:
 
-\begin{lemma} Let $G = \{1, g_1, g_2, \ldots, g_n\}$, $s(G, 1)$ and $s(G, g_1)$ differ by a minimal prefix of length at most 1.
+\begin{lemma} Let $C = \{1, c_1, c_2, \ldots, c_n\}$, $s(C, 1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
 \end{lemma}
 \begin{proof}
-Let $a_i$ refer to the $i$th term of $s(G, 1)$ and $b_i$ refer to the $i$th term of $s(G, g_1)$.
-If we use our algorithm for generating the sequence, we note that $a_1 = 1$, and so $a_2 = g_1$, and that this is the beginning of a new block of length $g_1$.
-Furthermore, $b_1 = g_1$ and it indicated the beginning of a block of length $g_1$.
+Let $a_i$ refer to the $i$th term of $s(C, 1)$ and $b_i$ refer to the $i$th term of $s(C, c_1)$.
+If we use our algorithm for generating the sequence, we note that $a_1 = 1$, and so $a_2 = c_1$, and that this is the beginning of a new block of length $c_1$.
+Furthermore, $b_1 = c_1$ and it indicated the beginning of a block of length $c_1$.
 Note that our algorithm is ``memory-less'' in the sense that it does not maintain any state about previous elements.
@@
 The algorithm will generate the same sequence, shifted by 1.
-Therefore, $s(G,1)$ and $s(G, g_1)$ differ by a minimal prefix of length at most 1.
+Therefore, $s(C,1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
 \end{proof}
-We now give a 
+
 \begin{theorem}
 \label{independentsequences}
-Let $G = \{ 1, a, b\}$ where $1<a<b$. $s(G, a)$ and $s(G, b)$ are independent.
+Let $C = \{ 1, a, b\}$ where $1<a<b$. $s(C, a)$ and $s(C, b)$ are independent.
 \end{theorem}
@@
 If the two sequences are not independent then they must differ by some minimal prefix of length $n$.
-If we let $a_i$ be the sequence beginning with $a$ and $b_i$ be the sequence beginning with $b$ then
+If we let $a_i$ be the sequence beginning with $a$ and $b_i$ be the sequence beginning with $b$ the
 \begin{equation*}
@@
 \end{equation*}
-We know that the first $b$ terms of $b_i$ are $b$'s.
+We know that the first $b$ terms of the sequence which begins with a $b$ are $b$'s.
 Because $b>a>1$ no block can be longer then $b$.
-Thus the terms immediately preceding and following this block in  $a_i$ cannot be $b$'s.
-Because of this the first block of $b_i$ is the $m+1$th block of $a_i$, for some $m$ a positive integer.
-Since the value of the sequence at position $k$ is the length of the $k$th block this means that $a_{m+1}=b_1$.
+Thus the terms immediately preceding and following this block in the sequence $a_i$ must not be $b$'s.
+Because of this the first block of $b_i$ is the $m+1$th block of $a_i$, where $m$ is some positive integer.
+Since the value of the sequence at position $k$ is also the length of the $k$th block this means that $a_{m+1}=b_1$.
 In fact the $k$th block of $b_i$ is the $(m+k)$th block of $a_i$ and so
@@
 We now claim that the prefix has length strictly larger than $m$. Assume otherwise so $n \leq m$. 
-This means that $a_{m+1}$ which gives the length of the $m+1$th block occurs at or after the beginning of the $m+1$th block, which begins at position $a_{n+1}$.
+This means the entry in the $a_i$ which gives the length of the $m+1$th block occurs at or after the beginning of the $m+1$th block.
 By our proof that self-referential sequences are well defined we know that the second case cannot occur.
-Therefore we must have the $m+1$th block start at position $m+1$.
-This would require every block to have length exactly 1, but this is a contradiction because the first block has length $a_1=a>1$.
+Therefore the only other option would be for the $m+1$th block to start at position $m+1$, but this would require every block to have length exactly 1, but this is a contradiction because $a_1=a>1$, that is the first block has length greater than one.
 Therefore the prefix must have length strictly larger than $m$.
@@
 In this section, we consider a slightly weaker definition of equivalence.
-Consider, for example, sequences generated by $G = \{1,2,3\}$.
+Consider, for example, sequences generated by $C = \{1,2,3\}$.
 Theorem~\ref{independentsequences} shows that the sequence starting with 2
@@
 Lets look at the state of the algorithm that generates both sequences.
-After the first block, the execution of the algorithm of $s(G, 2)$ right after placing the first $3$'s (in the 3rd and 4th position), the algorithm is in state corresponding to:
+After the first block, the execution of the algorithm of $s(C, 2)$ right after placing the $3$'s (in the 4th and 5th position), the algorithm is in state corresponding to:
 \begin{itemize}
 \item block size $3$
-\item no elements in the current block
+\item no elements on the current block
 \item starting number is $1$
 \end{itemize}.   
-If we look at the execution of the algorithm that generates $s(G, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
+If we look at the execution of the algorithm that generates $s(C, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
 \begin{itemize}
 \item block size $3$
-\item no elements in the current block
+\item no elements on the current block
 \item starting number is $1$
 \end{itemize}
-Therefore, since the states are the same, the sequence generated after the 4th position of $s(G, 2)$ and after the $3rd$ position of $s(G, 3)$ will be the same.
-By inspection, we see we can also include two of the 3's.
+Therefore, since the states are the same, the same sequence will be generated after the 4th position of $s(C, 2)$ and after the $3rd$ position of $s(C, 3)$ will be the same.
+By inspection, we can see we can include $2$ of the 3's.
 
-Thus, since both sequences are essentially the same if a prefix is removed from each of them, we call these sequences \emph{equivalent}.
+Thus, since both sequences are essentially the same if certain prefixes are removed, we call these sequences \emph{equivalent}.
 To formalize, this, we say that two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n,k \geq 0$ such that
@@
 For generating sets with larger numbers, we have been unable to verify or refute the conjecture.
-However, in Section~\ref{density}, we show that for some generating sets the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
+However, in Section~\ref{Density}, we show that the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
 
 We now introduce a notion which will help us categorize sequences as equivalent. 
-\begin{definition} Consider a sequence $a$. We define the read distance for an element $a_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
+\begin{definition} Consider a sequence $x$.
+We define the read distance for an element $x_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
+We denote this as $Rd(x_i)$.
 \end{definition}
@@
 \end{align*}
-the read distance for $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number.
+the read distance for the $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number.
 We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0,1,2,4,6,6,6,6,7,8,9,\ldots$.
 From this we might suspect that the read distance for any self referencing sequence will also be monotonically increasing.
-We can in fact compute the read distance exactly.
+We can in fact show a much stronger result.
 \begin{lemma} 
 \label{readdistformula}
-Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$, for $i>1$, is 
+Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$ for $i>1$ is 
 \begin{equation*}
@@
 We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number.
-When $i=2$ we must show that the read distance is $a_1-1$.
-The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$.
+For the base case $i=2$ we must show that the read distance is $a_1-1$.
+Well we know that the first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at \emph{position} $a_1$.
 The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.
 
-Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the start of the $l$th block and position $l$ is 
+Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the $l$th block and the number at position $l$ is 
 \begin{equation*}
@@
 Also note that $a_{l+1}$ is one position to the right of $a_l$.
-Therefore the beginning of the $(l+1)$th block is
+Therefore the beginning of the $l+1$th block is
 \begin{equation*}
@@
 \end{proof}
-Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing.
-Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a simple bound on the number of consecutive times the read distance can stay constant.
+Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and the read distance is monotonically increasing.
+Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a bound on the number of consecutive times the read distance can stay constant.
 
-\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max(a_1, a_2, \ldots, a_n)$.
+\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max{ a_1, a_2, \ldots, a_n}$.
 Suppose the read distance for $x_i$ is $d$, then the read distance for $x_{i+a}$ is at least $d+1$.
@@
 Assume otherwise.
-By Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$.
-Thus it must be exactly $d$.
-Lemma ~\ref{readdistformula} also tells us that if this is the case then $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$.
+By our Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$.
+By our assumption it must be exactly $d$.
+Lemma ~\ref{readdistformula} also tells us that $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$.
 But this means that we have a block of length $a+1$, which is a contradiction of the self referencing property of the sequence.
@@
 One possible iterative process is as follows.
-Let the sequence $s^0$ be 2, and let the sequence $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$.
+Let the sequence $\{s^0\}$ be 2, and let the sequence $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$.
 This gives:
 \begin{align*}
-s^{(0)}&= 2\\
-s^{(1)} &= 22\\
-s^{(2)} &= 2211\\
-s^{(3)}&= 221121\\
-s^{(4)} &= 221121221\\
-&\vdots
+\{s^{(0)}\} &= 2\\
+\{s^{(1)}\} &= 2,2\\
+\{s^{(2)}\} &= 2,2,1,1\\
+\{s^{(3)}\} &= 2,2,1,1,2,1\\
+\{s^{(4)}\} &= 2,2,1,1,2,1,2,2,1\\
+&\cdots
 \end{align*}
 Let $s^\infty$ be the limit of this process.
-By limit, we mean the limit with respect to finitely many terms $s^{(i)}_j \rightarrow s^\infty_j$ as $i \rightarrow \infty$.
+By limit, we mean the limit with respect to finitely many terms $s^{(i)}_j \rightarrow s^\infty_j$ as$i \rightarrow \infty$.
 We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2.
-In fact we can prove a more general result about this iterative process.
+\begin{proof}
+We prove this with induction.
+We show that if $s^{(i)}$ is a prefix of $s(\{1,2\},2)$, then $s^{(i+1)}$ is a longer prefix of $s(\{1,2\},2)$.
+The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(\{1,2\},2)$.
+
+For the inductive step, observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence.
+We know that it describes a longer prefix because the prefix it describes has length equal to the sum of the original prefix.
+In other words, the length of $s^{(i+1)}$ must be equal to the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1).
+This holds true because $s^{(i)}$ merely contains a collection of lengths, which are summed up to give the next prefix.
+\end{proof}
+
+It turns out that this same argument holds in general.
 
 \begin{theorem}
-Let $s^0 = a$ for some $a \in A$ where $a \neq 1$, and let $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$.
+Let $\{s^0\} = a$ for some $a \in A$ where $a \neq 1$, and let $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$.
 Then the limit of this process, $s^\infty$, is the sequence $s(A,a)$.
@@
 
-\begin{proof}
-We prove this with induction.
-We show that if $s^{(i)}$ is a prefix of $s(A,a)$, then $s^{(i+1)}$ is a longer prefix of $s(A,a)$.
-The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(A,a)$.
+Note that we must specify that $a \neq 1$, because that is the only case in which the length of $s^{(i+1)}$ is not strictly greater than the length of $s^{(i)}$.
+The proof of this theorem is omitted, as it is essentially the same proof as for the $A = \{1,2\}, a = 2$ case.
 
-For the inductive step, we observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence.
-We know that the sequence it describes is longer because the sequence it describes has length equal to the sum of the original prefix.
-In other words, the length of $s^{(i+1)}$ equals the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1).
-
-The resultant sequence is a prefix of the self referencing sequence because if we apply our process to any prefix of the self referencing sequence it will produce a prefix of the self referencing sequence.
-\end{proof}
-
-Note that we must specify that $a \neq 1$, because that is the only case in which the length of $s^{(i+1)}$ is not strictly greater than the length of $s^{(i)}$.
-
-It turns out that we can be less restrictive about what our starting sequence is.
+It turns out that we can be less restrictive about what the starting sequence is.
 Going back to the case of $\{1,2\}$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2.
 For example, suppose $s^{(0)} = 2,1,1,1,2$.
-The rules for generating $s^{(i)}$ give:
+The rules for generating large $s^{(i)}$ give:
 \begin{align*}
-s^{(0)}&= \mathbf{2},1,1,1,2\\
-s^{(1)} &= \mathbf{2,2,1},2,1,2,2\\
-s^{(2)} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
-s^{(3)} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
-&\vdots
+\{s^{(0)}\} &= \mathbf{2},1,1,1,2\\
+\{s^{(1)}\} &= \mathbf{2,2,1},2,1,2,2\\
+\{s^{(2)}\} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
+\{s^{(3)}\} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
+&\cdots
 \end{align*}
-Here, we bold the numbers that form a prefix of $s(\{1,2\},2)$.
-We note that $s^\infty = s(A,a)$ because because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(A,a)$ in the limit.
-We call the rest of each $s^{(i)}$ that is not a prefix of $s({1,2},2)$ the \emph{tail}.
+Here, we highlight in bold the numbers that form a prefix of $s(\{1,2\},2)$.
+We note that $s^\infty = s(A,a)$ simply because $s^{(0)}$ contains a 2 at the beginning, whch then causes a 2,2 to appear at the start of $s^{(1)}$, which causes $2,2,1,1$ to appear at the start of $s^{(2)}$, and so on.
+The rest of each $s^{(i)}$ that is not a prefix of $s({1,2},2)$ we call the \emph{tail}.
 
@@
 However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(A,a)$.
-Thus, we formalize two types of convergence
-\begin{definition} We say that a sequence $s^{i}$ is weakly convergent to $s(A,a)$ if $s^\infty = s(A,a)$.
-We say that the sequence $s^{i}$ is strongly convergent if $s^{(i)}$ is a prefix of $s(A,a)$ for all $i \geq k$ for some $k$.
-\end{definition}
-It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence, although based on numerical simulation we make the following conjecture.
-
-\begin{conjecture} A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 2)$ if the starting sequence begins with a $2$ and contains no more than 3 elements.
-A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 1)$ if the starting sequence begins with a $1$ and contains between 2 and 3 elements.
-A sequence $s^i$ is weakly convergent but not strongly convergent to $s(\{1,2\}, a)$ if the starting sequence begins with an $a \in \{1,2,\}$ and contains 4 or more elements.
-\end{conjecture}
-We make this conjecture on the basis on computer simulations which executes the iterative process to at least 20 iterations.
-More iterations were not completed because after this point the simulations began to take impractically long amounts of time to complete.
+Thus, we can formalize two types of convergence: \emph{weak convergence} and \emph{strong convergence}.
+Weak convergence is satisfied if $s^\infty = s(A,a)$, and strong convergence is satisfied if $s^{(i)}$ is a prefix of $s(A,a)$ for all $i \geq k$ for some $k$.
 
@@
 
-We would like to know if we can do this for any generating set $G$ and starting number $g$.
-In particular we would like to be able to simulate our original algorithm with a substitution rule.
-Previously we established that our algorithm had a state corresponding to
+So we can ask, when can we do this? We would like to be able to simulate our original algorithm with this current one. We had established that our algorithm had a state corresponding to 
 \begin{itemize}
-\item current block size
-\item number of elements already written down in the current block
-\item what number the current block is composed of 
+\item current block length number
+\item how many elements already written down in the block
+\item which number we need to write.
 \end{itemize}
 
-When we apply our substitution rules we should not need to know what happened to blocks other than the one we are currently substituting for.
-Therefore, each rule must somehow encode the length of  blocks, the number of elements already written in the current block, and what number the current block is composed of.
-The first condition is simple to implement, as the rules will specify the length of blocks.
-In order to keep track of the numbers already written, each rule should be self-contained.
-That is each rule should write complete blocks.
-In order to comply with the third condition it is enough for the size of the generating set to divide the length of the rules.
+In an algorithm with rules, we cannot remember anything about the previous rules.
+Therefore, each rule needs to encompas the block length numbers, the number of elements already written, and which number we need to write.
+The first condition is trivial to keep, since the rules themselves will say the block length numbers, in order to keep track of the numbers already written, each rule should be self-contained.
+That means that the rules should read off the block numbers.
+We would also need the length of the rules to be divisible by the size of the generating set.
 This is so that at the start of each rule, we always know which numbers to write.
@@
 \begin{itemize}
-\item each rule writes blocks in the same order of $G$
+\item each rule reads off the block lengths in the same order of $G$
 \item $\sum_{i \in r} i$ is divisible by $l$
@@
 
-Another example, we can have $G = \{ 2, 4 \}$, and we can have that the rules be
+Another exaple, we can have $G = \{ 2, 4 \}$, and we can have that the rules be
 \begin{align*}
@@
 However, generalizing the procedure to improve the bound seems dificult.
+
 \end{document}

### 2014-10-31 draft.tex

--- 2014-10-28 draft.tex
@@
 Consider the sequence
-\begin{align*}
+\begin{equation}
+\label{eq:onetwo}
 1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
-\end{align*}
-We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers:
-\begin{align*}
-[1],[2,2],[1,1],[2],[1],[2,2],[1],[2,2],[1,1],[2],[1,1],[2,2],\dots
-\end{align*}
-The block lengths, read from left to right, reproduce the original sequence. Thus, this sequence is \emph{self-referential}. For the remainder of this paper we study properties of such sequences.
+\end{equation}
+We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers.
+\begin{figure}
+\label{fig:sequenceblocks}
+\center
+\includegraphics[width=0.5\linewidth]{sequenceBlocks.pdf}
+\caption{Sequence broken up into contiguous blocks with lengths generating same sequence.}
+\end{figure}
+The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). 
+Thus, this sequence is \emph{self-referential}.
 
-We say that this sequence is \emph{generated} by the set of numbers used in the sequence, so the above sequence is generated by $\{1,2\}$.
-We will refer to this set as the \emph{generating set}.
-Observe that it is unambiguous which number to use next.
-After a block of 1's, the next block must be a block of 2's.
-There does exist ambiguity in a sequence with generating set $\{1,2,3\}$, since after a block of 1's, ther can be a block of 2's or a block of 3's.
-We assign an order to the generating set and resolve the ambiguity.
-The order we will use will be the order of the elements presented, for example, $\{1, 2, 3 \}$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.
-Note that because order matters, the \emph{generating set} is technically not a set, but we do not call it a ``sequence" to avoid possible confusion with its corresponding self-referencing sequence.
+The sequence~\ref{eq:onetwo} contains the numbers $1$ and $2$.
+In general, we want to make self-refering sequences containing other numbers as well.
+Once there is more than two numbers in the sequence, generating the sequence becomes a bit ambiguous.
+Take the simple example of a self-referencing sequence with numbers 1, 2, and 3.
 
-With this rule, a self-referencing sequence generated by $\{1,2,3\}$ has blocks of 1's followed by blocks of 2's followed by blocks of 3's, followed by blocks of 1's, and so on.
-Thus, $\{1,2,3\}$ can generate
-\begin{align*}
-1,2,2,3,3,1,1,1,2,2,2,3,2,1,2,2,3,3,1,1,\dots
-\end{align*}
-but not
-\begin{align*}
-1,3,3,3,2,2,2,1,1,1,3,3,2,2,1,1,3,2,1,3,\dots
-\end{align*}
-as the second sequence has the numbers in the wrong order.
+If the first number is 1, then we can make the second number either a 2, or a 3. In particular, both:
+\begin{equation}
+1, 3, 3, 3, 2, 2, 2, \dots
+\end{equation}
+\begin{equation}
+1, 2, 2, 1, 1, 3, 1, \dots\\
+\end{equation}
+are in some sense ``self-referencing'' sequences.
+Each time we need to specify a new block, we have a choice of two elements.
+We would like to avoid having a large number of choices in the generation of the sequence, and so we require that the sequence contain elements from a \emph{cycle}, and each time we start another block, the block will contain the next number in the cycle.
+In the case of sequence~\ref{eq:onetwo}, the cycle is $(1,2)$.
+
+The self-referencing sequence with cycle $(1, 2, 3)$ that begins with $1$ is
+\begin{figure}
+\label{fig:onetwothree}
+\center
+\includegraphics[width=0.5\linewidth]{sequenceBlocks123.pdf}
+\caption{Sequence generated by $(1,2,3)$ broken up into contiguous blocks.}
+\end{figure}
+
+Formally, we say that a sequence is \emph{generated} by the set of numbers used in the sequence.
+So sequence~\ref{eq:onetwo} is generated by the set $\{1, 2\}$.
+We will refer to the cycle used in the sequence as the \emph{generating cycle}.
+
+Using the cycle, it is no longer ambiguous which number to use in the next block.
 
@@
 
-In this paper, we address a variety of questions regarding these sequences.
+We address a variety of questions regarding these sequences. Our main results are:
+\begin{itemize}
+\item The self-referencing sequence of a  given generating cycle is uniquely determined by its first number.
+\item Some characterizations of sequences with the same generating cycle.
+\item There exists an equivalent formulation of self-referencing sequences using substitution rules for some criteria of the generating cycles.
+\item One can compute the limiting density for the numbers in the sequence if a limiting density exists, for generating cycles that have substitution rules. 
+\item The limiting densities of the self-referencing sequences of generating cycles $(1,3)$ if it exists.
+\end{itemize}
+
 In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined.
-After that, we consider how the starting number for the sequence affects the long term behavior of the sequence.
+We consider how the starting number for the sequence affects the long term behavior of the sequence.
 In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers.
-In the following Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.
+In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.
 
@@
 In Section~\ref{substitutionrules}, we show how a large number of these sequences can actually be generated with a simple set of substitution rules.
-We then introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating sets.
+We introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating sets.
 Unfortunately, we find that the open problem of determining the density for the sequence of 1's and 2's is hard to solve using our methods.
@@
 \end{document}  
+>>>>>>> 94de583213b09230c51ca57ed5c9897bcabc4666

### 2014-11-01 draft.tex
edited up to prefixes, added a lemma showing sequences are unique

@@
 We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number.
-To see why this is true, we give an algorithm to build the sequence.
-Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$.
-If we ever write down some $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
+
+\begin{lemma}
+Let $C$ be a generating cycle. Suppose $(a_i)$ and $(b_i)$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
+\end{lemma}
+
+\begin{proof}
+Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $(a_i)$ and $(b_i)$ are different.
+Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $(a_i)$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $(b_i)$, we get that $b_j$ comes after $a_{j-1}$.
+This is a contradiction since $a_j \neq b_j$.
+So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block.
+Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block.
+However, this means that $a_k \geq b_k + 1$.
+So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$.
+Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.
+
+\begin{figure}{hb}
+\label{fig:unequalsequences}
+\center
+\includegraphics[width=0.5\linewidth]{lemma1.pdf}
+\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
+\end{figure}
+
+This implies that every block before the $k$th block just contains 1's since each block refered it itself.
+This means that either $j = 2$, or the sequence is not self-referential.
+But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle.
+\end{proof}
+
+We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element.
+Assume the generating cycle is $(a_1,a_2,\dots,a_m)$ and that the starting element is some $a_i$.
+For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
 \begin{quote}
@@
 
-Thus, a self-referencing sequence is uniquely determined by the generating set and the starting number.
-From this point on, we will denote by $s(G,g)$ the unique sequence generated by set $G$ with starting number $g \in G$.
+From this point on, we will denote by $s(C,a)$ the unique sequence with generating cycle $C$ and with starting number $a \in C$.
 

### 2014-11-01 draft.tex
edited up to iterative process, removed prefixes

@@
 
-\section{Prefixes}
-\label{prefixes}
-\emph{Written by Perry Kleinhenz, edited by Fermi Ma and Erik Waingarten}
-\begin{definition} We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a prefix of length $n$ if 
-\begin{equation*}
-a_{n+i} = b_{i},
-\end{equation*}
-for $i$ any positive integer. We say that two sequences are independent if they do not differ by a prefix for any $n$, in either order. 
-\end{definition}
-Note that a pair of sequences can differ by multiple prefixes of different length. For example if we have 
-\begin{align*}
-&\{a_i\} = \{1,2,3,4,5,6,1,2,3,4,5,6, \ldots \} \\  
-&\{b_i\}= \{7,8,9,1,2,3,4,5,6,1,2,3,4,5,6, \ldots\},
-\end{align*}
- then $a_i$ and $b_i$ differ by prefixes of length $3+6k$ for $k$ a nonnegative integer. Because of this we make the following definition.
- \begin{definition} 
- We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a minimal prefix of length $n$ if they differ by a prefix of length $n$, but do not differ by a prefix of length $m$ for all $0<m<n$.
- \end{definition}
-We note that if a pair of sequences differs by a prefix they differ by a minimal prefix.
-In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$.
-The following lemma is a direct consequence of our algorithm for generating self-referential sequences:
-
-\begin{lemma} Let $G = \{1, g_1, g_2, \ldots, g_n\}$, $s(G, 1)$ and $s(G, g_1)$ differ by a minimal prefix of length at most 1.
-\end{lemma}
-\begin{proof}
-Let $a_i$ refer to the $i$th term of $s(G, 1)$ and $b_i$ refer to the $i$th term of $s(G, g_1)$.
-If we use our algorithm for generating the sequence, we note that $a_1 = 1$, and so $a_2 = g_1$, and that this is the beginning of a new block of length $g_1$.
-Furthermore, $b_1 = g_1$ and it indicated the beginning of a block of length $g_1$.
-Note that our algorithm is ``memory-less'' in the sense that it does not maintain any state about previous elements.
-Because the number and the begining of the block is the same.
-The algorithm will generate the same sequence, shifted by 1.
-Therefore, $s(G,1)$ and $s(G, g_1)$ differ by a minimal prefix of length at most 1.
-\end{proof}
-We now give a 
-\begin{theorem}
-\label{independentsequences}
-Let $G = \{ 1, a, b\}$ where $1<a<b$. $s(G, a)$ and $s(G, b)$ are independent.
-\end{theorem}
-\begin{proof}
-We proceed by contradiction.
-If the two sequences are not independent then they must differ by some minimal prefix of length $n$.
-If we let $a_i$ be the sequence beginning with $a$ and $b_i$ be the sequence beginning with $b$ then
-\begin{equation*}
-a_{n+k} = b_{k}.
-\end{equation*}
-We know that the first $b$ terms of $b_i$ are $b$'s.
-Because $b>a>1$ no block can be longer then $b$.
-Thus the terms immediately preceding and following this block in  $a_i$ cannot be $b$'s.
-Because of this the first block of $b_i$ is the $m+1$th block of $a_i$, for some $m$ a positive integer.
-Since the value of the sequence at position $k$ is the length of the $k$th block this means that $a_{m+1}=b_1$.
-In fact the $k$th block of $b_i$ is the $(m+k)$th block of $a_i$ and so
-\begin{equation*}
-a_{m+k} = b_k
-\end{equation*}
-
-We now claim that the prefix has length strictly larger than $m$. Assume otherwise so $n \leq m$. 
-This means that $a_{m+1}$ which gives the length of the $m+1$th block occurs at or after the beginning of the $m+1$th block, which begins at position $a_{n+1}$.
-By our proof that self-referential sequences are well defined we know that the second case cannot occur.
-Therefore we must have the $m+1$th block start at position $m+1$.
-This would require every block to have length exactly 1, but this is a contradiction because the first block has length $a_1=a>1$.
-Therefore the prefix must have length strictly larger than $m$.
-
-We note that based on a previous step the two sequences differ by a prefix starting at $m$.
-Since $m<n$ we have produced a prefix which starts before $n$ which is a contradiction.
-Therefore the two sequences are independent.
-\end{proof}
-
 \section{Equivalence}
@@
 
-In this section, we consider a slightly weaker definition of equivalence.
-Consider, for example, sequences generated by $G = \{1,2,3\}$.
+In this section, we consider a different definition of equivalence.
+Consider, for example, sequences generated by $C = (1,2,3)$.
 Theorem~\ref{independentsequences} shows that the sequence starting with 2
@@
 \begin{align*}
-3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+2,2,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
+3,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
 \end{align*}
 in both cases.
-Lets look at the state of the algorithm that generates both sequences.
-After the first block, the execution of the algorithm of $s(G, 2)$ right after placing the first $3$'s (in the 3rd and 4th position), the algorithm is in state corresponding to:
-\begin{itemize}
-\item block size $3$
-\item no elements in the current block
-\item starting number is $1$
-\end{itemize}.   
-If we look at the execution of the algorithm that generates $s(G, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
-\begin{itemize}
-\item block size $3$
-\item no elements in the current block
-\item starting number is $1$
-\end{itemize}
-Therefore, since the states are the same, the sequence generated after the 4th position of $s(G, 2)$ and after the $3rd$ position of $s(G, 3)$ will be the same.
-By inspection, we see we can also include two of the 3's.
 
 Thus, since both sequences are essentially the same if a prefix is removed from each of them, we call these sequences \emph{equivalent}.
-To formalize, this, we say that two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n,k \geq 0$ such that
+To formalize, this, we say that two sequences $(a_i)$ and $(b_i)$ are equivalent if there exists some integers $n,k \geq 0$ such that
 \begin{align*}
@@
 \end{align*}
-
-We now give the following conjecture
+for all positive integers $i$. This definition is saying that the sequences are eventually equivalent.
 
@@
 
-This conjecture is partly motivated by the fact that the property holds for $\{1,2,3\}$.
-For generating sets with larger numbers, we have been unable to verify or refute the conjecture.
+This conjecture is partly motivated by the fact that the property holds for self-referencing sequences from the generating cycle $C= (1,2,3)$.
+For generating cycles with larger numbers, we have been unable to verify or refute the conjecture..
 However, in Section~\ref{density}, we show that for some generating sets the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
@@
 We now introduce a notion which will help us categorize sequences as equivalent. 
-\begin{definition} Consider a sequence $a$. We define the read distance for an element $a_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
+\begin{definition} Consider a sequence $(a_i)$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
 \end{definition}
-The read distance can be any nonnegative integer. For example if we consider the sequence $a_i$ generated by $\{1,2,3\}$ beginning with a 2  
+The read distance can be any nonnegative integer. For example if we consider the sequence $(a_i)$ generated by $(1,2,3)$ beginning with a 1  
 \begin{align*}
-2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+1,2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
 \end{align*}
-the read distance for $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number.
-We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0,1,2,4,6,6,6,6,7,8,9,\ldots$.
-From this we might suspect that the read distance for any self referencing sequence will also be monotonically increasing.
-We can in fact compute the read distance exactly.
+the read distance for $1$ is always $0$ since in a self-referencing sequence, the first number must refer to the first block.
+The read distance for $2$ is $0$ as the second block of numbers starts at the second number.
+We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0, 0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.
+The read distance, can be thought of as the horizontal shift of the arrows in Figure~\ref{fig:onetwothree}.
+
 \begin{lemma} 
 \label{readdistformula}
-Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$, for $i>1$, is 
+Let $(a_i)$ be any self referencing sequence generated by some set of positive integers. Then the read distance for $i>1$, is 
 \begin{equation*}
@@
 \end{equation*} 
+and the read distance for $1$ is always 0.
 \end{lemma}
@@
 We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number.
-When $i=2$ we must show that the read distance is $a_1-1$.
+When $i=2$ the read distance is $a_1-1$.
 The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$.
@@
 
-\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max(a_1, a_2, \ldots, a_n)$.
-Suppose the read distance for $x_i$ is $d$, then the read distance for $x_{i+a}$ is at least $d+1$.
+\begin{lemma} Consider a self referencing sequence $(a_i)$ generated by $(c_1, c_2, \ldots, c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$.
+Suppose the read distance for $i$ is $d$, then the read distance for $i+m$ is at least $d+1$.
 \end{lemma}
 \begin{proof}
-Assume otherwise.
-By Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$.
-Thus it must be exactly $d$.
-Lemma ~\ref{readdistformula} also tells us that if this is the case then $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$.
-But this means that we have a block of length $a+1$, which is a contradiction of the self referencing property of the sequence.
-Therefore the read distance for $x_{i}$ is at least $d+1$.
+The read distances can stay constant from $i$ to $i+1$ only if there is a $a_{i+1} = 1$.
+There can be at most $m$ 1's in a contiguous block, so the read distance of $i+m$ is bigger than the read distance of $i$.
 \end{proof}
@@
 We also have a sufficient condition for equivalence 
-\begin{lemma} If two sequences $a_i, b_i$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are the same and 
+\begin{lemma} If two sequences $(a_i), (b_i)$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are the same and 
 \begin{equation*}
@@
 \end{equation*}
-where $k$ ranges from 0 to the read distance for $a_m$ then $a_i$ and $b_i$ are equivalent. 
+where $k$ ranges from 0 to the read distance for $a_m$ then $(a_i)$ and $(b_i)$ are equivalent. 
 \end{lemma}
@@
 Therefore we have $a_{m+k+1} = b_{l+k+1}$ for this case as well.
-This completes our induction and so $a_i$ and $b_i$ are equivalent.
+This completes our induction and so $(a_i)$ and $(b_i)$ are equivalent.
 \end{proof}
+
+The following is a simple observation of sequences that begin with $1$.
+
+\begin{corollary} Let $C = (1, c_1, c_2, \ldots, c_n)$. $s(C, 1)$ and $s(C, c_1)$ are equivalent.
+\end{corollary}
+\begin{proof}
+Let $(a_i) = s(C, 1)$ and $(b_i) = s(C, c_1)$.
+The read distance of sequence $(a_i)$ for $i=2$ is 0, and the read distance of sequence $(b_i)$ for $i = 1$ is 0.
+In addition, $a_2 = b_1 = c_1$, which means the two sequences are equivalent.
+\end{proof}
+
 These conditions for equivalence lead us to make the following conjecture on a sufficient condition for nonequivalence.
 
-\begin{conjecture} If two sequences $a_i$ and $b_i$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
+\begin{conjecture} If two sequences $(a_i)$ and $(b_i)$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
 \end{conjecture}

### 2014-11-02 draft.tex
Made Changes to Introduction

@@
 
-\subsection{Problem Setup}
 Consider the sequence
@@
 \end{equation}
-We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers.
+We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers:
 \begin{figure}
@@
 \end{figure}
-The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). 
-Thus, this sequence is \emph{self-referential}.
 
-The sequence~\ref{eq:onetwo} contains the numbers $1$ and $2$.
-In general, we want to make self-refering sequences containing other numbers as well.
-Once there is more than two numbers in the sequence, generating the sequence becomes a bit ambiguous.
-Take the simple example of a self-referencing sequence with numbers 1, 2, and 3.
+The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). Thus, this sequence is \emph{self-referential}.
 
+Observe that in this case it is unambiguous which number to use next. After a block of 1's, the next block must be a block of 2's. However ambiguity exists in a sequence generated by $(1,2,3)$.
 If the first number is 1, then we can make the second number either a 2, or a 3. In particular, both:
@@
 Each time we need to specify a new block, we have a choice of two elements.
-We would like to avoid having a large number of choices in the generation of the sequence, and so we require that the sequence contain elements from a \emph{cycle}, and each time we start another block, the block will contain the next number in the cycle.
-In the case of sequence~\ref{eq:onetwo}, the cycle is $(1,2)$.
+We would like to avoid having a large number of choices in the generation of the sequence, and so we will use the convention that the order in which the elements are presented gives the order they are used in the sequence.
+For example, $(1, 2, 3 )$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.
 
-The self-referencing sequence with cycle $(1, 2, 3)$ that begins with $1$ is
+With this rule, the self-referencing sequence generated by $(1,2,3)$ that begins with a $1$ is
 \begin{figure}
@@
 
-Formally, we say that a sequence is \emph{generated} by the set of numbers used in the sequence.
-So sequence~\ref{eq:onetwo} is generated by the set $\{1, 2\}$.
-We will refer to the cycle used in the sequence as the \emph{generating cycle}.
+We call the set the sequence is generated over, along with its order the \emph{generating cycle}.
 
-Using the cycle, it is no longer ambiguous which number to use in the next block.
-
-\subsection{Overview}
-
-We address a variety of questions regarding these sequences. Our main results are:
+Our main results are:
 \begin{itemize}
-\item The self-referencing sequence of a  given generating cycle is uniquely determined by its first number.
+\item The self-referencing sequence for a given generating cycle is uniquely determined by its first number.
 \item Some characterizations of sequences with the same generating cycle.
-\item There exists an equivalent formulation of self-referencing sequences using substitution rules for some criteria of the generating cycles.
-\item One can compute the limiting density for the numbers in the sequence if a limiting density exists, for generating cycles that have substitution rules. 
-\item The limiting densities of the self-referencing sequences of generating cycles $(1,3)$ if it exists.
+\item For some generating cycles there exists an equivalent formulation of self-referencing sequences using substitution rules
+\item If a generating cycle has substitution rules then one can compute the limiting density for the numbers in the sequencem if a limiting density exists.
+\item The limiting densities of the self-referencing sequences of  $(1,3)$ if such densities exists.
 \end{itemize}
@@
 We consider how the starting number for the sequence affects the long term behavior of the sequence.
-In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers.
 In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.

### 2014-11-03 draft.tex
edited iterative process

@@
 
-We change focus in this section and consider how to generate self-referential sequences with an iterative process. We first look at the sequences generated by $\{1,2\}$.
+We change focus in this section and consider how to generate self-referential sequences with an iterative process. We first look at the sequences generated by the cycle $C = (1,2)$.
 
@@
 Let $s^\infty$ be the limit of this process. 
-By limit, we mean the limit with respect to finitely many terms $s^{(i)}_j \rightarrow s^\infty_j$ as $i \rightarrow \infty$.
-We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2.
-In fact we can prove a more general result about this iterative process.
+By limit, we mean the limit with respect to each term seperately, so we can define
+\[ s^{(i)}_j \rightarrow s^\infty_j \]
+as $i \rightarrow \infty$. We claim that $s^\infty$ is the unique self-referential sequence $s(C,2)$. In fact we can prove a more general result about this iterative process.
 
 \begin{theorem}
-Let $s^0 = a$ for some $a \in A$ where $a \neq 1$, and let $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$.
-Then the limit of this process, $s^\infty$, is the sequence $s(A,a)$.
+Let $C$ be some generating cycle.
+Let $s^0 = a$ for some $a \in C$ where $a \neq 1$, and let $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$.
+Then the limit of this process, $s^\infty$, is the sequence $s(C,a)$.
 \end{theorem}
@@
 \begin{proof}
-We prove this with induction.
-We show that if $s^{(i)}$ is a prefix of $s(A,a)$, then $s^{(i+1)}$ is a longer prefix of $s(A,a)$.
-The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(A,a)$.
+We prove this by induction on $i$.
+We show that if $s^{(i)}$ is a prefix of $s(C,a)$, then $s^{(i+1)}$ is a longer prefix of $s(C,a)$.
+The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(C,a)$.
 
@@
 
-It turns out that we can be less restrictive about what our starting sequence is.
-Going back to the case of $\{1,2\}$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2.
-For example, suppose $s^{(0)} = 2,1,1,1,2$.
-The rules for generating $s^{(i)}$ give:
+It turns out that we can be less restrictive about what our starting sequence, $s^{(0)}$, is.
+Going back to the case of $C = (1,2)$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2.
+And still, the limiting sequence will be the self-referential sequence.
+For example, suppose $s^{(0)} = 2,1,1,1,2$. The rules for generating $s^{(i)}$ give:
 \begin{align*}
@@
 \end{align*}
-Here, we bold the numbers that form a prefix of $s(\{1,2\},2)$.
-We note that $s^\infty = s(A,a)$ because because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(A,a)$ in the limit.
-We call the rest of each $s^{(i)}$ that is not a prefix of $s({1,2},2)$ the \emph{tail}.
+Here, we bold the numbers that form a prefix of $s(C,2)$.
+We note that $s^\infty = s(C,a)$ because because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(C,a)$ in the limit.
+We call the rest of each $s^{(i)}$ that is not a prefix of $s(C,2)$ the \emph{tail}.
 
 We now consider two different forms of convergence.
-We know that $s^{(i)}$ in some sense approaches $s(A,a)$, as arbitrarily many elements at the beginning of the sequence will match up with the beginning of $s(A,a)$ as $i$ increases.
-However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(A,a)$.
+We know that $s^{(i)}$ in some sense approaches $s(C,a)$, as arbitrarily many elements at the beginning of the sequence will match up with the beginning of $s(C,a)$ as $i$ increases.
+However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(C,a)$.
 Thus, we formalize two types of convergence
-\begin{definition} We say that a sequence $s^{i}$ is weakly convergent to $s(A,a)$ if $s^\infty = s(A,a)$.
-We say that the sequence $s^{i}$ is strongly convergent if $s^{(i)}$ is a prefix of $s(A,a)$ for all $i \geq k$ for some $k$.
+\begin{definition} We say that a sequence $s^{i}$ is weakly convergent to $s(C,a)$ if $s^\infty = s(C,a)$.
+We say that the sequence $s^{i}$ is strongly convergent if $s^{(i)}$ is a prefix of $s(C,a)$ for all $i \geq k$ for some $k$.
 \end{definition}
@@
 
-\begin{conjecture} A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 2)$ if the starting sequence begins with a $2$ and contains no more than 3 elements.
-A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 1)$ if the starting sequence begins with a $1$ and contains between 2 and 3 elements.
-A sequence $s^i$ is weakly convergent but not strongly convergent to $s(\{1,2\}, a)$ if the starting sequence begins with an $a \in \{1,2,\}$ and contains 4 or more elements.
+\begin{conjecture} Let $C = (1,2)$.
+A sequence $s^{(i)}$ is strongly convergent to $s(C, 2)$ if the starting sequence begins with a $2$ and contains no more than 3 elements.
+A sequence $s^{(i)}$ is strongly convergent to $s(C, 1)$ if the starting sequence begins with a $1$ and contains between 2 and 3 elements.
+A sequence $s^{(i)}$ is weakly convergent but not strongly convergent to $s(C, a)$ if the starting sequence begins with an $a \in \{1,2\}$ and contains 4 or more elements.
 \end{conjecture} 
@@
 \end{document}
->>>>>>> 94de583213b09230c51ca57ed5c9897bcabc4666

### 2014-11-04 draft.tex
Edits to the first 3 pages

@@
 
-The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). Thus, this sequence is \emph{self-referential}.
+The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). We call such a sequence \emph{self-referential}.
 
-Observe that in this case it is unambiguous which number to use next. After a block of 1's, the next block must be a block of 2's. However ambiguity exists in a sequence generated by $(1,2,3)$.
-If the first number is 1, then we can make the second number either a 2, or a 3. In particular, both:
+Observe that it was unambiguous which number a new block begins with.
+After a block of 1's, the next block must be a block of 2's and vice versa.
+However if our sequence was composed of $1$'s, $2'$s and $3'$s there would be ambiguity over which number a new block should begin with.
+For example if the first number is 1, then we could make the second number either a 2, or a 3. In particular, both:
 \begin{equation}
@@
 We would like to avoid having a large number of choices in the generation of the sequence, and so we will use the convention that the order in which the elements are presented gives the order they are used in the sequence.
-For example, $(1, 2, 3 )$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.
+For example, $(123 )$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.
 
-With this rule, the self-referencing sequence generated by $(1,2,3)$ that begins with a $1$ is
+With this rule, the self-referencing sequence generated by $(123)$ that begins with a $1$ is
 \begin{figure}
@@
 \includegraphics[width=0.5\linewidth]{sequenceBlocks123.pdf}
-\caption{Sequence generated by $(1,2,3)$ broken up into contiguous blocks.}
+\caption{Sequence generated by $(123)$ broken up into contiguous blocks.}
 \end{figure}
@@
 \item For some generating cycles there exists an equivalent formulation of self-referencing sequences using substitution rules
-\item If a generating cycle has substitution rules then one can compute the limiting density for the numbers in the sequencem if a limiting density exists.
-\item The limiting densities of the self-referencing sequences of  $(1,3)$ if such densities exists.
+\item If a generating cycle has substitution rules then one can compute the limiting density for the numbers in the sequence if a limiting density exists.
+\item The limiting densities of the self-referencing sequences of  $(13)$ if such densities exists.
 \end{itemize}
 
-In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined.
-We consider how the starting number for the sequence affects the long term behavior of the sequence.
-In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.
+In Section~\ref{determinism} we show that if we are given a generating cycle and a starting number for the sequence, the sequence is uniquely determined.
+In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating cycle, and not the starting number.
 
-In Section~\ref{iterativeprocess}, we take a different approach to analyzing these sequences by showing that there is an alternate way of generating them with an iterative process.
+In Section~\ref{iterativeprocess}, we show that there is an alternate way of generating them with an iterative process.
 In Section~\ref{substitutionrules}, we show how a large number of these sequences can actually be generated with a simple set of substitution rules.
-We introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating sets.
-Unfortunately, we find that the open problem of determining the density for the sequence of 1's and 2's is hard to solve using our methods.
+We introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating cycles.
+Unfortunately, we find that the open problem of determining the density of the sequences generated by $(12)$ is hard to solve using our methods.
 
@@
 
-We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number.
+We claim that a self-referencing sequence is uniquely determined by the generating cycle and the starting number.
 
@@
 
-In this section, we consider a different definition of equivalence.
-Consider, for example, sequences generated by $C = (1,2,3)$.
-Theorem~\ref{independentsequences} shows that the sequence starting with 2
+In this section, we consider long term equivalence of sequences. We introduce a definition for this
+\begin{definition}
+We say that two sequence $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n, k \geq 0$ such that 
+\begin{align*}
+a_{n+i} = b_{k+i}
+\end{align*}
+for all positive integers $i$. 
+\end{definition}
+
+ Consider, for example, sequences generated by $C = (123)$. Theorem~\ref{independentsequences} shows that the sequence starting with 2
 \begin{align*}
@@
 
-Thus, since both sequences are essentially the same if a prefix is removed from each of them, we call these sequences \emph{equivalent}.
-To formalize, this, we say that two sequences $(a_i)$ and $(b_i)$ are equivalent if there exists some integers $n,k \geq 0$ such that
-\begin{align*}
-a_{n+i} = b_{k+i}
-\end{align*}
-for all positive integers $i$. This definition is saying that the sequences are eventually equivalent.
-
-\begin{conjecture} Any sequence generated by a set $S$ is equivalent to any other sequence generated by the same set $S$.
+\begin{conjecture} Any sequence generated by a cycle $C$ is equivalent to any other sequence generated by the same cycle $C$.
 \end{conjecture}
 
-This conjecture is partly motivated by the fact that the property holds for self-referencing sequences from the generating cycle $C= (1,2,3)$.
+This conjecture is partly motivated by the fact that the property holds for self-referencing sequences from the generating cycle $C= (123)$.
 For generating cycles with larger numbers, we have been unable to verify or refute the conjecture..
-However, in Section~\ref{density}, we show that for some generating sets the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
+However, in Section~\ref{density}, we show that for some generating cycles the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
 

### 2014-11-04 draft.tex
Edited Theorem 4 of Equivalence

@@
 \begin{proof}
-Assume otherwise. So if the two sequences are $a_j$ and $b_j$ and we have 
+Assume otherwise. Suppose the two sequences are $\{a_j\}$ and $\{b_j\}$ and we have 
 \begin{equation*}
@@
 \end{equation*}
-such that if $m$ or $l$ were made smaller this relation would not hold.
-Suppose the read distance for $b_{1+k}$ is $x$ and the read distance for $a_{1+l}$ is $y$.
-Without loss of generality let us assume that $x>y$.
-Therefore we know there are two separate blocks of length $c_1$ in the sequence, one which corresponds to $a_{i+l}$ and another which corresponds to $b_{i+k}$.
-Because the two sequences are identical there exists some index $n$ such that $c_{n+1}=a_{m}$ is the element of sequence $a$ which corresponds to the second block of length $c_1$. 
-Because the sequences are self referential we know that $c_1, c_2, \ldots, c_{n}$  must appears in $b$ prior to element $i+k$.
-In fact the self referential nature of the sequence guarantees that $b_{i+k-1}=c_{n}$, that is this subsequence must end just before the two sequences become the same.
-Thus our sequence $b$ is
+for all $i>1$, such that if $m$ or $l$ were made smaller this relation would not hold.
+Suppose the read distance for $b_{k+1}$ is $x$ and the read distance for $a_{1+l}$ is $y$.
+Without loss of generality let us assume that $x>y$ (note that $x \neq y$ as per our assumption).
+
+Since the two read distances are different we know there are two separate blocks of length $c_1$ in the sequence, one which corresponds to $a_{i+l}$ and another which corresponds to $b_{i+k}$. 
+Because the two sequences are equal after a certain point and $\{a_j\}$ is self referential there exists some index $n$ such that $c_{n+1}=a_{n+l}$ is the element of $\{a_j\}$ which corresponds to the second block of length $c_1$.
+
+We now claim that $c_1, c_2, \ldots, c_{n}$  must appear in $\{b_j\}$ such that $b_{i+k-1}=c_{n}$.
+This is true because $\{b_j\}$ is self referential, and so the length of the blocks of the sequence $\{c_i\}$ must reproduce the terms of $\{b_j\}$.
+In particular the lengths of the blocks which occur before the second block of length $c_1$ must appear in $\{b_j\}$, and must end right next to the number which refers to the second block of length $c_1$.
+
+Thus our sequence $\{b_j\}$ is  
 \begin{equation*}
@@
 \end{equation*}
-where the prefix begins at the second $c_1$. We now claim that the common sequence cycles through
+where the prefix ends at the second $c_1$. We now claim that the common sequence cycles through
 \begin{equation*}
@@
 \end{equation*}
-Well by Lemma ~\ref{readdistformula} since the read distance for $b_{1+k}$ is greater than the read distance for $a_{1+l}$, and the sequences are identical for subsequent indices, we know that the read distance for $b_{i+k}$ is longer than the read distance for $a_{i+l}$ by exactly $x-y$.
+Since the read distance for $b_{1+k}$ is greater than the read distance for $a_{1+l}$, and the sequences are identical for subsequent indices by Lemma ~\ref{readdistformula}, we know that the read distance for $b_{i+k}$ is longer than the read distance for $a_{i+l}$ by exactly $x-y$.
 Thus $b_{k+i}$ corresponds to the same block as $a_{l+(x-y)+i}$.
 But $b_{k+i} = c_{i}$ and $a_{l+(x-y)+i}=c_{x-y+i}$ and so $c_{i} = c_{x-y+i}$ and so the common sequence cycles.
-Because the common sequence cycles and $b$ has the same cycle, but one which begins before the common sequence $l$ could be made smaller and have the equivalence relation hold.
-This is a contradiction and so the first element of the shared sequence must have the same read distance in both sequences.
+Because the common sequence cycles and $\{b_j\}$ has the same cycle, but one which begins before the common sequence, $l$ could be made smaller and have the equivalence relation hold.
+This is contradicts our statement that $l$ is as small as possible.
+Therefore the first element of the shared sequence must have the same read distance in both sequences.
 \end{proof}

### 2014-11-04 draft.tex
Finished editing Equivalence

@@
 We also have a sufficient condition for equivalence 
-\begin{lemma} If two sequences $(a_i), (b_i)$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are the same and 
+\begin{proposition} 
+\label{equivsuff}
+If two sequences $(a_i), (b_i)$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
 \begin{equation*}
@@
 \end{equation*}
-where $k$ ranges from 0 to the read distance for $a_m$ then $(a_i)$ and $(b_i)$ are equivalent. 
-\end{lemma}
+where $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
+\end{proposition}
 \begin{proof}
-We will prove this using a simple induction.
-Suppose the read distance for $a_m$ is $d$ then know that $a_{m+d} = b_{l+d}$ because $a_{m+d-1} = b_{l+d-1}$ and $a_{m+d}$ is the beginning of a new block.
-Furthermore we have that the read distance for $a_{m+d}$ equals the read distance for $b_{l+d}$.
-Now suppose that we have for $k \leq j$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$ and we would like to show that this holds for $k = j+1$.
-By Lemma ~\ref{readdistformula} we actually already know that the read distance for the two is the same, as it is the same for the preceding terms, which are equal.
+We will prove this by  induction.
+In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$.
+By our assumptions we already have this for $0 \leq k \leq d-1$ and so we begin with $k=d$.
+
+Since a new block begins at $a_{m+d}$ and $b_{l+d}$ and $a_{m+d-1} = b_{l+d-1}$ we have $a_{m+d} = b_{l+d}$.
+Furthermore, by Proposition ~\ref{readdistformula} we have that the read distance for $a_{m+d}$ equals the read distance for $b_{l+d}$.
+
+Now suppose that  for $k \leq j$ we have $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$.
+We would like to show that this holds for $k = j+1$.
+By Proposition ~\ref{readdistformula} we actually already know that the read distance for the two is the same, as it is the same for the preceding terms, which are equal.
 
@@
 
-Otherwise $a_{m+k}$ and $a_{m+k+1}$ are in the same block, then we only need to show that $b_{l+k}$ and $b_{l+k+1}$ are in the same block as well.
-But as we argued above the length of blocks is determined by terms earlier in the sequence which we know to be equal and have the same read distance.
+Otherwise $a_{m+k}$ and $a_{m+k+1}$ are in the same block, and so we only need to show that $b_{l+k}$ and $b_{l+k+1}$ are in the same block as well.
+But as we argued above the length of blocks is determined by terms earlier in the sequence, which we know to be equal and have the same read distance.
 Therefore we have $a_{m+k+1} = b_{l+k+1}$ for this case as well.
-This completes our induction and so $(a_i)$ and $(b_i)$ are equivalent.
+This completes our induction and so $\{a_i\}$ and $\{b_i\}$ are equivalent.
 \end{proof}
@@
 
-\begin{corollary} Let $C = (1, c_1, c_2, \ldots, c_n)$. $s(C, 1)$ and $s(C, c_1)$ are equivalent.
+\begin{corollary} Let $C = (1 c_1 c_2 \cdots c_n)$. $s(C, 1)$ and $s(C, c_1)$ are equivalent.
 \end{corollary}
 \begin{proof}
-Let $(a_i) = s(C, 1)$ and $(b_i) = s(C, c_1)$.
-The read distance of sequence $(a_i)$ for $i=2$ is 0, and the read distance of sequence $(b_i)$ for $i = 1$ is 0.
-In addition, $a_2 = b_1 = c_1$, which means the two sequences are equivalent.
+Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$.
+The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0.
+In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
 \end{proof}

### 2014-11-05 draft.tex
changed sequences to curly braces

@@
 \begin{lemma}
-Let $C$ be a generating cycle. Suppose $(a_i)$ and $(b_i)$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
+Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
 \end{lemma}
@@
 \begin{proof}
-Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $(a_i)$ and $(b_i)$ are different.
-Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $(a_i)$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $(b_i)$, we get that $b_j$ comes after $a_{j-1}$.
+Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different.
+Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$.
 This is a contradiction since $a_j \neq b_j$.
@@
 We now introduce a notion which will help us categorize sequences as equivalent. 
-\begin{definition} Consider a sequence $(a_i)$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
+\begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
 \end{definition}
-The read distance can be any nonnegative integer. For example if we consider the sequence $(a_i)$ generated by $(1,2,3)$ beginning with a 1  
+The read distance can be any nonnegative integer. For example if we consider the sequence $\{a_i\}$ generated by $(1,2,3)$ beginning with a 1  
 \begin{align*}
@@
 \label{readdistformula}
-Let $(a_i)$ be any self referencing sequence generated by some set of positive integers. Then the read distance for $i>1$, is 
+Let $\{a_i\}$ be any self referencing sequence generated by some set of positive integers. Then the read distance for $i>1$, is 
 \begin{equation*}
@@
 
-\begin{lemma} Consider a self referencing sequence $(a_i)$ generated by $(c_1, c_2, \ldots, c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$.
+\begin{lemma} Consider a self referencing sequence $\{a_i\}$ generated by $(c_1, c_2, \ldots, c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$.
 Suppose the read distance for $i$ is $d$, then the read distance for $i+m$ is at least $d+1$.
@@
 \label{equivsuff}
-If two sequences $(a_i), (b_i)$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
+If two sequences $\{a_i\}, \{b_i\}$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
 \begin{equation*}
@@
 
-The following is a simple observation of sequences that begin with $1$.
-
-\begin{corollary} Let $C = (1 c_1 c_2 \cdots c_n)$. $s(C, 1)$ and $s(C, c_1)$ are equivalent.
+This result allows us to make a general statement for equivalence for sequences that begin with a $1$.
+\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
 \end{corollary}
@@
 
-\begin{conjecture} If two sequences $(a_i)$ and $(b_i)$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
+\begin{conjecture} If two sequences $\{a_i\}$ and $\{b_i\}$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
 \end{conjecture}

### 2014-11-05 draft.tex
made conjecture boldface by using begin conj and end conj

@@
 
-\makeatletter
-\let\c@lemma=\c@theorem
-\let\c@corollary=\c@theorem
-\let\c@fact=\c@theorem
-\makeatother
-
-\let\realendproof=\endproof
-\def\endproof{\hspace*{\fill}$\Box$\realendproof}
+%\makeatletter
+%\let\c@lemma=\c@theorem
+%\let\c@corollary=\c@theorem
+%\let\c@fact=\c@theorem
+%\makeatother
+%
+%\let\realendproof=\endproof
+%\def\endproof{\hspace*{\fill}$\Box$\realendproof}
+\newtheorem{conj}[theorem]{Conjecture} 
 \title{Self Referencing Sequences}
@@
 
-\begin{conjecture} Any sequence generated by a cycle $C$ is equivalent to any other sequence generated by the same cycle $C$.
-\end{conjecture}
+\begin{conj} Any sequence generated by a cycle $C$ is equivalent to any other sequence generated by the same cycle $C$.
+\end{conj}
 
@@
 
-\begin{conjecture} If two sequences $\{a_i\}$ and $\{b_i\}$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
-\end{conjecture}
+\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
+\end{conj}
 
@@
 
-\begin{conjecture} Let $C = (1,2)$.
+\begin{conj} Let $C = (1,2)$.
 A sequence $s^{(i)}$ is strongly convergent to $s(C, 2)$ if the starting sequence begins with a $2$ and contains no more than 3 elements.
@@
 A sequence $s^{(i)}$ is weakly convergent but not strongly convergent to $s(C, a)$ if the starting sequence begins with an $a \in \{1,2\}$ and contains 4 or more elements.
-\end{conjecture} 
+\end{conj} 
 We make this conjecture on the basis on computer simulations which executes the iterative process to at least 20 iterations.

### 2014-11-05 draft.tex
changed generating sets to generating cycles, made notation consistent

@@
 We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element.
-Assume the generating cycle is $(a_1,a_2,\dots,a_m)$ and that the starting element is some $a_i$.
+Assume the generating cycle is $(a_1 a_2 \dots a_m)$ and that the starting element is some $a_i$.
 For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
@@
 \end{definition}
-The read distance can be any nonnegative integer. For example if we consider the sequence $\{a_i\}$ generated by $(1,2,3)$ beginning with a 1  
+The read distance can be any nonnegative integer. For example if we consider the sequence $\{a_i\}$ generated by $C = (1 2 3)$ beginning with a 1  
 \begin{align*}
@@
 
-\begin{lemma} Consider a self referencing sequence $\{a_i\}$ generated by $(c_1, c_2, \ldots, c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$.
+\begin{lemma} Consider a self referencing sequence $\{a_i\}$ generated by $(c_1 c_2 \ldots c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$.
 Suppose the read distance for $i$ is $d$, then the read distance for $i+m$ is at least $d+1$.
@@
 
-We change focus in this section and consider how to generate self-referential sequences with an iterative process. We first look at the sequences generated by the cycle $C = (1,2)$.
+We change focus in this section and consider how to generate self-referential sequences with an iterative process. We first look at the sequences generated by the cycle $C = (1 2)$.
 
@@
 It turns out that we can be less restrictive about what our starting sequence, $s^{(0)}$, is.
-Going back to the case of $C = (1,2)$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2.
+Going back to the case of $C = (1 2)$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2.
 And still, the limiting sequence will be the self-referential sequence.
@@
 The iterative process given in Section~\ref{iterativeprocess} is one way to generate self-referencing sequences.
-However, for certain generating sets $A$, there is a simpler ways to perform this iteration using a fixed set of substitution rules.
+However, for certain generating cycles $A$, there is a simpler ways to perform this iteration using a fixed set of substitution rules.
 Take $A = \{1,3\}$ as an example, and consider the sequence that starts with a 3:
@@
 
-We would like to know if we can do this for any generating set $G$ and starting number $g$.
+We would like to know if we can do this for any generating cycle $G$ and starting number $g$.
 In particular we would like to be able to simulate our original algorithm with a substitution rule.
@@
 That is each rule should write complete blocks.
-In order to comply with the third condition it is enough for the size of the generating set to divide the length of the rules.
+In order to comply with the third condition it is enough for the size of the generating cycle to divide the length of the rules.
 This is so that at the start of each rule, we always know which numbers to write.
 
-Therefore, we can formulate a set of sufficient conditions to make rules to generate the self-referential sequence with generating set $G$.
-We want to have a set of rules $R$ where each rule $r \in R$ has length $|r| = l$, such that all $|G|^{l}$ strings are in the rule set and for each $r \in R$
+Therefore, we can formulate a set of sufficient conditions to make rules to generate the self-referential sequence with generating cycle $C$.
+We want to have a set of rules $R$ where each rule $r \in R$ has length $|r| = l$, such that all $|C|^{l}$ strings are in the rule set and for each $r \in R$
 \begin{itemize}
-\item each rule writes blocks in the same order of $G$
+\item each rule writes blocks in the same order of $C$
 \item $\sum_{i \in r} i$ is divisible by $l$
-\item $l$ is divisible by $|G|$
+\item $l$ is divisible by $|C|$
 \end{itemize}
@@
 
-Another example, we can have $G = \{ 2, 4 \}$, and we can have that the rules be
+Another example, we can have $C = (2 4)$, and we can have that the rules be
 \begin{align*}
@@
 
-One of the reasons why generating a rule set for $\{1, 2\}$ is complicated is that since $1, ..., 1$ must be in the rule set, the length of the rules must be even (since $|G| = |\{1, 2\}| = 2$).
+One of the reasons why generating a rule set for $(1 2)$ is complicated is that since $1, ..., 1$ must be in the rule set, the length of the rules must be even (since $|C| = 2$).
 On the other hand, $2, 1, ..., 1$ must also be in the set and so the rule length must be odd.
@@
 
-A topic of interest when considering a self referencing sequence is the relative density of the various elements of the generating set.
-It is unclear if such a density even exists and if so how to compute it for arbitrary generating sets.
+A topic of interest when considering a self referencing sequence is the relative density of the various elements of the generating cycle.
+It is unclear if such a density even exists and if so how to compute it for arbitrary generating cycles.
 In this section, we show how to use substitution rules to compute relative densities for certain sequences if we assume such densities exists.
-We explain the method and compute the relative density of 3's for $G = \{ 1, 3\}$.
-We will also produce bounds on the density of $2$'s $G = \{ 1, 2\}$ in a way that suggests that the density converges to 2; however, this method is computationally difficult for smaller bounds since the number of cases to check increases exponentially.
+We explain the method and compute the relative density of 3's for $C = (1 3)$.
+We will also produce bounds on the density of $2$'s $C = (1 2)$ in a way that suggests that the density converges to 2; however, this method is computationally difficult for smaller bounds since the number of cases to check increases exponentially.
 
-To begin we will discuss a technique that can be used to compute the densities for some generating sets.
-In particular, suppose we have a generating set $G$ with a set of rules $R$ that satisfy the conditions from the section above.
+To begin we will discuss a technique that can be used to compute the densities for some generating cycles.
+In particular, suppose we have a generating cycle $C$ with a set of rules $R$ that satisfy the conditions from the section above.
 For each rule $r \in R$, we will say denote the rule as $r_l \rightarrow r_r$ as the inputs and the outputs.
@@
 
-As discussed in the previous section, it is unclear how to establish a set of rules for $G = \{1 , 2\}$; however, there is some structure that we can exploit.
+As discussed in the previous section, it is unclear how to establish a set of rules for $C = (1 2)$; however, there is some structure that we can exploit.
 In particular, we know that the sequences $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence.
@@
 \item In Section~\ref{equivalence}, we propose the idea that in some sense, the number that we start a sequence with does not matter.
-In other words, any sequence that uses a certain generating set is equivalent to any other sequence created by the same generating set.
+In other words, any sequence that uses a certain generating cycle is equivalent to any other sequence created by the same generating cycle.
 \item In the same section, we propose a condition for determining when two sequences are not equivalent.
@@
 Most of what we know about them comes from analyzing numerical data.
-\item In Section~\ref{substitutionrules}, we look at substitution rules for certain ``nice" generating sets of numbers, such as $\{1,3\}$ and $\{2,4\}$.
+\item In Section~\ref{substitutionrules}, we look at substitution rules for certain ``nice" generating cycles of numbers, such as $(13)$ and $(24)$.
 These sets allow for substitution rules because they satisfy the trio of properties we outline in that section.
@@
 We leave unresolved the question of determining precisely which sets allow for substitution rules and which ones ones do not.
-The major difficulty here seems to come in proving that a certain generating set does \emph{not} allow for substitution rules.
-\item In Section~\ref{density} we were able to compute the density of any sequence with generating set $\{1,3\}$, and we were able to give bounds on the density of any sequence with generating set $\{1,2\}$.
-However, we leave unresolved the question of what the density for $\{1,2\}$ actually is, or even if we can obtain tighter bounds.
+The major difficulty here seems to come in proving that a certain generating cycle does \emph{not} allow for substitution rules.
+\item In Section~\ref{density} we were able to compute the density of any sequence with generating cycle $(13)$, and we were able to give bounds on the density of any sequence with generating cycle $(12)$.
+However, we leave unresolved the question of what the density for $(12)$ actually is, or even if we can obtain tighter bounds.
 \end{itemize}

### 2014-11-05 draft.tex
Restructuring Equivalence section

@@
 
-In this section, we consider long term equivalence of sequences. We introduce a definition for this
+In this section, we consider long term equivalence of sequences. We introduce a formal definition of this and establish results that allow us to determine when two sequences are equivalent.
+
+We would first like to make a formal definition of equivalence. 
 \begin{definition}
@@
 
- Consider, for example, sequences generated by $C = (123)$. Theorem~\ref{independentsequences} shows that the sequence starting with 2
+As an example let us consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
 \begin{align*}
-2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
 \end{align*}
-and the sequence starting with 3
+and the sequence starting with 3 is 
 \begin{align*}
-3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ .
 \end{align*}
-are independent in the sense that neither sequence is exactly contained in the other.
-However, if we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we get
+If we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we see 
 \begin{align*}
-2,2,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
-3,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+2,2,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
+3,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
 \end{align*}
-in both cases.
+that the resultant sequences appear to be identical and thus equivalent.
+At an intuitive level it seems obvious that these two truncated sequences are identical as they both begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers.
 
-\begin{conj} Any sequence generated by a cycle $C$ is equivalent to any other sequence generated by the same cycle $C$.
-\end{conj}
-
-This conjecture is partly motivated by the fact that the property holds for self-referencing sequences from the generating cycle $C= (123)$.
-For generating cycles with larger numbers, we have been unable to verify or refute the conjecture..
-However, in Section~\ref{density}, we show that for some generating cycles the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
-
-We now introduce a notion which will help us categorize sequences as equivalent. 
+In order to formalize and generalize this idea we introduce the notion of a read distance.  
 \begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
@@
 We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0, 0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.
-The read distance, can be thought of as the horizontal shift of the arrows in Figure~\ref{fig:onetwothree}.
 
+This monotonicity is true in general, and we can in fact make a stronger statement on the value of the read distance at a given index.
 \begin{lemma} 
@@
 Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing.
-Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a simple bound on the number of consecutive times the read distance can stay constant.
 
-\begin{lemma} Consider a self referencing sequence $\{a_i\}$ generated by $(c_1 c_2 \ldots c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$.
+%Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a simple bound on the number of consecutive times the read distance can stay constant.
+%\begin{corollary} Consider a self referencing sequence $\{a_i\}$ generated by $(c_1 c_2 \ldots c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$.
 Suppose the read distance for $i$ is $d$, then the read distance for $i+m$ is at least $d+1$.
-\end{lemma}
+%\end{corollary}
+%\begin{proof}
+%The read distances can stay constant from $i$ to $i+1$ only if there is a $a_{i+1} = 1$.
+There can be at most $m$ 1's in a contiguous block, so the read distance of $i+m$ is bigger than the read distance of $i$.
+%\end{proof}
+
+Now that we have a nice expression for the value of the read distance at a particular index we can show our first major result which allows us to check only finitely many indices in order to establish that two sequences are equivalent.
+\begin{theorem} 
+\label{equivsuff}
+If two sequences $\{a_i\}, \{b_i\}$ generated over the same generating cycle are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
+\begin{equation*}
+a_{m+k} = b_{l+k}
+\end{equation*}
+where $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
+\end{theorem}
 \begin{proof}
-The read distances can stay constant from $i$ to $i+1$ only if there is a $a_{i+1} = 1$.
-There can be at most $m$ 1's in a contiguous block, so the read distance of $i+m$ is bigger than the read distance of $i$.
+We will prove this by  induction.
+In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$.
+By our assumptions we already have this for $0 \leq k \leq d-1$ and so we begin with $k=d$.
+
+Since a new block begins at $a_{m+d}$ and $b_{l+d}$ and $a_{m+d-1} = b_{l+d-1}$ we have $a_{m+d} = b_{l+d}$.
+Furthermore, by Proposition ~\ref{readdistformula} we have that the read distance for $a_{m+d}$ equals the read distance for $b_{l+d}$.
+
+Now suppose that  for $k \leq j$ we have $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$.
+We would like to show that this holds for $k = j+1$.
+By Proposition ~\ref{readdistformula} we actually already know that the read distance for the two is the same, as it is the same for the preceding terms, which are equal.
+
+If $a_{m+k}$ is the last term in a block then $b_{m+k}$ must also be the last term in its block, because the length of blocks is determined by earlier terms which we know to be equal and to have the same read distance.
+Since $a_{m+k} = b_{l+k}$  in this case we have $a_{m+k+1} = b_{l+k+1}$.
+
+Otherwise $a_{m+k}$ and $a_{m+k+1}$ are in the same block, and so we only need to show that $b_{l+k}$ and $b_{l+k+1}$ are in the same block as well.
+But as we argued above the length of blocks is determined by terms earlier in the sequence, which we know to be equal and have the same read distance.
+Therefore we have $a_{m+k+1} = b_{l+k+1}$ for this case as well.
+This completes our induction and so $\{a_i\}$ and $\{b_i\}$ are equivalent.
 \end{proof}
 
+This result also allows us to make a general statement for equivalence for sequences that begin with a $1$.
+\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
+\end{corollary}
+\begin{proof}
+Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$.
+The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0.
+In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
+\end{proof}
+
+We also can demonstrate a sufficient condition for equivalence. 
 \begin{theorem} If two sequences are equivalent then the first element of the shared sequence must have the same read distance in both sequences.
@@
 
-We also have a sufficient condition for equivalence 
-\begin{proposition} 
-\label{equivsuff}
-If two sequences $\{a_i\}, \{b_i\}$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
-\begin{equation*}
-a_{m+k} = b_{l+k}
-\end{equation*}
-where $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
-\end{proposition}
-\begin{proof}
-We will prove this by  induction.
-In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$.
-By our assumptions we already have this for $0 \leq k \leq d-1$ and so we begin with $k=d$.
-
-Since a new block begins at $a_{m+d}$ and $b_{l+d}$ and $a_{m+d-1} = b_{l+d-1}$ we have $a_{m+d} = b_{l+d}$.
-Furthermore, by Proposition ~\ref{readdistformula} we have that the read distance for $a_{m+d}$ equals the read distance for $b_{l+d}$.
-
-Now suppose that  for $k \leq j$ we have $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$.
-We would like to show that this holds for $k = j+1$.
-By Proposition ~\ref{readdistformula} we actually already know that the read distance for the two is the same, as it is the same for the preceding terms, which are equal.
-
-If $a_{m+k}$ is the last term in a block then $b_{m+k}$ must also be the last term in its block, because the length of blocks is determined by earlier terms which we know to be equal and to have the same read distance.
-Since $a_{m+k} = b_{l+k}$  in this case we have $a_{m+k+1} = b_{l+k+1}$.
-
-Otherwise $a_{m+k}$ and $a_{m+k+1}$ are in the same block, and so we only need to show that $b_{l+k}$ and $b_{l+k+1}$ are in the same block as well.
-But as we argued above the length of blocks is determined by terms earlier in the sequence, which we know to be equal and have the same read distance.
-Therefore we have $a_{m+k+1} = b_{l+k+1}$ for this case as well.
-This completes our induction and so $\{a_i\}$ and $\{b_i\}$ are equivalent.
-\end{proof}
-
-This result allows us to make a general statement for equivalence for sequences that begin with a $1$.
-\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
-\end{corollary}
-\begin{proof}
-Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$.
-The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0.
-In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
-\end{proof}
-
 These conditions for equivalence lead us to make the following conjecture on a sufficient condition for nonequivalence.
@@
 \end{conj}
+
+\begin{conj} Any sequence generated by a cycle $C$ is equivalent to any other sequence generated by the same cycle $C$.
+\end{conj}
+
+This conjecture is partly motivated by the fact that the property holds for self-referencing sequences from the generating cycle $C= (123)$.
+For generating cycles with larger numbers, we have been unable to verify or refute the conjecture..
+However, in Section~\ref{density}, we show that for some generating cycles the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
 

### 2014-11-05 draft.tex
updating the determinism section

@@
 
-We claim that a self-referencing sequence is uniquely determined by the generating cycle and the starting number.
+We claim that a self-referencing sequence is uniquely determined by the generating cycle $C$ and the starting number $a \in C$.
+We will prove this by showing that if we are given the first number of the sequence, there is only one way to extend the sequence.
+In order to do this, we formalize the idea of breaking up a sequence into blocks.
 
-\begin{lemma}
-Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
-\end{lemma}
+\begin{definition}
+A block $B$ is any maximal length subsequence of repeated numbers in a sequence. We denote by $B_i$ the $i$th block of the sequence.
+\end{definition}
 
+Here, \emph{maximal length} is a local property that means that the the block cannot be extended to include the numbers preceding or succeeding the block, as they will not be the repeated number.
+Thus, if a sequence $\{s_i\}$ is
+\begin{align*}
+1,3,3,3,1,1,1,3\dots \,
+\end{align*}
+the first block is $B_1 = \{1\}$, the second is $B_2 = \{3,3,3\}$, the third is $B_3 = \{1,1,1\}$, and so on.
+
+It is obvious that a sequence can be specified by giving the numbers of the sequence individually, or by specifying all the blocks $B_i$. This is enough to make the following claim.
+
+\begin{proposition}
+A self-referencing sequence $s$ is uniquely specified by the generating cycle $C$, and the starting number $a \in C$. In other words, such an $s$ is guaranteed to exist and is unique.
+\end{proposition}
 \begin{proof} 
-Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different.
+The proof is straightforward. The contents of block $B_i$ are determined by the value of $i$ and $s_i$. The value of $i$ determines which number is repeated throughout the block. 
+\end{proof}
+%
+%\begin{lemma}
+%Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
+%\end{lemma}
+%
+%\begin{proof}
+%Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different.
 Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$.
@@
 Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.
-
-\begin{figure}{hb}
-\label{fig:unequalsequences}
-\center
-\includegraphics[width=0.5\linewidth]{lemma1.pdf}
-\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
-\end{figure}
-
-This implies that every block before the $k$th block just contains 1's since each block refered it itself.
+%
+%\begin{figure}{hb}
+%\label{fig:unequalsequences}
+%\center
+%\includegraphics[width=0.5\linewidth]{lemma1.pdf}
+%\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
+%\end{figure}
+%
+%This implies that every block before the $k$th block just contains 1's since each block refered it itself.
 This means that either $j = 2$, or the sequence is not self-referential.
 But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle.
-\end{proof}
-
-We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element.
+%\end{proof}
+%
+%
+%
+%We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element.
 Assume the generating cycle is $(a_1 a_2 \dots a_m)$ and that the starting element is some $a_i$.
 For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
-\begin{quote}
-\textbf{Algorithm:}\\
-Suppose the first number of the sequence is $a_i$.
+%\begin{quote}
+%\textbf{Algorithm:}\\
+%Suppose the first number of the sequence is $a_i$.
 Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block.
@@
 If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
-
-This procedure continues deterministically so long as there is an element to read.
+%
+%This procedure continues deterministically so long as there is an element to read.
 Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
-\end{quote}
+%\end{quote}
 

### 2014-11-05 draft.tex
finished the determinism section

@@
 
-Here, \emph{maximal length} is a local property that means that the the block cannot be extended to include the numbers preceding or succeeding the block, as they will not be the repeated number.
-Thus, if a sequence $\{s_i\}$ is
+Here, \emph{maximal length} is a local property that means that the block cannot be extended to include the numbers preceding or succeeding the block, as they will not be the repeated number.
+Thus, if a sequence $\{a_i\}$ is
 \begin{align*}
@@
 
-It is obvious that a sequence can be specified by giving the numbers of the sequence individually, or by specifying all the blocks $B_i$. This is enough to make the following claim.
+Clearly, a sequence can be specified by giving the numbers of the sequence individually, or by specifying all the blocks $B_i$. This is enough to make the following claim:
 
 \begin{proposition}
-A self-referencing sequence $s$ is uniquely specified by the generating cycle $C$, and the starting number $a \in C$. In other words, such an $s$ is guaranteed to exist and is unique.
+A self-referencing sequence $\{a_i\}$ is uniquely specified by the generating cycle $C$ and the starting number $a \in C$. In other words, such an $\{a_i\}$ is guaranteed to exist and is unique.
 \end{proposition}
 \begin{proof} 
-The proof is straightforward. The contents of block $B_i$ are determined by the value of $i$ and $s_i$. The value of $i$ determines which number is repeated throughout the block. 
+The proof is constructive.
+The self-referential property of the sequence implies that the contents of block $B_i$ are entirely determined by the values of $a$, $i$, and $s_i$.
+The value of $i$ determines which number is repeated throughout the block.
+For example, when $i = 1$, the number is the starting number $a$, and when $i = 2$, we use the number following the starting number in $C$, and so on for larger values of $i$.
+The value of $s_i$ simply dictates the length of block $B_i$.
+
+Thus, the sequence is uniquely given by first ``writing down" $B_1$, followed by $B_2$, followed by $B_3$, etc.
+One detail is that it remains to check that $s_i$ is known when $B_i$ is written down.
+However, writing down a block always involves writing down at least one element of the sequence, so the value of $s_i$ is always known when $B_i$ is written down.
+Note that there is an edge case where $s_i$ is not yet written down if the starting number is 1 and $i =2$.
+However, in this case, the value of $s_i$ is still known (as it is given by the rule) and thus $B_i$ can still be written down.
 \end{proof}

### 2014-11-05 draft.tex
went through all formal statements

@@
 
+We will study a sequences which are self-referencing.
+That is, each number in the sequence will correspond to a previous number in the same sequence.
+In particular, we will analyze self-referencing sequences with respect to block lengths.
 Consider the sequence
@@
 
-Observe that it was unambiguous which number a new block begins with.
-After a block of 1's, the next block must be a block of 2's and vice versa.
+Above was an example where the blocks always have elements be either $1$ or $2$.
+Since we are self-referincing with respect to block lengths, the particular value that the block takes is not as important as the block lengths.
+Whenever the sequence can only takes two possible numbers, the values of the blocks is unambiguous; the block values simply alternate.
+As in the sequence above, after a block of 1's, the next block must be a block of 2's and vice versa.
 However if our sequence was composed of $1$'s, $2'$s and $3'$s there would be ambiguity over which number a new block should begin with.
+
 For example if the first number is 1, then we could make the second number either a 2, or a 3. In particular, both:
@@
 \end{equation}
-are in some sense ``self-referencing'' sequences.
+are in some sense self-referencing sequences.
 Each time we need to specify a new block, we have a choice of two elements.
@@
 
-With this rule, the self-referencing sequence generated by $(123)$ that begins with a $1$ is
+With this rule, the self-referencing sequence generated by $(123)$ that begins with a $1$ is shown in Figure~\ref{fig:onetwothree}.
 \begin{figure}
@@
 \item The self-referencing sequence for a given generating cycle is uniquely determined by its first number.
-\item Some characterizations of sequences with the same generating cycle.
+% This doesn't really mean anything... - erik
+%
+%\item Some characterizations of sequences with the same generating cycle.
 \item For some generating cycles there exists an equivalent formulation of self-referencing sequences using substitution rules
@@
 \label{readdistformula}
-Let $\{a_i\}$ be any self referencing sequence generated by some set of positive integers. Then the read distance for $i>1$, is 
+Let $\{a_i\}$ be any self referencing sequence generated by some cycle of positive integers. Then the read distance for $i>1$, is 
 \begin{equation*}
@@
 \label{equivsuff}
-If two sequences $\{a_i\}, \{b_i\}$ generated over the same generating cycle are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
+Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle.If there are integers $m,l$ with the read distance for $a_m$ and $b_l$ are both $d$ and 
 \begin{equation*}
@@
 \end{theorem}
+
 \begin{proof}
@@
 We also can demonstrate a sufficient condition for equivalence. 
-\begin{theorem} If two sequences are equivalent then the first element of the shared sequence must have the same read distance in both sequences.
+\begin{theorem}
+\label{thm:readfirst}
+If two sequences are equivalent then the first element of the shared sequence must have the same read distance in both sequences.
 \end{theorem}
+
 \begin{proof}
@@
 
+\begin{corollary}
+If two self-referencing sequences are equivalent, then their read distance of the shared subsequence is the same everywhere.
+\end{corollary}
+
+\begin{proof}
+This follows from Theorem~\ref{thm:readfirst} and Lemma~\ref{readdistformula}.
+\end{proof}
+
 These conditions for equivalence lead us to make the following conjecture on a sufficient condition for nonequivalence.
 
-\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
+\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent, then whenever two elements have the same read distances, the two elements are equal. 
 \end{conj}
@@
 Let $C$ be some generating cycle.
-Let $s^0 = a$ for some $a \in C$ where $a \neq 1$, and let $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$.
-Then the limit of this process, $s^\infty$, is the sequence $s(C,a)$.
+Let $s^0 = a$ for some $a \in C$ where $a \neq 1$, and let $s^{(i)}$ be the sequences generated by the iterative process.
+Then the limit of this process $s^\infty$ exists and is the sequence $s(C,a)$.
 \end{theorem}
@@
 The iterative process given in Section~\ref{iterativeprocess} is one way to generate self-referencing sequences.
-However, for certain generating cycles $A$, there is a simpler ways to perform this iteration using a fixed set of substitution rules.
-Take $A = \{1,3\}$ as an example, and consider the sequence that starts with a 3:
+However, for certain generating cycles $C$, there is a simpler ways to perform this iteration using a fixed set of substitution rules.
+Take $C = (1 3)$ as an example, and consider the sequence that starts with a 3:
 \begin{align*}
@@
 
-We would like to know if we can do this for any generating cycle $G$ and starting number $g$.
-In particular we would like to be able to simulate our original algorithm with a substitution rule.
-Previously we established that our algorithm had a state corresponding to
-\begin{itemize}
-\item current block size
-\item number of elements already written down in the current block
-\item what number the current block is composed of 
-\end{itemize}
+We would like to know if we can do this for any generating cycle $C$ and starting number $a$. In particular we would like to be able to simulate our iterative process with substitution rules.
 
@@
 Therefore, we can formulate a set of sufficient conditions to make rules to generate the self-referential sequence with generating cycle $C$.
+
 We want to have a set of rules $R$ where each rule $r \in R$ has length $|r| = l$, such that all $|C|^{l}$ strings are in the rule set and for each $r \in R$
 \begin{itemize}
-\item each rule writes blocks in the same order of $C$
+\item each rule writes blocks in the same order of $C$ and each rule has the same starting number
 \item $\sum_{i \in r} i$ is divisible by $l$

### 2014-11-05 draft.tex
added another poem

@@
 (and the whole paper).
+
+\subsection{Poem}
+
+\noindent A sequence refering to its past\\
+warrants many questions to be asked.\\
+These sequences are deterministic \\
+and their read distances an imporant statistic.
+\\
+
+\noindent We want to compute the limiting density\\
+so we've defined a process.\\
+But one would be a fool to not follow our rule.\\
+And compute the density for 1, 3.\\
+
+\noindent The big question remains unanswered.\\
+The density of 1,2 which seems to be one half.\\
+Our rules are for numbers of a certain class.\\
+Unfortunetly, dealing with 1,2 is a pain in the a**.\\
+
 \end{document}

### 2014-11-06 draft.tex
Adding New Figures
Edit Single Typo in Determinism
Otherwise Determinism is Done

@@
 \center
-\includegraphics[width=0.5\linewidth]{sequenceBlocks.pdf}
+\includegraphics[width=0.4\linewidth]{sequenceBlocks}
 \caption{Sequence broken up into contiguous blocks with lengths generating same sequence.}
@@
 \center
-\includegraphics[width=0.5\linewidth]{sequenceBlocks123.pdf}
+\includegraphics[width=0.4\linewidth]{sequenceBlocks123}
 \caption{Sequence generated by $(123)$ broken up into contiguous blocks.}
@@
 
-We claim that a self-referencing sequence is uniquely determined by the generating cycle $C$ and the starting number $a \in C$.
+We claim that a self-referencing sequence is uniquely determined by the generating cycle $C$ and the starting number $c \in C$.
 We will prove this by showing that if we are given the first number of the sequence, there is only one way to extend the sequence.
@@
 \begin{proposition}
-A self-referencing sequence $\{a_i\}$ is uniquely specified by the generating cycle $C$ and the starting number $a \in C$. In other words, such an $\{a_i\}$ is guaranteed to exist and is unique.
+A self-referencing sequence $\{a_i\}$ is uniquely specified by the generating cycle $C$ and the starting number $c \in C$. In other words, such an $\{a_i\}$ is guaranteed to exist and is unique.
 \end{proposition}
@@
 The proof is constructive.
-The self-referential property of the sequence implies that the contents of block $B_i$ are entirely determined by the values of $a$, $i$, and $s_i$.
+The self-referential property of the sequence implies that the contents of block $B_i$ are entirely determined by the values of $c$, $i$, and $a_i$.
 The value of $i$ determines which number is repeated throughout the block.
-For example, when $i = 1$, the number is the starting number $a$, and when $i = 2$, we use the number following the starting number in $C$, and so on for larger values of $i$.
-The value of $s_i$ simply dictates the length of block $B_i$.
+For example, when $i = 1$, the number is the starting number $c$, and when $i = 2$, we use the number following the starting number in $C$, and so on for larger values of $i$.
+The value of $a_i$ simply dictates the length of block $B_i$.
 
 Thus, the sequence is uniquely given by first ``writing down" $B_1$, followed by $B_2$, followed by $B_3$, etc.
-One detail is that it remains to check that $s_i$ is known when $B_i$ is written down.
-However, writing down a block always involves writing down at least one element of the sequence, so the value of $s_i$ is always known when $B_i$ is written down.
-Note that there is an edge case where $s_i$ is not yet written down if the starting number is 1 and $i =2$.
-However, in this case, the value of $s_i$ is still known (as it is given by the rule) and thus $B_i$ can still be written down.
+One detail is that it remains to check that $a_i$ is known when $B_i$ is written down.
+However, writing down a block always involves writing down at least one element of the sequence, so the value of $a_i$ is always known when $B_i$ is written down.
+Note that there is an edge case where $a_i$ is not yet written down if the starting number is 1 and $i =2$.
+However, in this case, the value of $a_i$ is still known (as it is given by the rule) and thus $B_i$ can still be written down.
 \end{proof}

### 2014-11-06 draft.tex
Edits to Equivalence

@@
 \begin{align*}
-a_{n+i} = b_{k+i}
+a_{n+i} = b_{k+i},
 \end{align*}
@@
 \end{align*}
-and the sequence starting with 3 is 
+and the sequence that starts with a 3 is 
 \begin{align*}
@@
 that the resultant sequences appear to be identical and thus equivalent.
-At an intuitive level it seems obvious that these two truncated sequences are identical as they both begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers.
+At an intuitive level it seems obvious that these two truncated sequences are identical as they begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers.
 
-In order to formalize and generalize this idea we introduce the notion of a read distance.  
+In order to formalize this idea we introduce the notion of a read distance.  
 \begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
 \end{definition}
-The read distance can be any nonnegative integer. For example if we consider the sequence $\{a_i\}$ generated by $C = (1 2 3)$ beginning with a 1  
+The read distance can be any nonnegative integer. For example if we again consider generated by $C = (1 2 3)$ that starts with a $2$
 \begin{align*}
-1,2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
+2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
 \end{align*}
-the read distance for $1$ is always $0$ since in a self-referencing sequence, the first number must refer to the first block.
-The read distance for $2$ is $0$ as the second block of numbers starts at the second number.
-We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0, 0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.
+the read distance for the first number, $2$, is $0$ since in a self-referencing sequence, the first number must refer to the first block.
+The read distance for the second $2$ is $1$ as the second block of numbers starts at the third number.
+We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.
 
@@
 \label{readdistformula}
-Let $\{a_i\}$ be any self referencing sequence generated by some cycle of positive integers. Then the read distance for $i>1$, is 
+Let $\{a_i\}$ be any self referencing sequence generated by some cycle of positive integers. Then the read distance for the number at index $i>1$, is 
 \begin{equation*}
-\sum_{k=1}^{i-1}( a_k - 1)
+\sum_{k=1}^{i-1}( a_k - 1),
 \end{equation*} 
-and the read distance for $1$ is always 0.
+and the read distance for the number at index $1$ is always 0.
 \end{lemma}
@@
 \end{proof}
-Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing.
+Note that because $\{a_i\}$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing.
 
@@
 \label{equivsuff}
-Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle.If there are integers $m,l$ with the read distance for $a_m$ and $b_l$ are both $d$ and 
+Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle.If there are integers $m,l$ such that the read distance for $a_m$ and $b_l$ are both $d$ and 
+\begin{equation*}
+a_{m+k} = b_{l+k},
+\end{equation*}
+for all $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
+\end{theorem}
+
+\begin{proof}
+We will proceed by contradiction so assume otherwise. That is there exists some $j \geq d$ such that 
+\begin{equation*}
+a_{m+j} \neq b_{l+j}
+\end{equation*}
+but 
 \begin{equation*}
@@
 \end{equation*} 
-where $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
-\end{theorem}
+for all $0 \leq k < j$. 
+Since the two sequences are both generated by the same cycle this implies that either $a_{m+j}$ or $b_{l+j}$ is part of the same block as the element proceeding it, and the other is not.
+Without loss of generality we can assume that $a_{m+j}=a_{m+j-1}$ and $b_{l+j} \neq b_{l+j-1}$. 
+The length of the block that $a_{m+j-1}$ is a part of is given by some element $a_{m+n}$, for $0 \leq n < j-1$.
+Because the two sequences are identical for $0 \leq k<j$ the length of the block that $b_{l+j-1}$ is a part of is given by the element $b_{l+n}$.
+We note that $a_{m+n}=b_{l+n}$ and so the two block including $a_{m+j-1}$ must be of the same length as the block including $b_{l+j-1}$. 
+Since the two sequences are the same for $0 \leq k<j$ the number of elements in the block containing $a_{m+j-1}$ and in the block containing $b_{l+j-1}$ which have $k<j$ must be the same.
+But this means that $b_{l+j}$ is a part of the same block as the element proceeding it which is a contradiction.
+Therefore the two sequences must be equivalent.
 
-\begin{proof}
-We will prove this by  induction.
+%We will prove this by  induction.
 In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$.
@@
 
-Since a new block begins at $a_{m+d}$ and $b_{l+d}$ and $a_{m+d-1} = b_{l+d-1}$ we have $a_{m+d} = b_{l+d}$.
-Furthermore, by Proposition ~\ref{readdistformula} we have that the read distance for $a_{m+d}$ equals the read distance for $b_{l+d}$.
+%Since a new block begins at $a_{m+d}$ and $b_{l+d}$ and $a_{m+d-1} = b_{l+d-1}$ we have $a_{m+d} = b_{l+d}$.
+Furthermore, by Lemma ~\ref{readdistformula} we have that the read distance for $a_{m+d}$ equals the read distance for $b_{l+d}$.
 
-Now suppose that  for $k \leq j$ we have $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$.
+%Now suppose that  for $k \leq j$ we have $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$.
 We would like to show that this holds for $k = j+1$.
-By Proposition ~\ref{readdistformula} we actually already know that the read distance for the two is the same, as it is the same for the preceding terms, which are equal.
+By Lemma ~\ref{readdistformula} we actually already know that the read distance for the two is the same, as it is the same for the preceding terms, which are equal.
 
-If $a_{m+k}$ is the last term in a block then $b_{m+k}$ must also be the last term in its block, because the length of blocks is determined by earlier terms which we know to be equal and to have the same read distance.
+%If $a_{m+k}$ is the last term in a block then $b_{m+k}$ must also be the last term in its block, because the length of blocks is determined by earlier terms which we know to be equal and to have the same read distance.
 Since $a_{m+k} = b_{l+k}$  in this case we have $a_{m+k+1} = b_{l+k+1}$.
 
-Otherwise $a_{m+k}$ and $a_{m+k+1}$ are in the same block, and so we only need to show that $b_{l+k}$ and $b_{l+k+1}$ are in the same block as well.
+%sOtherwise $a_{m+k}$ and $a_{m+k+1}$ are in the same block, and so we only need to show that $b_{l+k}$ and $b_{l+k+1}$ are in the same block as well.
 But as we argued above the length of blocks is determined by terms earlier in the sequence, which we know to be equal and have the same read distance.
@@
 
-This result also allows us to make a general statement for equivalence for sequences that begin with a $1$.
+This result also allows us to make a general statement for equivalence of sequences that begin with a $1$.
 \begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.

### 2014-11-06 draft.tex
Edits to Equivalence

@@
 \begin{equation*}
-\sum_{k=1}^{i-1}( a_k - 1),
+\sum_{k=1}^{i-1}( a_k - 1).
 \end{equation*} 
-and the read distance for the number at index $1$ is always 0.
+The read distance for the number at index $1$ is always 0.
 \end{lemma}
@@
 \end{equation*} 
-This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block.
-Also note that $a_{l+1}$ is one position to the right of $a_l$.
+This $l$th block has length $a_l$ and so the $(l+1)$th block will begin $a_l$ positions to the right of the beginning of the $l$th block.
+Additionally $a_{l+1}$ is one position to the right of $a_l$.
 Therefore the beginning of the $(l+1)$th block is
 \begin{equation*}
-a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
+(a_l-1 )+ \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
 \end{equation*} 
-positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
+positions to the right of $a_{l+1}$. But this is exactly the desired read distance for $a_{l+1}$ and so we have our result.
 \end{proof}
-Note that because $\{a_i\}$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing.
+Note that because $\{a_i\}$ is generated by positive integers we always have $a_k-1\geq 0$ and so read distance is increases monotonically.
 
@@
 
-Now that we have a nice expression for the value of the read distance at a particular index we can show our first major result which allows us to check only finitely many indices in order to establish that two sequences are equivalent.
+Now that we have a nice expression for the value of the read distance at a particular index we can show our first major result.
+It will allow us to check only finitely many indices in order to establish that two sequences are equivalent.
 \begin{theorem} 
 \label{equivsuff}
-Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle.If there are integers $m,l$ such that the read distance for $a_m$ and $b_l$ are both $d$ and 
+Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle. If there are integers $m,l$ such that the read distance for $a_m$ and $b_l$ are both $d$ and 
 \begin{equation*}
@@
 \end{equation*}
-for all $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
+for all $0 \leq k \leq d-1$, then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
 \end{theorem}
@@
 
-We also can demonstrate a sufficient condition for equivalence. 
+We can also demonstrate a sufficient condition for equivalence. 
 \begin{theorem}
@@
 \begin{corollary}
-If two self-referencing sequences are equivalent, then their read distance of the shared subsequence is the same everywhere.
+If two self-referencing sequences are equivalent, then the read distance of their shared subsequence is the same everywhere.
 \end{corollary}
@@
 
-\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent, then whenever two elements have the same read distances, the two elements are equal. 
+\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent, then whenever two elements have the same read distances, other than for $i=0$,  the two elements are equal. 
 \end{conj}

### 2014-11-06 draft.tex
More Updates to Equivalence

@@
 \begin{align*}
-a_{n+i} = b_{k+i},
+a_{n+i} = b_{k+i}, \quad i=1,2,3,\ldots 
 \end{align*}
-for all positive integers $i$. 
 \end{definition}
+Note that two sequences can only be equivalent if they are generated by the same cycle.
 
-As an example let us consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
+As a basic example of equivalence let us consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
 \begin{align*}
@@
 \end{align*}
-that the resultant sequences appear to be identical and thus equivalent.
-At an intuitive level it seems obvious that these two truncated sequences are identical as they begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers.
+that the resultant sequences appear to be identical and so the sequences should be equivalent.
+At an intuitive level it seems obvious that these two truncated sequences are identical as they begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers, but we cannot actually show that these two sequences are equivalent using just our definition as we would need to check equality for an infinite number of terms.
 
-In order to formalize this idea we introduce the notion of a read distance.  
+In order to develop a method to address this problem we introduce the notion of a read distance.  
 \begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
@@
 
-Now that we have a nice expression for the value of the read distance at a particular index we can show our first major result.
-It will allow us to check only finitely many indices in order to establish that two sequences are equivalent.
+Now that we have a nice expression for the value of the read distance at a given index we can show our first major result.
+It will allow us to check only finitely many indices in order to establish that two sequences are equivalent, thus solving the problem we encountered in our example.
 \begin{theorem} 
 \label{equivsuff}
-Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle. If there are integers $m,l$ such that the read distance for $a_m$ and $b_l$ are both $d$ and 
+Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle. If there are integers $m,l$ such that the read distance for $a_m$ and $b_l$ is $d$ and 
 \begin{equation*}
-a_{m+k} = b_{l+k},
+a_{m+k} = b_{l+k}, \quad 0 \leq k \leq d-1
 \end{equation*}
-for all $0 \leq k \leq d-1$, then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
+then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
 \end{theorem}
@@
 for all $0 \leq k < j$.
-Since the two sequences are both generated by the same cycle this implies that either $a_{m+j}$ or $b_{l+j}$ is part of the same block as the element proceeding it, and the other is not.
-Without loss of generality we can assume that $a_{m+j}=a_{m+j-1}$ and $b_{l+j} \neq b_{l+j-1}$. 
-The length of the block that $a_{m+j-1}$ is a part of is given by some element $a_{m+n}$, for $0 \leq n < j-1$.
-Because the two sequences are identical for $0 \leq k<j$ the length of the block that $b_{l+j-1}$ is a part of is given by the element $b_{l+n}$.
-We note that $a_{m+n}=b_{l+n}$ and so the two block including $a_{m+j-1}$ must be of the same length as the block including $b_{l+j-1}$. 
-Since the two sequences are the same for $0 \leq k<j$ the number of elements in the block containing $a_{m+j-1}$ and in the block containing $b_{l+j-1}$ which have $k<j$ must be the same.
-But this means that $b_{l+j}$ is a part of the same block as the element proceeding it which is a contradiction.
-Therefore the two sequences must be equivalent.
+
+Since the two sequences are both generated by the same cycle this implies that either $a_{m+j}$ or $b_{l+j}$ is part of the same block as the element proceeding it, while the other is not.
+Without loss of generality we will assume that $a_{m+j}=a_{m+j-1}$ and $b_{l+j} \neq b_{l+j-1}$.
+
+The length of the block that $a_{m+j-1}$ is a part of is $a_{m+n}$, for some $0 \leq n < j-1$.
+Because the two sequences are identical for $0 \leq k<j$ and by our assumption that the read distance is the same for $a_{m}$ and $b_{l}$, the read distance for $b_{l+n}$ equals the read distance for $a_{m+n}$.
+This combined with the fact that $a_{m+n} = b_{l+n}$ shows that the block including $a_{m+j-1}$ must be of the same length as the block including $b_{l+j-1}$.
+
+Since the two sequences are the same for $0 \leq k<j$ the number of elements in the block containing $a_{m+j-1}$, with $k<j$, and the number of elements in the block containing $b_{l+j-1}$, with $k<j$, must be the same.
+Because $a_{m+j}$ is part of the same block as the element preceding it, this means that $b_{l+j}$ must be a part of the same block as the element proceeding it as well. 
+This contradicts the fact that exactly one of the two elements is not in the same block as the element preceding it and so the two sequences must be equivalent.
 
@@
 The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0.
-In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
+In addition, $a_2 = b_1 = c_1$, so by Theorem ~\ref{equivsuff} the two sequences are equivalent.
 \end{proof}
@@
 
-\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent, then whenever two elements have the same read distances, other than for $i=0$,  the two elements are equal. 
+\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent, then whenever two elements have the same read distances, other than for $i=1$,  the two elements are equal. 
 \end{conj}

### 2014-11-06 draft.tex
rewrote densities section, broke it up into two sections

@@
 
-\section{Density}
-\label{density}
-\emph{Written by Erik Waingarten, edited by Perry Kleinhenz and Fermi Ma}
+\section{A Density Computation}
+\label{densitycomputation}
+\emph{Written by Erik Waingarten and Fermi Ma, edited by Perry Kleinhenz}
 
-A topic of interest when considering a self referencing sequence is the relative density of the various elements of the generating cycle.
-It is unclear if such a density even exists and if so how to compute it for arbitrary generating cycles.
-In this section, we show how to use substitution rules to compute relative densities for certain sequences if we assume such densities exists.
-We explain the method and compute the relative density of 3's for $C = (1 3)$.
-We will also produce bounds on the density of $2$'s $C = (1 2)$ in a way that suggests that the density converges to 2; however, this method is computationally difficult for smaller bounds since the number of cases to check increases exponentially.
+An open question regarding the self-referential sequence of 1's and 2's, $s((1 2),1)$, is what the limiting \emph{density} of 1's and 2's is.
+In this section, we consider the related problem of finding the density for sequences with generating cycle $C = (1 3)$ (note that the sequence that starts with a 1 differs by only one number from the sequence that starts with a 3, so we do not need to consider them separately).
+While we cannot prove that a limiting density exists, we give a numerical calculation of what the density is \emph{if it exists}.
+We use a general technique that takes advantage of the fact that substitution rules when $C = (1 3)$, as shown in Section~\ref{substitutionrules}.
 
-To begin we will discuss a technique that can be used to compute the densities for some generating cycles.
-In particular, suppose we have a generating cycle $C$ with a set of rules $R$ that satisfy the conditions from the section above.
-For each rule $r \in R$, we will say denote the rule as $r_l \rightarrow r_r$ as the inputs and the outputs.
+In Section~\ref{substitutionrules}, we found that repeated applications of the rules
+\begin{align*}
+\text{Rule }1: 3,3 &\to 3,3,3,1,1,1 \\
+\text{Rule }2: 3,1 &\to 3,3,3,1\\
+\text{Rule }3: 1,1 &\to 3,1
+\end{align*}
+to the initial sequence $\{3,1\}$ gives a growing sequence that converges to the self-referential sequence $s((1 3),3)$.
 
-Now we can count the fraction of times $r_r$ appears in the sequence at each step.
-Since we know the rules, these will become a set of $|R|$ equalities that reference each other.
-We can take their ratios and solve for the fraction of times that each $r_r$ appears and compute the density by assuming a stable state.
+If we let $A$ denote the subsequence $\{3,3\}$, $B$ denote $\{3,1\}$ and let $C$ denote $\{1,1\}$, then the rules can be rewritten as
+\begin{align*}
+\text{Rule }1: A &\to ABC \\
+\text{Rule }2: B &\to AB\\
+\text{Rule }3: C &\to B,
+\end{align*}
+where AB and ABC are to be interpreted as concatenations of the subsequences.
 
-For example, lets look at the rules we have determined for $G = \{ 1, 3 \}$. 
+Our technique for computing densities will work as follows.
+Let $N_i(A), N_i(B),$ and $N_i(C)$ denote the number of $A$'s, $B$'s, and $C$'s, respectively, in the sequence given by the $i$th application of the substitution rules.
+We keep track of these counts of $A$'s, $B$'s, and $C$'s at each round of rule applications, and we use these to determine limiting proportions of $A$'s, $B$'s and $C$'s.
+From these limiting proportions, we can deduce the densities of 1's and 3's.
+
+To start, consider the substitution process from the very first step. We start applying substitution rules to the initial sequence $\{3,1\}$. Since $\{3,1\} = B$, the initial counts are:
+\[\text{Sequence at step }1: \{3,1\},\text{ Counts}: N_1(A) = 0, N_1(B) = 1, N_1(C) = 0.\]
+
+After one application of the substitution rules, we have
+\[\text{Sequence at step }2: \{3,3,3,1\},\text{ Counts}: N_2(A) = 1, N_2(B) = 1, N_2(C) = 0.\]
+
+After one more application, we have
+\[\text{Sequence at step }3: \{3,3,3,1,1,1,3,3,3,1\}, \text{ Counts}: N_3(A) = 2, N_3(B) = 2, N_3(C) = 1.\]
+
+For the purposes of computing densities, it turns out that only the counts are relevant and we do not need to keep track of the sequence at each step.
+We can compute the counts $N_i$ from the values of $N_{i-1}$ as follows
 \begin{align*}
-\text{Rule }A: 3,3 &\to 3,3,3,1,1,1 \\
-\text{Rule }B: 3,1 &\to 3,3,3,1\\
-\text{Rule }C: 1,1 &\to 3,1
+N_i(A) &= N_{i-1}(A) + N_{i-1}(B)\\
+N_i(B) &= N_{i-1}(A) + N_{i-1}(B) + N_{i-1}(C)\\
+N_i(C) &= N_{i-1}(A).
 \end{align*}
-Then we can let $3,3,3,1,1,1 = A$, $3,3,3,1 = B$ and $3, 1 = C$. If $A_i$ is the number of occurrences of $A$ in $s^{(i)}$, then 
-\[ A_i = B_{i-1} + A_{i-1} \]
-Likewise, we have
-\[ B_i = A_{i-1} + B_{i-1} + C_{i-1} \]
-\[ C_i = A_{i-1} \]
-So now we can take the fraction by dividing by the total number of them:
-\[ A_i' = \dfrac{B_{i-1}' + A_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
-\[ B_i' = \dfrac{A_{i-1}' + B_{i-1}' + C_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
-\[ C_i' = \dfrac{A_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
-So if we assume a density, there will be a stable state which will be the density, so in the case where $A_i'=A_{i-1}', B_i' = B_{i-1}', C_i'=C_{i-1}'$, we can solve for the fractional appearance of each and get the density of the numbers.
+The above equations are arrived at by looking at the set of rules involving $A,B,$ and $C$.
+For example, an $A$ at the $i$th iteration is written down every time there is an $A$ or $B$ in the previous iteration, which leads to the first equation above.
 
-We can compute solutions to this set of quadratic equations and find the density of 3's is approximately $0.602$.
+For computing densities, however, the \emph{proportions} of $A$'s, $B$'s, and $C$'s are needed.
+Fortunately, we can derive equations for the proportions by simply normalizing the equations for the counts.
+Let $P_i(X)$ denote the proportion of the sequence after iteration $i$ that is made up of the $X$ subsequence.
+Normalizing gives the following equations:
+\begin{align*}
+P_i(A) &= \frac{P_{i-1}(A) + P_{i-1}(B)}{3P_{i-1}(A) + 2P_{i-1}(B) + P_{i-1}(C)}\\
+P_i(B) &= \frac{P_{i-1}(A) + P_{i-1}(B) + P_{i-1}(C)}{3P_{i-1}(A) + 2P_{i-1}(B) + P_{i-1}(C)}\\
+P_i(C) &= \frac{P_{i-1}(A)}{3P_{i-1}(A) + 2P_{i-1}(B) + P_{i-1}(C)}.
+\end{align*}
+The only difference is that each equation is divided on the right-hand side by $3P_{i-1}(A) + 2P_{i-1}(B) + P_{i-1}(C)$ to ensure that the proportions of $A$'s, $B$'s, and $C'$s at each step add up to 1.
+Now, assume that there exist limiting proportions (and hence, limiting densities) as $i \to \infty$.
+Then at this limit, $P_i(X) = P_{i-1}(X) = P_{\infty}(X)$.
+Making these substitutions and solving for $P_{\infty}(A), P_{\infty}(B),$ and $P_{\infty}(C)$ in \texttt{Mathematica} gives the following values:
+\begin{align*}
+P_{\infty}(A) &\approx \texttt{0.3760858894420931}\\
+P_{\infty}(B) &\approx \texttt{0.4533976515164037}\\
+P_{\infty}(C) &\approx \texttt{0.1705164590415032}.
+\end{align*}
 
-As discussed in the previous section, it is unclear how to establish a set of rules for $C = (1 2)$; however, there is some structure that we can exploit.
+Since $A = \{3,3\}$, $B = \{3,1\}$ and $C = \{1,1\}$, the fraction of 1's in the limiting sequence is $0.5P_{\infty}(B) + P_{\infty}(C)$, and the fraction of 3's in the limiting sequence is $P_{\infty}(A) + 0.5P_{\infty}(B)$.
+So the density values, if they exist, are
+\begin{align*}
+&\text{Fraction of 1's: }\texttt{0.397215284799705}\\
+&\text{Fraction of 3's: }\texttt{0.602784715200295}.
+\end{align*}
+We were able to support these calcuations with numerical simulations.
+We generated the $s((13),3)$ sequence up to 100,000 terms, and we found that the fraction of 1's was approximately $0.398$, and that the fraction of 3's was approximately 0.602.
+
+It is difficult to fully generalize this procedure for all substitution rule sets, as we cannot prescribe a general form for all substitution rules.
+However, we believe that this technique should work whenever all the inputs to the substitution rules are the same length (in this case, the inputs were all length 2), and the outputs are all lengths that are multiples of the inputs.
+We do not know if the resulting equations will always give exactly 1 solution, as they did in this case.
+
+\section{Density Bounds for $C = (12)$}
+\label{densitybounds}
+
+As discussed in the Section~\ref{substitutionrules}, it is unclear how to establish a set of rules for $C = (1 2)$; however, there is some structure that we can exploit.
 In particular, we know that the sequences $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence.

### 2014-11-06 draft.tex
Equivalence Edits

@@
 
-We can also demonstrate a sufficient condition for equivalence. 
+We can also demonstrate a necessary condition for equivalence. 
 \begin{theorem}
@@
 \begin{proof}
-Assume otherwise. Suppose the two sequences are $\{a_j\}$ and $\{b_j\}$ and we have 
+Suppose the two sequences are $\{a_j\}$ and $\{b_j\}$ and we have 
 \begin{equation*}
-a_{i+l} = b_{i+k} = c_{i}
+a_{m+i} = b_{l+i} = c_{i} \quad 1 \leq i
 \end{equation*}
-for all $i>1$, such that if $m$ or $l$ were made smaller this relation would not hold.
-Suppose the read distance for $b_{k+1}$ is $x$ and the read distance for $a_{1+l}$ is $y$.
-Without loss of generality let us assume that $x>y$ (note that $x \neq y$ as per our assumption).
+such that if $m$ or $l$ were made smaller this relation would not hold.
+We will proceed by contradiction so assume that the read distance for $a_{m+1}$ is $x$ and the read distance for $b_{l+1}$ is $y$ such that $x<y$.
 
-Since the two read distances are different we know there are two separate blocks of length $c_1$ in the sequence, one which corresponds to $a_{i+l}$ and another which corresponds to $b_{i+k}$. 
-Because the two sequences are equal after a certain point and $\{a_j\}$ is self referential there exists some index $n$ such that $c_{n+1}=a_{n+l}$ is the element of $\{a_j\}$ which corresponds to the second block of length $c_1$.
+%Since the two sequences are equivalent the block that begins at index $m+x+1$ in $\{a_i\}$ begins at index $l+x+1$ in $\{b_i\}$.
+The element which refers to this block in $\{b_i\}$ must appear before the index $l+1$, because the element at index $l+1$ refers to the block beginning at $l+y+1$ and read distance is non-decreasing.
+Furthermore this element of $\{b_i\}$ must have the same value as $a_{m+1}$, because the two blocks are of the same length.
 
-We now claim that $c_1, c_2, \ldots, c_{n}$  must appear in $\{b_j\}$ such that $b_{i+k-1}=c_{n}$.
-This is true because $\{b_j\}$ is self referential, and so the length of the blocks of the sequence $\{c_i\}$ must reproduce the terms of $\{b_j\}$.
-In particular the lengths of the blocks which occur before the second block of length $c_1$ must appear in $\{b_j\}$, and must end right next to the number which refers to the second block of length $c_1$.
+Since the two read distances are different we know there are two separate blocks of length $c_1$ in the common sequence, one which corresponds to $a_{m+1}$ and another which corresponds to $b_{l+1}$.
+Note that the block corresponding to $b_{l+1}$ is to the right of the block corresponding to $a_{m+1}$ because the read distance for $b_{l+1}$ is larger.
+Now because $\{a_i\}$ is self referential there exists some index $n$ such that $a_{m+n+1}$ is the element of $\{a_i\}$ which corresponds to the second block of length $c_1$.
 
-Thus our sequence $\{b_j\}$ is  
+We claim that
 \begin{equation*}
-\cdots c_1 c_2 c_3 \cdots c_{n} c_1 c_2 \cdots c_n \cdots
+c_i = b_{l-n+i} \quad 1 \leq i \leq n
 \end{equation*}
-where the prefix ends at the second $c_1$. We now claim that the common sequence cycles through
+That is the subsequence $c_1, c_2, \ldots, c_{n}$ appears in $\{b_i\}$ such that the last term is next to the beginning of the common sequence.
+This is true because the length of the blocks of $\{c_i\}$ which occur before the second block of length $c_1$ must appear as terms of $\{b_i\}$ and must end next to the number which refers to the second block of length $c_1$.
+
+Thus our sequence $\{b_i\}$ is  
 \begin{equation*}
-c_1 c_2 \cdots c_n.
+\cdots c_1 c_2 \cdots c_{n} c_1 c_2 \cdots c_n \cdots
 \end{equation*}
-Since the read distance for $b_{1+k}$ is greater than the read distance for $a_{1+l}$, and the sequences are identical for subsequent indices by Lemma ~\ref{readdistformula}, we know that the read distance for $b_{i+k}$ is longer than the read distance for $a_{i+l}$ by exactly $x-y$.
-Thus $b_{k+i}$ corresponds to the same block as $a_{l+(x-y)+i}$.
-But $b_{k+i} = c_{i}$ and $a_{l+(x-y)+i}=c_{x-y+i}$ and so $c_{i} = c_{x-y+i}$ and so the common sequence cycles.
-Because the common sequence cycles and $\{b_j\}$ has the same cycle, but one which begins before the common sequence, $l$ could be made smaller and have the equivalence relation hold.
-This is contradicts our statement that $l$ is as small as possible.
+where the beginning of the common sequence is the second $c_1$. We now claim that the common sequence is just the subsequence
+\begin{equation*}
+c_1 c_2 \cdots c_n,
+\end{equation*}
+repeated an infinite number of times.
+
+Since the read distance for $b_{l+1}$ is greater than the read distance for $a_{m+1}$, and the sequences are identical for subsequent indices, by Lemma ~\ref{readdistformula} we know that the read distance for $b_{l+i}$ is larger than the read distance for $a_{m+i}$ by exactly $y-x$.
+
+Thus $b_{l+i}$ and $a_{m+(y-x)+i}$ correspond to the same block. But $b_{l+i} = c_{i}$ and $a_{m+(x-y)+i}=c_{x-y+i}$. Therefore $c_{i} = c_{y-x+i}$ and so the common sequence repeats.
+
+Because the common sequence repeats and $\{b_i\}$ has the same cycle, but one which begins before the common sequence, $l$ could be made smaller and have the equivalence relation hold.
+This contradicts our statement that $l$ is as small as possible.
 Therefore the first element of the shared sequence must have the same read distance in both sequences.
@@
 \begin{corollary}
-If two self-referencing sequences are equivalent, then the read distance of their shared subsequence is the same everywhere.
+If two self-referencing sequences are generated over the same cycle and  then the read distance of their shared subsequence is the same everywhere.
 \end{corollary}
@@
 
-These conditions for equivalence lead us to make the following conjecture on a sufficient condition for nonequivalence.
-
-\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent, then whenever two elements have the same read distances, other than for $i=1$,  the two elements are equal. 
-\end{conj}
-
-\begin{conj} Any sequence generated by a cycle $C$ is equivalent to any other sequence generated by the same cycle $C$.
-\end{conj}
-
-This conjecture is partly motivated by the fact that the property holds for self-referencing sequences from the generating cycle $C= (123)$.
-For generating cycles with larger numbers, we have been unable to verify or refute the conjecture..
-However, in Section~\ref{density}, we show that for some generating cycles the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
+It is unclear to us whether all sequences generated by the same cycle are equivalent.
+This is certainly the case for $C=(123)$ and any two element cycles but we have been unable to generalize this to other three element or larger cycles.
+A key ingredient in establishing a counterexample would be more stringent necessary conditions for equivalence but it is unclear what these conditions should be or how to prove they are necessary.
+In order to establish that all sequences generated by the same cycle are in fact equivalent we would like to establish weaker or more generalizable sufficient conditions for equivalence, but based on our proof of Theorem ~\ref{eqsuff} it does not seem like weaker conditions will actually be sufficient.
 

### 2014-11-06 draft.tex
Typo fix

@@
 \begin{corollary}
-If two self-referencing sequences are generated over the same cycle and  then the read distance of their shared subsequence is the same everywhere.
+If two self-referencing sequences are generated over the same cycle  and are equivalent then the read distance of their shared subsequence is the same everywhere.
 \end{corollary}
@@
 A key ingredient in establishing a counterexample would be more stringent necessary conditions for equivalence but it is unclear what these conditions should be or how to prove they are necessary.
-In order to establish that all sequences generated by the same cycle are in fact equivalent we would like to establish weaker or more generalizable sufficient conditions for equivalence, but based on our proof of Theorem ~\ref{eqsuff} it does not seem like weaker conditions will actually be sufficient.
+In order to establish that all sequences generated by the same cycle are in fact equivalent we would like to establish weaker or more generalizable sufficient conditions for equivalence, but based on our proof of Theorem ~\ref{equivsuff} it does not seem like weaker conditions will actually be sufficient.
 

### 2014-11-06 draft.tex
edited first half of equivalence section

@@
 \end{proof}
-%
-%\begin{lemma}
-%Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
-%\end{lemma}
-%
-%\begin{proof}
-%Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different.
-Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$.
-This is a contradiction since $a_j \neq b_j$.
-So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block.
-Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block.
-However, this means that $a_k \geq b_k + 1$.
-So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$.
-Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.
-%
-%\begin{figure}{hb}
-%\label{fig:unequalsequences}
-%\center
-%\includegraphics[width=0.5\linewidth]{lemma1.pdf}
-%\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
-%\end{figure}
-%
-%This implies that every block before the $k$th block just contains 1's since each block refered it itself.
-This means that either $j = 2$, or the sequence is not self-referential.
-But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle.
-%\end{proof}
-%
-%
-%
-%We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element.
-Assume the generating cycle is $(a_1 a_2 \dots a_m)$ and that the starting element is some $a_i$.
-For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
-%\begin{quote}
-%\textbf{Algorithm:}\\
-%Suppose the first number of the sequence is $a_i$.
-Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block.
-If $a_i$ is not 1, then there is a second element of that block.
-We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence.
-If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
-%
-%This procedure continues deterministically so long as there is an element to read.
-Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
-%\end{quote}
 
@@
 
-In this section, we consider long term equivalence of sequences. We introduce a formal definition of this and establish results that allow us to determine when two sequences are equivalent.
+In this section, we consider equivalence of the limiting behavior of sequences.
+We introduce a formal definition of this and establish results that allow us to determine when two sequences are equivalent.
 
-We would first like to make a formal definition of equivalence. 
+We first make a formal definition of equivalence. 
 \begin{definition}
-We say that two sequence $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n, k \geq 0$ such that 
+We say that two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent if there exist integers $n, k \geq 0$ such that 
 \begin{align*}
-a_{n+i} = b_{k+i}, \quad i=1,2,3,\ldots 
+a_{n+i} = b_{k+i}, \quad i=1,2,3,\ldots \ .
 \end{align*}
@@
 
-As a basic example of equivalence let us consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
+As a basic example of equivalence, consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
 \begin{align*}
@@
 \end{align*}
-If we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we see 
+If we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, as follows
 \begin{align*}
@@
 \end{align*}
-that the resultant sequences appear to be identical and so the sequences should be equivalent.
-At an intuitive level it seems obvious that these two truncated sequences are identical as they begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers, but we cannot actually show that these two sequences are equivalent using just our definition as we would need to check equality for an infinite number of terms.
+we see that the resultant sequences are the same, and thus the original sequences are in some sense equivalent.
+At an intuitive level, it seems obvious that these two truncated sequences are identical since they begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers.
+We now formalize this intuition for claiming that two sequences are in fact equivalent.
 
-In order to develop a method to address this problem we introduce the notion of a read distance.  
-\begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
+To do this, we first introduce the \emph{read distance}.
+\begin{definition} Consider a sequence $\{a_i\}$.
+The read distance for an index $i$ is the difference $index(B_i) - i$, where $index(B_i)$ is the index of the first number in the $i$th block $B_i$ of numbers.
 \end{definition}
-The read distance can be any nonnegative integer. For example if we again consider generated by $C = (1 2 3)$ that starts with a $2$
+Note that the read distance can be any nonnegative integer. For example if we again consider the sequence generated by $C = (1 2 3)$ that starts with a $2$
 \begin{align*}
@@
 \end{align*}
-the read distance for the first number, $2$, is $0$ since in a self-referencing sequence, the first number must refer to the first block.
-The read distance for the second $2$ is $1$ as the second block of numbers starts at the third number.
-We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.
+the read distance for index 1 is $index(B_1) - 1 = 1 - 1 = 0$.
+The read distance for index 2 is $index(B_2) - 2 = 3 - 2 = 1$.
+We can continue to calculate the read distances at each index of this sequence and we see that they form a monotonically increasing sequence
+\[ 0,1,2,4,6,6,6,6,7,8,9,\ldots \]
 
-This monotonicity is true in general, and we can in fact make a stronger statement on the value of the read distance at a given index.
-\begin{lemma} 
+This monotonicity is present in sequences other than ones generated by $(123)$. We can explain this monotonicity phenomenon with the following proposition.
+\begin{proposition} 
 \label{readdistformula}
-Let $\{a_i\}$ be any self referencing sequence generated by some cycle of positive integers. Then the read distance for the number at index $i>1$, is 
+Let $\{a_i\}$ be any self-referencing sequence generated by some cycle of positive integers. Then the read distance for index $j$ is 0 if $j =1$, and when $j>1$, the read distance is
 \begin{equation*}
-\sum_{k=1}^{i-1}( a_k - 1).
+\sum_{k=1}^{j-1}( a_k - 1).
 \end{equation*} 
-The read distance for the number at index $1$ is always 0.
-\end{lemma}
+\end{proposition}
 \begin{proof}
-We will prove this using induction.
-We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number.
-When $i=2$ the read distance is $a_1-1$.
-The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$.
-The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.
+We prove this using induction.
+We know that for $i=1$, the read distance is $0$ as the first block always starts at the first number.
+When $i=2$, the read distance is $a_1-1$.
+The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at index $a_1+1$.
+So the read distance is $index(B_2) - 2 = a_1 -1$.
 
-Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the start of the $l$th block and position $l$ is 
+For the inductive step, suppose the formula holds for $i=l > 1$. This means that the read distance of index $l$ is
 \begin{equation*}
-\sum_{k=1}^{l-1}( a_k - 1).
+index(B_l) - l = \sum_{k=1}^{l-1}( a_k - 1).
 \end{equation*} 
-This $l$th block has length $a_l$ and so the $(l+1)$th block will begin $a_l$ positions to the right of the beginning of the $l$th block.
-Additionally $a_{l+1}$ is one position to the right of $a_l$.
-Therefore the beginning of the $(l+1)$th block is
-\begin{equation*}
-(a_l-1 )+ \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
-\end{equation*} 
-positions to the right of $a_{l+1}$. But this is exactly the desired read distance for $a_{l+1}$ and so we have our result.
+The $l$th block has length $a_l$, which implies that $index(B_{l+1}) - index(B_l) = a_l$. Thus, the read distance of index $l+1$ is
+\begin{align*}
+index(B_{l+1}) - (l+1) = a_l - 1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1),
+\end{align*}
+which compeltes the induction.
 \end{proof}
-Note that because $\{a_i\}$ is generated by positive integers we always have $a_k-1\geq 0$ and so read distance is increases monotonically.
+Note that because $\{a_i\}$ is generated by positive integers, we always have $a_k-1\geq 0$ and so the read distance increases monotonically.
 
@@
 
-Now that we have a nice expression for the value of the read distance at a given index we can show our first major result.
-It will allow us to check only finitely many indices in order to establish that two sequences are equivalent, thus solving the problem we encountered in our example.
+Now that we have a closed-form expression for the read distance at a given index, we prove our first major result: we can check only finitely many indices in order to establish that two sequences are equivalent.
 \begin{theorem} 
 \label{equivsuff}
-Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle. If there are integers $m,l$ such that the read distance for $a_m$ and $b_l$ is $d$ and 
+Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle. If there exist integers $m,l$ such that the read distances for $a_m$ and $b_l$ are both equal to $d$ and 
 \begin{equation*}
-a_{m+k} = b_{l+k}, \quad 0 \leq k \leq d-1
+a_{m+k} = b_{l+k}, \quad 0 \leq k \leq d-1,
 \end{equation*}
@@
 \begin{proof}
-We will proceed by contradiction so assume otherwise. That is there exists some $j \geq d$ such that 
+We will proceed by contradiction, so first assume otherwise. That is, assume there exists some $j \geq d$ such that 
 \begin{equation*}
@@
 
-Since the two sequences are both generated by the same cycle this implies that either $a_{m+j}$ or $b_{l+j}$ is part of the same block as the element proceeding it, while the other is not.
+Since the two sequences are both generated by the same cycle, this implies that either $a_{m+j}$ or $b_{l+j}$ is part of the same block as the element proceeding it, while the other is not.
 Without loss of generality we will assume that $a_{m+j}=a_{m+j-1}$ and $b_{l+j} \neq b_{l+j-1}$.

### 2014-11-06 draft.tex
finished modifying equivalence section

@@
 Because the two sequences are identical for $0 \leq k<j$ and by our assumption that the read distance is the same for $a_{m}$ and $b_{l}$, the read distance for $b_{l+n}$ equals the read distance for $a_{m+n}$.
-This combined with the fact that $a_{m+n} = b_{l+n}$ shows that the block including $a_{m+j-1}$ must be of the same length as the block including $b_{l+j-1}$.
+This combined with the fact that $a_{m+n} = b_{l+n}$ shows that the block containing $a_{m+j-1}$ must be of the same length as the block containing $b_{l+j-1}$.
 
-Since the two sequences are the same for $0 \leq k<j$ the number of elements in the block containing $a_{m+j-1}$, with $k<j$, and the number of elements in the block containing $b_{l+j-1}$, with $k<j$, must be the same.
+Since the two sequences are the same for $0 \leq k<j$, the number of elements in the block containing $a_{m+j-1}$, with $k<j$, and the number of elements in the block containing $b_{l+j-1}$, with $k<j$, must be the same.
 Because $a_{m+j}$ is part of the same block as the element preceding it, this means that $b_{l+j}$ must be a part of the same block as the element proceeding it as well.
@@
 
-This result also allows us to make a general statement for equivalence of sequences that begin with a $1$.
+This result also allows us to make a general statement regarding equivalence of sequences that begin with a $1$.
 \begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
@@
 \end{proof}
+Note that one can check this corollary easily by hand, but our formalism provides a way to mathematically verify this intuition.
 
@@
 \label{thm:readfirst}
-If two sequences are equivalent then the first element of the shared sequence must have the same read distance in both sequences.
+If two sequences are equivalent, then the first element of the shared sequence must have the same read distance in both sequences.
 \end{theorem}
@@
 
-Since the two read distances are different we know there are two separate blocks of length $c_1$ in the common sequence, one which corresponds to $a_{m+1}$ and another which corresponds to $b_{l+1}$.
+Since the two read distances are different, we know there are two separate blocks of length $c_1$ in the common sequence, one which corresponds to $a_{m+1}$ and another which corresponds to $b_{l+1}$.
 Note that the block corresponding to $b_{l+1}$ is to the right of the block corresponding to $a_{m+1}$ because the read distance for $b_{l+1}$ is larger.
-Now because $\{a_i\}$ is self referential there exists some index $n$ such that $a_{m+n+1}$ is the element of $\{a_i\}$ which corresponds to the second block of length $c_1$.
+Now because $\{a_i\}$ is self-referential, there exists some index $n$ such that $a_{m+n+1}$ is the element of $\{a_i\}$ which corresponds to the second block of length $c_1$.
 
@@
 \begin{equation*}
-c_i = b_{l-n+i} \quad 1 \leq i \leq n
+c_i = b_{l-n+i} \quad 1 \leq i \leq n.
 \end{equation*}
-That is the subsequence $c_1, c_2, \ldots, c_{n}$ appears in $\{b_i\}$ such that the last term is next to the beginning of the common sequence.
+That is, the subsequence $c_1, c_2, \ldots, c_{n}$ appears in $\{b_i\}$ such that the last term is next to the beginning of the common sequence.
 This is true because the length of the blocks of $\{c_i\}$ which occur before the second block of length $c_1$ must appear as terms of $\{b_i\}$ and must end next to the number which refers to the second block of length $c_1$.
@@
 \begin{equation*}
-\cdots c_1 c_2 \cdots c_{n} c_1 c_2 \cdots c_n \cdots
+\cdots c_1 c_2 \cdots c_{n} c_1 c_2 \cdots c_n \cdots \ ,
 \end{equation*}
@@
 \begin{corollary}
-If two self-referencing sequences are generated over the same cycle  and are equivalent then the read distance of their shared subsequence is the same everywhere.
+If two self-referencing sequences are generated over the same cycle and are equivalent, then the read distance of their shared subsequence is the same everywhere.
 \end{corollary}
@@
 
-It is unclear to us whether all sequences generated by the same cycle are equivalent.
-This is certainly the case for $C=(123)$ and any two element cycles but we have been unable to generalize this to other three element or larger cycles.
-A key ingredient in establishing a counterexample would be more stringent necessary conditions for equivalence but it is unclear what these conditions should be or how to prove they are necessary.
-In order to establish that all sequences generated by the same cycle are in fact equivalent we would like to establish weaker or more generalizable sufficient conditions for equivalence, but based on our proof of Theorem ~\ref{equivsuff} it does not seem like weaker conditions will actually be sufficient.
+It is unclear whether all sequences generated by the same cycle are equivalent.
+This is certainly the case for $C=(123)$ and any two element cycles, but we have been unable to generalize this to other three element or larger cycles.
+A key ingredient in establishing a counterexample would be more stringent necessary conditions for equivalence, but it is unclear what these conditions should be or how to prove they are necessary.
+If we want to establish that all sequences generated by the same cycle are in fact equivalent, we would need weaker or more generalizable sufficient conditions for equivalence.
 

### 2014-11-07 draft.tex
fixed up intro again

@@
 
-Above was an example where the blocks always have elements be either $1$ or $2$.
-Since we are self-referincing with respect to block lengths, the particular value that the block takes is not as important as the block lengths.
-Whenever the sequence can only takes two possible numbers, the values of the blocks is unambiguous; the block values simply alternate.
-As in the sequence above, after a block of 1's, the next block must be a block of 2's and vice versa.
-However if our sequence was composed of $1$'s, $2'$s and $3'$s there would be ambiguity over which number a new block should begin with.
+Above was an example where each block contains either $1$'s or $2$'s.
+Whenever the sequence can only take two possible numbers, the number in the block is unambiguous; the numbers simply alternate.
+As in the sequence above, after a block of 1's, the next block must be a block of 2's, then again 1's, and vice versa.
+However if our sequence was composed of $1$'s, $2$'s and $3$'s there would be ambiguity over which number a new block should begin with.
 
@@
 Each time we need to specify a new block, we have a choice of two elements.
-We would like to avoid having a large number of choices in the generation of the sequence, and so we will use the convention that the order in which the elements are presented gives the order they are used in the sequence.
-For example, $(123 )$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.
+We would like to avoid having a large number of choices in the generation of the sequence, and so we will give the possible elements as a cycle.
+For example, self-referencing sequences with elements from the cycle $(123 )$ contain the elements 1, 2, and 3 and will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.
 
@@
 
-We call the set the sequence is generated over, along with its order the \emph{generating cycle}.
+We call the cycle the self-referencing sequence is generated over the \emph{generating cycle}.
 
@@
 \item The self-referencing sequence for a given generating cycle is uniquely determined by its first number.
-% This doesn't really mean anything... - erik
-%
-%\item Some characterizations of sequences with the same generating cycle.
-\item For some generating cycles there exists an equivalent formulation of self-referencing sequences using substitution rules
-\item If a generating cycle has substitution rules then one can compute the limiting density for the numbers in the sequence if a limiting density exists.
+\item Some properties in the limiting behavior of self-referencing sequences generated by the same cycle.
+\item For some generating cycles there exists an equivalent formulation of self-referencing sequences using substitution rules.
+\item If a generating cycle has substitution rules then one can compute the limiting density for the numbers in the sequence if such a limiting density exists.
 \item The limiting densities of the self-referencing sequences of  $(13)$ if such densities exists.
@@
 In Section~\ref{determinism} we show that if we are given a generating cycle and a starting number for the sequence, the sequence is uniquely determined.
-In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating cycle, and not the starting number.
+In Section~\ref{equivalence}, we consider the limiting behavior of such sequences.
 
-In Section~\ref{iterativeprocess}, we show that there is an alternate way of generating them with an iterative process.
+In Section~\ref{iterativeprocess}, we give a procedure generating self-referencing sequences with an iterative process.
 In Section~\ref{substitutionrules}, we show how a large number of these sequences can actually be generated with a simple set of substitution rules.

### 2014-11-07 draft.tex
worked on density bounds for 12

@@
 
-As discussed in the Section~\ref{substitutionrules}, it is unclear how to establish a set of rules for $C = (1 2)$; however, there is some structure that we can exploit.
+As discussed in the Section~\ref{substitutionrules}, it is unclear how to establish a set of rules for $C = (1 2)$.
+However, there is some structure that we can exploit.
 In particular, we know that the sequences $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence.
 This is because such sequences would be part of a block of size at least $3$.
+There would either be a 3 earlier in the sequence, or $1,1,1$ or $2,2,2$ must be the first block.
+We know that neither option is possible.
 
-What does this mean? Well, we know that if there is a density of $1$'s and $2$'s, then that density $d_2$ is bounded between 
+What does this mean? Well, we know that if there is a limiting density of $1$'s and $2$'s, then that density $d_2$ is bounded between 
 \[ \dfrac{1}{3} \leq d_2 \leq \dfrac{2}{3} \]
 
-In general, this gives a procedure for computing the limiting density.
-For example, now we know that there must be at least one $2$ for every three numbers.
-So now we can look at how many $2$'s are in any given string of length $9$.
-We know there must be at least three $2$'s.
-There are only three cases where there are three 2's in a segment of $9$ numbers in the sequence, namely
+So now we know that there must be at least one $2$ for every three numbers.
+We can ask  the following similar question: how many $2$'s are in any consecutive substring of length $9$?
+We know there must be at least three $2$'s, but do we need more?
+There are only three cases where there are three 2's in a segment of $9$ numbers with every consecutive subsequence of length 3 containing at least one 2, namely
 \[ 2, 1, 1, 2, 1, 1 ,2 , 1, 1 \]
@@
 \[ 1, 1, 2, 1, 1, 2, 1, 1, 2 \]
-In each of these cases, we can read the subsequence that must have generated them part of them. In order, they are
+In each of these cases, we can look at the block lengths.
+Since the sequence is self-referential, the consecutive subsequence of block lengths must have appeared earlier in the sequence.
+The block lengths for the three above sequences are:
 \[ 2, 1, 2, 1, 2 \]
@@
 \[ 2, 1, 2, 1, 2 \] 
-but each of these are generated by $1, 1, 1$.
-This means that none of the sequences of length 9 are possible, so in addition to having one 2 for every group of 3, there is another 2 for every group of 9.
-Therefore, we have shown that the density must be bounded by
+We can look at the block lengths again.
+If the above segments were in the self-referencing sequence, then so is $1, 1, 1$, a contradiction.
+This means that none of the above consecutive subsequences of length 9 are possible.
+This means that in every consecutive subsequence of length $9$, there must be at least $4$ 2's.
+In other words, in addition to having one 2 for every group of 3, there is another 2 for every group of 9.
+Therefore, the density must be bounded by
 \[ \dfrac{1}{3} + \dfrac{1}{9} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} \]
-In general, we can continue this process, and we did, to show that
+In general, we can continue this process for higher powers of $3$. We did it once more for consecutive subsequences of length 27 and found you needed thirteen 2's. Thus, the density bounded by
 \[ \dfrac{1}{3} + \dfrac{1}{9} + \dfrac{1}{27} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} - \dfrac{1}{27} \]
-Of course, the number of cases to check increases very quickly: there are $3$ choices to put the $2$ in the groups of $3$, $3$ choices to put the 3 groups of 2's and 2 positions within these groups for the groups of 9, and so on.
-Its not hard to see that this number grows exponentially fast, making it computationally difficult to analyze more cases.
+The number of cases to check increases very quickly: there are $3$ choices to put the $2$ in the groups of $3$, $3$ choices to put the 3 groups of 2's and 2 positions within these groups for the groups of 9, and so on.
+Its not hard to see that this number grows exponentially fast, making it computationally difficult to bound the density.
 
-We believe that if there is a way to show this, we will be able to show that the density is $\frac{1}{2}$, since the density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number.
-However, generalizing the procedure to improve the bound seems difficult.
+If there is a pattern which we can exploit with an inductive proof, we will be able to show that the density exists and is $\frac{1}{2}$.
+The density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number.
+However, finding patterns in this procedure to improve the bound seems difficult.
 

### 2014-11-07 draft.tex
commented out conjecture

@@
 \end{definition}
-It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence, although based on numerical simulation we make the following conjecture.
-
-\begin{conj} Let $C = (1,2)$.
-A sequence $s^{(i)}$ is strongly convergent to $s(C, 2)$ if the starting sequence begins with a $2$ and contains no more than 3 elements.
-A sequence $s^{(i)}$ is strongly convergent to $s(C, 1)$ if the starting sequence begins with a $1$ and contains between 2 and 3 elements.
+It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence.%, although based on numerical simulation we make the following conjecture. 
+%
+%\begin{conj} Let $C = (1,2)$.
 A sequence $s^{(i)}$ is weakly convergent but not strongly convergent to $s(C, a)$ if the starting sequence begins with an $a \in \{1,2\}$ and contains 4 or more elements.
-\end{conj} 
-We make this conjecture on the basis on computer simulations which executes the iterative process to at least 20 iterations.
+%\end{conj} 
+%We make this conjecture on the basis on computer simulations which executes the iterative process to at least 20 iterations.
 More iterations were not completed because after this point the simulations began to take impractically long amounts of time to complete.
@@
 In other words, any sequence that uses a certain generating cycle is equivalent to any other sequence created by the same generating cycle.
-\item In the same section, we propose a condition for determining when two sequences are not equivalent.
+%\item In the same section, we propose a condition for determining when two sequences are not equivalent.
 We conjecture that if two sequences have unequal numbers with the same read distances, then they are not equivalent.
 It seems from trying cases that this condition is sufficient, but we have been unable to prove it.
-\item In Section~\ref{iterativeprocess} we make a conjecture on when sequences are strongly and weakly convergent.
+\item In Section~\ref{iterativeprocess} we explored when sequences are strongly and weakly convergent.
 We have not yet been able to investigate strong and weak convergence in great detail, as we have been unable to make precise mathematical statements about these types of convergence.

### 2014-11-07 draft.tex
Edits to the Introduction

@@
 
-We will study a sequences which are self-referencing.
-That is, each number in the sequence will correspond to a previous number in the same sequence.
+In this paper we study sequences which are self-referencing.
+That is, the numbers in the sequence in some way correspond to previous numbers in the same sequence.
 In particular, we will analyze self-referencing sequences with respect to block lengths.
@@
 
-Above was an example where each block contains either $1$'s or $2$'s.
-Whenever the sequence can only take two possible numbers, the number in the block is unambiguous; the numbers simply alternate.
-As in the sequence above, after a block of 1's, the next block must be a block of 2's, then again 1's, and vice versa.
+In our example above each block is composed of either $1$'s or $2$'s.
+When a self referential sequence can only take two possible numbers the number a new block should begin with is unambiguous; the numbers simply alternate.
+For instance in our example above after a block of 1's, the next block must be a block of 2's, then again 1's, and so on.
 However if our sequence was composed of $1$'s, $2$'s and $3$'s there would be ambiguity over which number a new block should begin with.
@@
 \end{equation}
-are in some sense self-referencing sequences.
+are self-referential sequences.
 Each time we need to specify a new block, we have a choice of two elements.
-We would like to avoid having a large number of choices in the generation of the sequence, and so we will give the possible elements as a cycle.
-For example, self-referencing sequences with elements from the cycle $(123 )$ contain the elements 1, 2, and 3 and will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.
+We would like to avoid having a large number of choices in the generation of the sequence, and so we will give an order to the elements a sequence will contain, and write it as a cycle.
+For example, self-referencing sequences with elements from the cycle $(123 )$ contain the elements 1, 2, and 3 and will have the order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.
 
-With this rule, the self-referencing sequence generated by $(123)$ that begins with a $1$ is shown in Figure~\ref{fig:onetwothree}.
+With this rule, the self-referencing sequence generated by $(123)$ that begins with a $1$ is shown in Figure 2 %~\ref{fig:onetwothree}.
 \begin{figure}
@@
 \includegraphics[width=0.4\linewidth]{sequenceBlocks123}
-\caption{Sequence generated by $(123)$ broken up into contiguous blocks.}
+\caption{Sequence generated by $(123)$ broken up into contiguous blocks with lengths generating same sequence.}
 \end{figure}
 
-We call the cycle the self-referencing sequence is generated over the \emph{generating cycle}.
+We call the cycle the self-referential sequence is generated over the \emph{generating cycle}.
 
@@
 \item The self-referencing sequence for a given generating cycle is uniquely determined by its first number.
-\item Some properties in the limiting behavior of self-referencing sequences generated by the same cycle.
+\item In order to establish long term equivalence of two sequences it is enough to check only finitely many terms 
 \item For some generating cycles there exists an equivalent formulation of self-referencing sequences using substitution rules.
@@
 In Section~\ref{determinism} we show that if we are given a generating cycle and a starting number for the sequence, the sequence is uniquely determined.
-In Section~\ref{equivalence}, we consider the limiting behavior of such sequences.
+In Section~\ref{equivalence}, we consider the limiting behavior of such sequences and show when the limiting behavior of two sequences is the same.
 
-In Section~\ref{iterativeprocess}, we give a procedure generating self-referencing sequences with an iterative process.
+In Section~\ref{iterativeprocess}, we give an iterative process that generates self-referencing sequences.
 In Section~\ref{substitutionrules}, we show how a large number of these sequences can actually be generated with a simple set of substitution rules.
-We introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating cycles.
-Unfortunately, we find that the open problem of determining the density of the sequences generated by $(12)$ is hard to solve using our methods.
+We introduce the density problem in Section~\ref{densitycomputation}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating cycles.
+In Section ~\ref{densitybounds}  we consider the open problem of determining the density of the sequences generated by $(12)$ and find it hard to solve using our methods.
 

### 2014-11-07 draft.tex
Edits to Iterative Process

@@
 
-\emph{Written by Fermi Ma, edited by Perry Kleinhenz and Erik Waingarten}
+\emph{Written by Fermi Ma and Perry Kleinhenz edited by Erik Waingarten}
 
@@
 \end{align*}
-Let $s^\infty$ be the limit of this process. 
-By limit, we mean the limit with respect to each term seperately, so we can define
-\[ s^{(i)}_j \rightarrow s^\infty_j \]
-as $i \rightarrow \infty$. We claim that $s^\infty$ is the unique self-referential sequence $s(C,2)$. In fact we can prove a more general result about this iterative process.
+We define the limit of this process, the sequence $s^{\infty}$ as 
+\begin{equation*}
+s_{j}^{\infty} = \lim_{i \rightarrow \infty} s_{j}^{(i)}.
+\end{equation*}
+We claim that this $s^{\infty}$ exists and is the unique self-referential sequence $s(C,2)$. In fact we can prove a more general result about this iterative process.
+
+Before we do so however we will clarify two different notions of convergence, that will allow us to state our result more precisely. 
+\begin{definition} We say that a sequence $s^{(i)}$ is weakly convergent to $s(C,a)$ if $s^{\infty}$ exists and $s^{\infty} = s(C,a)$.
+We say that the sequence $s^{(i)}$ is strongly convergent to $s(C,a)$ if $s^{(i)}$ is a prefix of $s(C,a)$ for all $i \geq k$ for some $k\geq 0$.
+\end{definition}
+Note that all strongly convergent sequences are also weakly convergent.
 
@@
 Let $s^0 = a$ for some $a \in C$ where $a \neq 1$, and let $s^{(i)}$ be the sequences generated by the iterative process.
-Then the limit of this process $s^\infty$ exists and is the sequence $s(C,a)$.
+Then $s^{(i})$ is strongly convergent to $s(C,a)$.
 \end{theorem}
@@
 For the inductive step, we observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence.
-We know that the sequence it describes is longer because the sequence it describes has length equal to the sum of the original prefix.
-In other words, the length of $s^{(i+1)}$ equals the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1).
+We know that the sequence it describes is longer because the length of $s^{(i+1)}$ equals the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1).
 
-The resultant sequence is a prefix of the self referencing sequence because if we apply our process to any prefix of the self referencing sequence it will produce a prefix of the self referencing sequence.
+%The resultant sequence is a prefix of the self referencing sequence because if we apply our process to any prefix of the self referencing sequence it will produce a prefix of the self referencing sequence.
+
+Since very $s^{(i)}$ is a prefix of $s(C,a)$, we know that once $s^{(i)}_{j}$ has been assigned a value, $s^{(k)}_{j}$ will have the same value for all $k \geq i$.
+Because of this $\lim_{i\rightarrow \infty} s^{(i)}_{j}$ is well defined and so $s^{\infty}$ exists.
+As the limit of the process we know that $s^{\infty}$ is a prefix of $s(C,a)$, but since at every step of the process the length of the sequence strictly increases we must have that $s^{\infty}$ has infinite length, and so $s^{\infty} = s(C,a)$.
+
 \end{proof}
 
-Note that we must specify that $a \neq 1$, because that is the only case in which the length of $s^{(i+1)}$ is not strictly greater than the length of $s^{(i)}$.
+Note that we must specify that $a \neq 1$, otherwise $s^{(i)}$ is always $1$, and so the length of the sequence does not grow.
 
-It turns out that we can be less restrictive about what our starting sequence, $s^{(0)}$, is.
-Going back to the case of $C = (1 2)$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2.
-And still, the limiting sequence will be the self-referential sequence.
+Above we only considered starting sequences which have length one,  but we can be less restrictive about what our starting sequence, $s^{(0)}$, is.
+Going back to the case of $C = (1 2)$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2 and the limiting sequence will still be the self-referential sequence.
 For example, suppose $s^{(0)} = 2,1,1,1,2$. The rules for generating $s^{(i)}$ give:
@@
 We note that $s^\infty = s(C,a)$ because because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(C,a)$ in the limit.
-We call the rest of each $s^{(i)}$ that is not a prefix of $s(C,2)$ the \emph{tail}.
+This is true in general.
 
-We now consider two different forms of convergence.
-We know that $s^{(i)}$ in some sense approaches $s(C,a)$, as arbitrarily many elements at the beginning of the sequence will match up with the beginning of $s(C,a)$ as $i$ increases.
-However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(C,a)$.
-Thus, we formalize two types of convergence
-\begin{definition} We say that a sequence $s^{i}$ is weakly convergent to $s(C,a)$ if $s^\infty = s(C,a)$.
-We say that the sequence $s^{i}$ is strongly convergent if $s^{(i)}$ is a prefix of $s(C,a)$ for all $i \geq k$ for some $k$.
-\end{definition}
-It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence.%, although based on numerical simulation we make the following conjecture. 
+\begin{corollary}
+Let $C$ be some generating cycle, let $s^(0)$ be any sequence of elements of $C$ which begins with some $a \in C$ such that $a \neq 1$. Then $s^{(i)}$ is weakly convergent to $s(C,a)$. 
+\end{corollary}
+\begin{proof} 
+\end{proof}
+
+It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence, although based on numerical simulation we suspect that applying the process to the starting sequence $2222$
 %
@@
 \label{densitybounds}
+\emph{Written by Erik Waingarten, edited by Fermi Ma and Perry Kleinhenz}
 
@@
 The major difficulty here seems to come in proving that a certain generating cycle does \emph{not} allow for substitution rules.
-\item In Section~\ref{density} we were able to compute the density of any sequence with generating cycle $(13)$, and we were able to give bounds on the density of any sequence with generating cycle $(12)$.
-However, we leave unresolved the question of what the density for $(12)$ actually is, or even if we can obtain tighter bounds.
+\item In Section~\ref{densitycomputation} we were able to compute the density of any sequence with generating cycle $(13)$, and in Section ~\ref{densitybounds} were able to give bounds on the density of any sequence with generating cycle $(12)$.
+However, we leave unresolved the question of what the density for $(12)$ actually is, or if we can obtain tighter bounds.
 \end{itemize}

### 2014-11-07 draft.tex
Edits to Iterative Process

@@
 Note that all strongly convergent sequences are also weakly convergent.
+We now can show that if we specify a generating cycle any one element starting sequence (other than $1$) is strongly convergent to the self referencing sequence that begins with the element.
 
 \begin{theorem}
+\label{iteratetheorem}
 Let $C$ be some generating cycle.
-Let $s^0 = a$ for some $a \in C$ where $a \neq 1$, and let $s^{(i)}$ be the sequences generated by the iterative process.
-Then $s^{(i})$ is strongly convergent to $s(C,a)$.
+Let $s^{(0)} = a$ for some $a \in C$ where $a \neq 1$, and let $s^{(i)}$ be the sequences generated by the iterative process.
+Then $s^{(i)}$ is strongly convergent to $s(C,a)$.
 \end{theorem}
@@
 For the inductive step, we observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence.
-We know that the sequence it describes is longer because the length of $s^{(i+1)}$ equals the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1).
+We know that the sequence it describes is longer because the length of $s^{(i+1)}$ equals the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$.
 
@@
 Since very $s^{(i)}$ is a prefix of $s(C,a)$, we know that once $s^{(i)}_{j}$ has been assigned a value, $s^{(k)}_{j}$ will have the same value for all $k \geq i$.
-Because of this $\lim_{i\rightarrow \infty} s^{(i)}_{j}$ is well defined and so $s^{\infty}$ exists.
+Because of this the limit is well defined and so $s^{\infty}$ exists.
 As the limit of the process we know that $s^{\infty}$ is a prefix of $s(C,a)$, but since at every step of the process the length of the sequence strictly increases we must have that $s^{\infty}$ has infinite length, and so $s^{\infty} = s(C,a)$.
-
 \end{proof}
@@
 
-Above we only considered starting sequences which have length one,  but we can be less restrictive about what our starting sequence, $s^{(0)}$, is.
+Above we only considered starting sequences which have length one,  but we would like to consider arbitrary starting sequences.
 Going back to the case of $C = (1 2)$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2 and the limiting sequence will still be the self-referential sequence.
@@
 Here, we bold the numbers that form a prefix of $s(C,2)$.
-We note that $s^\infty = s(C,a)$ because because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(C,a)$ in the limit.
+We note that $s^\infty = s(C,a)$ because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(C,a)$ in the limit.
 This is true in general.
@@
 \begin{corollary}
-Let $C$ be some generating cycle, let $s^(0)$ be any sequence of elements of $C$ which begins with some $a \in C$ such that $a \neq 1$. Then $s^{(i)}$ is weakly convergent to $s(C,a)$. 
+Let $C$ be some generating cycle, let $s^{(0)}$ be any sequence of elements of $C$ which begins with some $a \in C$ such that $a \neq 1$. Then $s^{(i)}$ is weakly convergent to $s(C,a)$. 
 \end{corollary}
 \begin{proof} 
+This follows directly from Theorem \ref{iteratetheorem}.
+That is because $s^{(0)}$ begins with an $a$, at each step of the process $s^{(i)}$ will be a prefix of $s(C,a)$ followed by some terms that disagree with $s(C,a)$.
+In the limit the prefix of $s(C,a)$ becomes $s(C,a)$ and so $s^{\infty} = s(C,a)$.
 \end{proof}
+Note that we do not guarantee strong convergence as there are some starting sequences, such as $2222$ which in numerical simulation appears to produce sequences which disagree with the $s( (12), 2)$ on a constant fraction of indices.
+We would like to be able to produce a general result that distinguishes between strongly and weekly convergent subsequences but it is unclear that there is a nice categorization of them.
 
-It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence, although based on numerical simulation we suspect that applying the process to the starting sequence $2222$
 %

### 2014-11-07 draft.tex
edited rules section

@@
 
+\begin{comment}
 Why does this work?
-Well, first of all, notice that the rules are the natural applications of the self-referential sequence.
+Well, first of all, notice that each rule of the form $x \rightarrow y$ has the .
 Also, note that we have decided to make the rules of length 2; if we had made the self-referential rules $1 \rightarrow 1$ and $3 \rightarrow 333$, this would not have worked.
@@
 This is so that at the start of each rule, we always know which numbers to write.
+\end{comment}
 
-Therefore, we can formulate a set of sufficient conditions to make rules to generate the self-referential sequence with generating cycle $C$.
+We can formulate a set of sufficient conditions for rules that generate the self-referencing sequences.
+Formally, we will set $R$ to be a rule set, which is a set of strings of the form $x \rightarrow y$.
 
-We want to have a set of rules $R$ where each rule $r \in R$ has length $|r| = l$, such that all $|C|^{l}$ strings are in the rule set and for each $r \in R$
-\begin{itemize}
-\item each rule writes blocks in the same order of $C$ and each rule has the same starting number
-\item $\sum_{i \in r} i$ is divisible by $l$
-\item $l$ is divisible by $|C|$
-\end{itemize}
-Given these conditions on every rule, we can guarantee that the self-referential sequence is generated.
-Note that in this case, the above example does not satisfy our rules until we say that $1, 3 \to 3, 1, 1, 1$.
+\begin{proposition}
+\label{ruleprop}
+Suppose $C$ is a cycle.
+For any $l \in \mathbb{N}$ divisible by $|C|$ and $a \in C$, let $R$ be a rule set.
+If for each string $s$ of elements in $C$ of length $l$, there exists a unique rule $x \rightarrow y$ in $R$ with $x = s$ and:
+\begin{enumerate}
+\item $x$ reads the block lengths of $y$ from left to right
+\item $y$ starts with $a$
+\item $|y|$ is divisible by $l$
+\end{enumerate}
+Then sucessively applying the rules in $R$, starting with a sequence of length $l$ generates a sequence which is weakly convergent to $s(C, a)$.
+\end{proposition}
 
-Another example, we can have $C = (2 4)$, and we can have that the rules be
+\begin{proof}
+Let $s^{(i)}$ be the sequence generated from the process from Section~\ref{iterativeprocess}.
+I will show by induction on $i$ that $s^{(i)}$ is the result of applying the rules in $R$ $i$ times to a starting sequence $s^{(0)}$ of length $l$.
+By letting the starting sequence be $s^{(0)}$, we know it works for $i = 0$.
+
+Suppose that applying the rules in $R$ to $s^{(0)}$ $i$ times gives $s^{(i)}$.
+Note that $|s^{(i)}|$ is divisible by $l$ since each iteration applies the rule to a segment of length $l$ and by condition 3, it writes a sequence of length a multiple of $l$.
+So we can partition $s^{(i)}$ uniquely into disjoint chunks of size $l$.
+
+In the iterative process, each chunks of size $l$ will generate $l$ blocks, and since $|C|$ divides $l$, the block referenced by each of these chunks will have the same number.
+If the rules begin with $a$, then this number will be $a$.
+
+So applying the rules to each chunk is equivalent to applying the iterative process, where we  group elements into chunks of size $l$.
+Therefore, $s^{(i+1)}$ is the result of applying the rules to $s^{(i)}$.
+\end{proof}
+
+Given Proposition~\ref{ruleprop} we can generate rules a large class of generating cycles $C$. Take the following example, where $C = (2 4)$, and the rules are
 \begin{align*}
@@
 \end{align*}
+These rules satisfy the conditions of Proposition~\ref{ruleprop}, and so successive application of these rules will generate a self-referencing sequence which is weakly convergent to $s(C, 2)$.
 
-One of the reasons why generating a rule set for $(1 2)$ is complicated is that since $1, ..., 1$ must be in the rule set, the length of the rules must be even (since $|C| = 2$).
-On the other hand, $2, 1, ..., 1$ must also be in the set and so the rule length must be odd.
+\begin{proposition}
+\label{norules12}
+There is no set of rules $R$ satisfying the conditions of Proposition~\ref{ruleprop} for $C=(12)$.
+\end{proposition}
+
+\begin{proof}
+Suppose $l$ is a positive integer.
+Then $l$ must be even.
+Let $s$ be the string with one 2 and $l-1$ 1's.
+Suppose $x \rightarrow y$ is the rule for $s$, so $x = s$.
+Then by condition 1 of Proposition~\ref{ruleprop}, $|y| = l+1$, and by condition 3, $l = 1$.
+Which contradicts the fact that $l$ is divisible by $|C| = 2$.
+\end{proof}
 

### 2014-11-07 draft.tex
density bounds for c = 12 finished

@@
 In particular, we know that the sequences $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence.
-This is because such sequences would be part of a block of size at least $3$.
-There would either be a 3 earlier in the sequence, or $1,1,1$ or $2,2,2$ must be the first block.
-We know that neither option is possible.
+This is because such sequences either imply the existence of a 3 earlier in the sequence, or make up the first block of the sequence.
+Neither option is possible.
 
-What does this mean? Well, we know that if there is a limiting density of $1$'s and $2$'s, then that density $d_2$ is bounded between 
+This simple observation implies that if there is a limiting density of $1$'s and $2$'s, then that density $d_2$ is bounded between 
 \[ \dfrac{1}{3} \leq d_2 \leq \dfrac{2}{3} \]
 
-So now we know that there must be at least one $2$ for every three numbers.
+So we know there must be at least one $2$ for every three consecutive terms of the sequence.
 We can ask  the following similar question: how many $2$'s are in any consecutive substring of length $9$?
-We know there must be at least three $2$'s, but do we need more?
-There are only three cases where there are three 2's in a segment of $9$ numbers with every consecutive subsequence of length 3 containing at least one 2, namely
+The answer is (trivially) at least 3, by our original observation.
+However, there are only three cases where there are exactly three 2's in a segment of $9$ numbers.
+In particular, they are
 \[ 2, 1, 1, 2, 1, 1 ,2 , 1, 1 \]
 \[ 1, 2, 1, 1, 2, 1, 1, 2, 1 \]
-\[ 1, 1, 2, 1, 1, 2, 1, 1, 2 \]
-In each of these cases, we can look at the block lengths.
+\[ 1, 1, 2, 1, 1, 2, 1, 1, 2 .\]
+Note that we rule out all strings of 9 numbers with a consecutive run of three 1's or three 2's.
+
+In each of the above cases, we can look at the block lengths.
 Since the sequence is self-referential, the consecutive subsequence of block lengths must have appeared earlier in the sequence.
@@
 \[ 1, 2, 1, 2, 1\]
-\[ 2, 1, 2, 1, 2 \] 
-We can look at the block lengths again.
+\[ 2, 1, 2, 1, 2 .\] 
+We repeat this reasoning.
 If the above segments were in the self-referencing sequence, then so is $1, 1, 1$, a contradiction.
-This means that none of the above consecutive subsequences of length 9 are possible.
-This means that in every consecutive subsequence of length $9$, there must be at least $4$ 2's.
+This means that none of the above consecutive subsequences of length 9 are possible, which implies that in every consecutive subsequence of length $9$, there must be at least four 2's.
 In other words, in addition to having one 2 for every group of 3, there is another 2 for every group of 9.
@@
 \[ \dfrac{1}{3} + \dfrac{1}{9} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} \]
-In general, we can continue this process for higher powers of $3$. We did it once more for consecutive subsequences of length 27 and found you needed thirteen 2's. Thus, the density bounded by
+We can try to continue this process for higher powers of $3$.
+For subsequences of length 27, our result for subsequences of length 9 tells us that we must have at least twelve 2's.
+But we were able to show that a contradictino is reached in each of these cases, similar to the way a contradiction is reached above.
+So for consecutive subsequences of length 27, there must be at least thirteen 2's.
+Thus, the density bounded by
 \[ \dfrac{1}{3} + \dfrac{1}{9} + \dfrac{1}{27} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} - \dfrac{1}{27} \]
-The number of cases to check increases very quickly: there are $3$ choices to put the $2$ in the groups of $3$, $3$ choices to put the 3 groups of 2's and 2 positions within these groups for the groups of 9, and so on.
-Its not hard to see that this number grows exponentially fast, making it computationally difficult to bound the density.
+Unfortunately, we have not been able to generalize this reasoning, as the number of cases that must be checked increases exponentially fast.
+There are $3$ choices for where to put the $2$ in the groups of $3$, $3$ choices for where to put the 3 groups of 2's and 2 positions within these groups for the groups of 9, and so on.
 
-If there is a pattern which we can exploit with an inductive proof, we will be able to show that the density exists and is $\frac{1}{2}$.
-The density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number.
-However, finding patterns in this procedure to improve the bound seems difficult.
+If there exists a pattern that we can make general statements about for arbitrary powers of 3, then this will imply that the density is bounded below and above by
+\[ \sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2} \]
+Note that this would also imply the existence of a density.
+The previous claims can only hope to bound the density, if it exists.
+The issue is that the ``limiting" density may not exist as it may oscillate between the two bounds.
+But if the bound is tight (upper and lower bounded at $1/2$) then there is no room for oscillation, and so the density would exist.
 

### 2014-11-08 draft.tex
new poem

@@
 We were able to support these calcuations with numerical simulations.
-We generated the $s((13),3)$ sequence up to 100,000 terms, and we found that the fraction of 1's was approximately $0.398$, and that the fraction of 3's was approximately 0.602.
+We generated the $s((13),3)$ sequence up to 100,000 terms, and we found that the fraction of 1's was approximately $0.397$, and that the fraction of 3's was approximately 0.603.
 
@@
 
-\section{Haikus}
-\emph{Written by Fermi Ma}\\
+\section{Poetry}
+\emph{Written by Fermi Ma and Erik Waingarten}\\
 
-\noindent One, Two, Two, One, One,\\
-Two, One, Two, Two, One, Two, Two,\\
-One, One, Two, One, One...\\
+\noindent A sequence referring to its past\\
+seems quite simple --- not so fast!\\
+Sure, it is deterministic,\\
+but tell me the density statistic.\\
 
-\noindent For the final draft,\\
-we will improve the haikus\\
-(and the whole paper).
+\noindent Well, try (one, three) and you will see\\
+the density* is point six-oh-three.\\
+Just write the substitution rule.\\
+To miss the answer, you'd be a fool.\\
 
-\subsection{Poem}
+\noindent Now do the same for one and two,\\
+Look at the rule and now you ---\\
+Hold on, this doesn't work for numbers of different parity.\\
 
-\noindent A sequence refering to its past\\
-warrants many questions to be asked.\\
-These sequences are deterministic \\
-and their read distances an imporant statistic.
-\\
-
-\noindent We want to compute the limiting density\\
-so we've defined a process.\\
-But one would be a fool to not follow our rule.\\
-And compute the density for 1, 3.\\
-
-\noindent The big question remains unanswered.\\
-The density of 1,2 which seems to be one half.\\
-Our rules are for numbers of a certain class.\\
-Unfortunetly, dealing with 1,2 is a pain in the a**.\\
+*if it exists
 

### 2014-11-08 draft.tex
modified open problems section

@@
 We were able to support these calcuations with numerical simulations.
-We generated the $s((13),3)$ sequence up to 100,000 terms, and we found that the fraction of 1's was approximately $0.397$, and that the fraction of 3's was approximately 0.603.
+We generated the $s((13),3)$ sequence using the iterative method to 20 iterations, and we found that the fraction of 1's was approximately $0.397$, and that the fraction of 3's was approximately 0.603.
 
@@
 
-Our work leaves a number of unresolved problems.
+If we had more time to work on this project, we would investigate the following unresolved questions.
 
 \begin{itemize}
-\item In Section~\ref{equivalence}, we propose the idea that in some sense, the number that we start a sequence with does not matter.
-In other words, any sequence that uses a certain generating cycle is equivalent to any other sequence created by the same generating cycle.
-%\item In the same section, we propose a condition for determining when two sequences are not equivalent.
-We conjecture that if two sequences have unequal numbers with the same read distances, then they are not equivalent.
-It seems from trying cases that this condition is sufficient, but we have been unable to prove it.
-\item In Section~\ref{iterativeprocess} we explored when sequences are strongly and weakly convergent.
-We have not yet been able to investigate strong and weak convergence in great detail, as we have been unable to make precise mathematical statements about these types of convergence.
-Most of what we know about them comes from analyzing numerical data.
+\item In Section~\ref{equivalence}, we considered the idea that perhaps the number that we start a sequence with does not matter.
+In other words, it may be true that any sequence that uses a certain generating cycle is equivalent to any other sequence that uses the same generating cycle.
+Alternatively, it may be the case that the same generating cycle may yield two sequences that are not equivalent.
+We would like to know which is actually true.
 \item In Section~\ref{substitutionrules}, we look at substitution rules for certain ``nice" generating cycles of numbers, such as $(13)$ and $(24)$.
-These sets allow for substitution rules because they satisfy the trio of properties we outline in that section.
-However, we do not know that these are the only properties.
-We leave unresolved the question of determining precisely which sets allow for substitution rules and which ones ones do not.
+However, we leave unresolved the question of determining precisely which sets allow for substitution rules and which ones ones do not.
 The major difficulty here seems to come in proving that a certain generating cycle does \emph{not} allow for substitution rules.
-\item In Section~\ref{densitycomputation} we were able to compute the density of any sequence with generating cycle $(13)$, and in Section ~\ref{densitybounds} were able to give bounds on the density of any sequence with generating cycle $(12)$.
-However, we leave unresolved the question of what the density for $(12)$ actually is, or if we can obtain tighter bounds.
+It seems very hard to come up with a set of substitution rules for $(12)$ and a proof for why this is impossible would be interesting.
+\item In Section~\ref{densitybounds} we were able to give certain bounds on the density, if it exists, for the original self-referential sequence of 1's and 2's.
+Given more time, we would want to see if we could extend our reasoning to the case of length 81 sequences, and perhaps even further.
+This would, of course, require a more clever way of proving our claims than the way we have used so far.
 \end{itemize}
@@
 Look at the rule and now you ---\\
-Hold on, this doesn't work for numbers of different parity.\\
+hold on, this doesn't work for numbers of different parity.\\
 

### 2014-11-08 draft.tex
heavily modified substitution rules section

@@
 
-Before we do so however we will clarify two different notions of convergence, that will allow us to state our result more precisely. 
+Before we do so however we will give two different notions of convergence that will allow us to state our result more precisely. 
 \begin{definition} We say that a sequence $s^{(i)}$ is weakly convergent to $s(C,a)$ if $s^{\infty}$ exists and $s^{\infty} = s(C,a)$.
@@
 Note that all strongly convergent sequences are also weakly convergent.
-We now can show that if we specify a generating cycle any one element starting sequence (other than $1$) is strongly convergent to the self referencing sequence that begins with the element.
+We can now show that if we specify a generating cycle, any length 1 starting sequence $s^0$ (other than $1$) is strongly convergent to the self referencing sequence that begins with the element.
 
-\begin{theorem}
+\begin{proposition}
 \label{iteratetheorem}
@@
 Then $s^{(i)}$ is strongly convergent to $s(C,a)$.
-\end{theorem}
+\end{proposition}
 
 \begin{proof}
-We prove this by induction on $i$.
+This claim is straightforward, but we will formally prove it with induction on $i$.
 We show that if $s^{(i)}$ is a prefix of $s(C,a)$, then $s^{(i+1)}$ is a longer prefix of $s(C,a)$.
@@
 We know that the sequence it describes is longer because the length of $s^{(i+1)}$ equals the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$.
-
-%The resultant sequence is a prefix of the self referencing sequence because if we apply our process to any prefix of the self referencing sequence it will produce a prefix of the self referencing sequence.
-
-Since very $s^{(i)}$ is a prefix of $s(C,a)$, we know that once $s^{(i)}_{j}$ has been assigned a value, $s^{(k)}_{j}$ will have the same value for all $k \geq i$.
+We know that the sequence it describes is a prefix, since the first number matches up with $s(C,a)$, and so do the block lengths.
+This completes the induction.
+%
+%%The resultant sequence is a prefix of the self referencing sequence because if we apply our process to any prefix of the self referencing sequence it will produce a prefix of the self referencing sequence.
+%
+%Since very $s^{(i)}$ is a prefix of $s(C,a)$, we know that once $s^{(i)}_{j}$ has been assigned a value, $s^{(k)}_{j}$ will have the same value for all $k \geq i$.
 Because of this the limit is well defined and so $s^{\infty}$ exists.
@@
 
-Note that we must specify that $a \neq 1$, otherwise $s^{(i)}$ is always $1$, and so the length of the sequence does not grow.
+Note that we must specify that $a \neq 1$, because otherwise $s^{(i)}$ is always $1$, and so the length of the sequence does not grow.
 
@@
 \begin{align*}
-s^{(0)}&= \mathbf{2},1,1,1,2\\
-s^{(1)} &= \mathbf{2,2,1},2,1,2,2\\
-s^{(2)} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
-s^{(3)} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
+s^{(0)}&= 2,1,1,1,2\\
+s^{(1)} &= 2,2,1,2,1,2,2\\
+s^{(2)} &= 2,2,1,1,2,1,1,2,1,1,2,2\\
+s^{(3)} &= 2,2,1,1,2,1,2,2,1,2,1,1,2,1,2,2,1,1\\
 &\vdots
 \end{align*}
-Here, we bold the numbers that form a prefix of $s(C,2)$.
-We note that $s^\infty = s(C,a)$ because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(C,a)$ in the limit.
-This is true in general.
+We claim that $s^\infty = s(C,a)$, which we state as the following corollary.
 
@@
 \begin{proof} 
-This follows directly from Theorem \ref{iteratetheorem}.
+This follows directly from Proposition \ref{iteratetheorem}.
 That is because $s^{(0)}$ begins with an $a$, at each step of the process $s^{(i)}$ will be a prefix of $s(C,a)$ followed by some terms that disagree with $s(C,a)$.
@@
 \end{proof}
-Note that we do not guarantee strong convergence as there are some starting sequences, such as $2222$ which in numerical simulation appears to produce sequences which disagree with the $s( (12), 2)$ on a constant fraction of indices.
+Note that we do not guarantee strong convergence as there are some starting sequences, such as $2222$, which in numerical simulation appears to produce sequences which disagree with the $s( (12), 2)$ on a constant fraction of indices.
 We would like to be able to produce a general result that distinguishes between strongly and weekly convergent subsequences but it is unclear that there is a nice categorization of them.
@@
 The iterative process given in Section~\ref{iterativeprocess} is one way to generate self-referencing sequences.
-However, for certain generating cycles $C$, there is a simpler ways to perform this iteration using a fixed set of substitution rules.
+However, for certain generating cycles $C$, there is a simpler way to perform this iteration using a fixed set of substitution rules.
 Take $C = (1 3)$ as an example, and consider the sequence that starts with a 3:
@@
 \begin{align*}
-\text{Rule }A: 3,3 &\to 3,3,3,1,1,1 \\
-\text{Rule }B: 3,1 &\to 3,3,3,1\\
-\text{Rule }C: 1,1 &\to 3,1
+\text{Rule }1: 3,3 &\to 3,3,3,1,1,1 \\
+\text{Rule }2: 3,1 &\to 3,3,3,1\\
+\text{Rule }3: 1,1 &\to 3,1
 \end{align*}
 This is to be interpreted as follows.
-Starting from 3,3, the only possible rule to apply is Rule A, which gives $3,3,3,1,1,1$ at the following iteration.
-We then apply Rule A to $3,3$, Rule B to $3,1$ and Rule C to $1,1$.
+Starting from 3,3, the only possible rule to apply is Rule 1, which gives $3,3,3,1,1,1$ at the following iteration.
+We then apply Rule 1 to $3,3$, Rule 2 to $3,1$ and Rule 3 to $1,1$.
 Note that at every step, we break up the sequence into chunks of 2 and then simultaneously apply \emph{all} the rules.
@@
 
-\begin{comment}
-Why does this work?
-Well, first of all, notice that each rule of the form $x \rightarrow y$ has the .
-Also, note that we have decided to make the rules of length 2; if we had made the self-referential rules $1 \rightarrow 1$ and $3 \rightarrow 333$, this would not have worked.
-In addition, we know that each block length preserves parity.
+We can formulate a set of sufficient conditions for rules that generate the self-referencing sequences.
 
-We would like to know if we can do this for any generating cycle $C$ and starting number $a$. In particular we would like to be able to simulate our iterative process with substitution rules.
-
-When we apply our substitution rules we should not need to know what happened to blocks other than the one we are currently substituting for.
-Therefore, each rule must somehow encode the length of  blocks, the number of elements already written in the current block, and what number the current block is composed of.
-The first condition is simple to implement, as the rules will specify the length of blocks.
-In order to keep track of the numbers already written, each rule should be self-contained.
-That is each rule should write complete blocks.
-In order to comply with the third condition it is enough for the size of the generating cycle to divide the length of the rules.
-This is so that at the start of each rule, we always know which numbers to write.
-\end{comment}
-
-We can formulate a set of sufficient conditions for rules that generate the self-referencing sequences.
-Formally, we will set $R$ to be a rule set, which is a set of strings of the form $x \rightarrow y$.
+\textbf{Reader's Note:} The following formalism might be hard to follow, but the general idea of the formalism states that we should do what we did in the $C = (13)$ case.
+In that example, we wrote down all possible rules that transform strings of a certain length (2), such that each rule was self referential (the input to the rule specifies the block lengths of the output to the rule).
 
@@
 \label{ruleprop}
-Suppose $C$ is a cycle.
-For any $l \in \mathbb{N}$ divisible by $|C|$ and $a \in C$, let $R$ be a rule set.
-If for each string $s$ of elements in $C$ of length $l$, there exists a unique rule $x \rightarrow y$ in $R$ with $x = s$ and:
+Suppose $C$ is a generating cycle.
+Fix a length $l$ that is divisible by $|C|$ as well as some $a \in C$.
+Then for all possible substrings $x$ (using numbers in $C$) of length $l$, write down the unique rule, if it exists, of the form $x \to y$ that satisfies the following properties:
 \begin{enumerate}
-\item $x$ reads the block lengths of $y$ from left to right
-\item $y$ starts with $a$
-\item $|y|$ is divisible by $l$
+\item $x$ is the sequence of the block lengths of $y$.
+\item $y$ starts $a$
+\item $|y|$ is divisible by $l$.
 \end{enumerate}
-Then sucessively applying the rules in $R$, starting with a sequence of length $l$ generates a sequence which is weakly convergent to $s(C, a)$.
+If any string $x$ does not have a corresponding $y$ that satisfies the above properties, then the rule set does not work.
+If we can write down a rule for each $x$, then successively applying these rules to a starting sequence length $l$ will give sequences that weakly converge to $s(C,a)$.
 \end{proposition}
@@
 \begin{proof}
-Let $s^{(i)}$ be the sequence generated from the process from Section~\ref{iterativeprocess}.
-I will show by induction on $i$ that $s^{(i)}$ is the result of applying the rules in $R$ $i$ times to a starting sequence $s^{(0)}$ of length $l$.
-By letting the starting sequence be $s^{(0)}$, we know it works for $i = 0$.
+Let $s^{(i)}$ be the sequence generated by the iterative process from Section~\ref{iterativeprocess} given some starting sequence $s^{(0)}$ of length $l$.
+We show by induction on $i$ that applying the above rules to the same starting sequence simulates the iterative process.
 
-Suppose that applying the rules in $R$ to $s^{(0)}$ $i$ times gives $s^{(i)}$.
+The base case is trivially satisfied, since we start with the same starting sequence $s^{(0)}$ in both the iterative process and this substitution process.
+
+For the inductive step, suppose that applying the rules in $i$ times to $s^{(0)}$ gives $s^{(i)}$, the sequence given at the $i$th step of the iterative process.
 Note that $|s^{(i)}|$ is divisible by $l$ since each iteration applies the rule to a segment of length $l$ and by condition 3, it writes a sequence of length a multiple of $l$.
@@
 
-In the iterative process, each chunks of size $l$ will generate $l$ blocks, and since $|C|$ divides $l$, the block referenced by each of these chunks will have the same number.
-If the rules begin with $a$, then this number will be $a$.
+In the iterative process, each subsequence of size $l$ ``generates" $l$ blocks at the next iteration, and since $|C|$ divides $l$, the block referenced by each of these subsequences will start with the same number the subsequence started with.
+Additionally, since all the rules begin with $a$, this number will be $a$.
 
-So applying the rules to each chunk is equivalent to applying the iterative process, where we  group elements into chunks of size $l$.
+So applying the rules to each chunk is equivalent to applying the iterative process, where we group elements into subsequences of size $l$.
 Therefore, $s^{(i+1)}$ is the result of applying the rules to $s^{(i)}$.
@@
 \begin{align*}
-\text{Rule }A: 2,2 &\to 2, 2, 4, 4 \\
-\text{Rule }B: 2,4 &\to 2, 2, 4,4,4,4\\
-\text{Rule }C: 4,2 &\to 2,2,2,2,4,4\\
-\text{Rule }D: 4,4 &\to 2, 2, 2, 2, 4, 4, 4, 4
+\text{Rule }1: 2,2 &\to 2, 2, 4, 4 \\
+\text{Rule }2: 2,4 &\to 2, 2, 4,4,4,4\\
+\text{Rule }3: 4,2 &\to 2,2,2,2,4,4\\
+\text{Rule }4: 4,4 &\to 2, 2, 2, 2, 4, 4, 4, 4.
 \end{align*}
 These rules satisfy the conditions of Proposition~\ref{ruleprop}, and so successive application of these rules will generate a self-referencing sequence which is weakly convergent to $s(C, 2)$.
+
+Unfortunately, this proposition will not help us resolve the density problem for $C = (12)$.
 
@@
 Suppose $l$ is a positive integer.
-Then $l$ must be even.
+Then, by our definition of $l$ in Proposition~\ref{ruleprop}, $l$ must be even.
 Let $s$ be the string with one 2 and $l-1$ 1's.
 Suppose $x \rightarrow y$ is the rule for $s$, so $x = s$.
-Then by condition 1 of Proposition~\ref{ruleprop}, $|y| = l+1$, and by condition 3, $l = 1$.
-Which contradicts the fact that $l$ is divisible by $|C| = 2$.
+Then by condition 1 of Proposition~\ref{ruleprop}, $|y| = l+1$, so $|y|$ is odd, but then it is not divisible by $l$, as condition 3 requires.
+This is a contradiction.
 \end{proof}
@@
 
-An open question regarding the self-referential sequence of 1's and 2's, $s((1 2),1)$, is what the limiting \emph{density} of 1's and 2's is.
-In this section, we consider the related problem of finding the density for sequences with generating cycle $C = (1 3)$ (note that the sequence that starts with a 1 differs by only one number from the sequence that starts with a 3, so we do not need to consider them separately).
+We consider the idea of density, which is the proportion of a self-referential sequence that is equal to a certain number.
+We build off of the substitution rules discovered in the previous section to study the densities in sequences with generating cycle $C = (1 3)$ (note that the sequence that starts with a 1 differs by only one number from the sequence that starts with a 3, so we do not need to consider them separately).
 While we cannot prove that a limiting density exists, we give a numerical calculation of what the density is \emph{if it exists}.
